---
title: "R Notebook"
output: html_notebook
---

# packages

```{r}
rm(list = ls())

library(ggplot2)
library(plyr)
library(MASS)
library(reshape)
library(ggpubr)
library(ggforce)

library(VGAM)
library(fitPS)

#library(rjson)

library(scico)

today = Sys.Date()
```

# functions

```{r}
# expected number of segregating sites
ES = function(N, mu, L, n){
  4*N*mu*L*sum( 1/(1:(n-1)) )
}

# reverse a string
strReverse <- function(x) sapply(lapply(strsplit(x, NULL), rev), paste, collapse="")

# HWE heterozygosity at a single site
diversity = function(p){
  1 - (p^2 + (1-p)^2)
}

# fucntion to calculate rates and accuracy
calc_rates = function(truth, test, sample_space){
  
  negatives = sample_space[!(sample_space %in% truth)]
  positives = truth
  
  true_positives = sum(test %in% positives)
  false_positives = sum(!(test %in% positives))
  
  true_negatives = sum(!(negatives %in% test))
  false_negatives = sum(!(positives %in% test))
  
  # accuracy
  acc = (true_positives + true_negatives)/(true_positives + true_negatives + false_positives + false_negatives)
  
  # true positives rate (sensitivity)
  tpr = true_positives/(true_positives + false_negatives)

  # false positive rate (false alarm)
  fpr = false_positives/(false_positives + true_negatives)

  # true negative rate (specificity)
  tnr = true_negatives/(true_negatives + false_positives)

  # false negative rate (miss)
  fnr = false_negatives/(true_positives + false_negatives)

  return(c(length(positives), length(negatives), true_positives, false_positives, true_negatives, false_negatives, acc, tpr, fpr, tnr, fnr))
}

#calc_rates(true_af$pos, varscan$pos, 1:1e6)

#calc_rates(true_af$pos, compare_maf_kmer$pos, 1:1e6)

# Posterior distribution of allele count i, given minor het-mer count X and total het-mer coverage Z
# n = pool size
# C = average k-mer coverage
# alpha and beta = prior distribution
posterior = function(x,z,n,C,alpha,beta){
  i=1:(n-1)
  p=i/n
  lambdax = i*C
  lambday = (n-i)*C
  
  likelihood_times_prior = (p^(x+alpha-1))*((1-p)^((z-x)+(beta-1)))/((exp(lambdax)-1)*(exp(lambday)-1))
  
  total_probability = sum(likelihood_times_prior)
  
  post = likelihood_times_prior/total_probability
  
  #return(which.max(post)/n)
  return(post)
}

# calculate posterior where the minor k-mer count is variable
truncation_constant = function(c, lambda){
    x = 0:(c-1)
    # truncation constants
    result = 1.0 - sum( (exp(-lambda)*(lambda^x))/factorial(x) )
    return(result)
}

posterior_min_kmer_count = function(x,z,n,C,c,alpha,beta){
  i=c:(n-c)
  #print(i)
  p=i/n
  #print(p)
  lambdax = i*C
  #print(lambdax)
  lambday = (n-i)*C
  #print(lambday)
  
  Tx = unlist(lapply(lambdax, FUN = truncation_constant, c = c))
  #print(Tx)
  Ty = unlist(lapply(lambday, FUN = truncation_constant, c = c))
  #print(Ty)
  
  likelihood_times_prior = (p^(x + alpha - 1))*(1-p)^((z-x) + (beta - 1)) / (Tx*Ty)
  
  total_probability = sum(likelihood_times_prior)
  
  post = likelihood_times_prior/total_probability
  
  return(which.max(post) + c - 1)
  #return(list(which.max(post) + c - 1, post))
}

#posterior_min_kmer_count(5, 200, 200, 1, 5, 1, 1)
```

# Simulations

## k-mer counts in pool-seq

```{r}

sterlings_approximation = function(n){
  sqrt(2*pi*n)*((n/exp(1))^n)
}

ln_ster_approx = function(n){
  n*log(n) - n
}

choose_ster_approx = function(z,x){
  z*log(z)-x*log(x)-(z-x)*log(z-x)  
}



pdf_2_ztp = function(z, i, n, N, L, k, G){
  C = ((L - k + 1)/L)*((N*L)/(n*G))

  lambx = i*C

  lamby = (n - i)*C
  
  x = 1:(z-1)
  
  #const = (1/((exp(lambx) - 1)*(exp(lamby) - 1)))

  #sum_terms = ((lambx^x)*(lamby^(z - x)))/(factorial(x)*factorial(z-x))

  const = (1/((exp(lambx) - 1)*(exp(lamby) - 1)*factorial(z)))
  
  #binom = unlist(lapply(x, combn, x = z))
  
  sum_terms = choose(z,x)*(lambx^x)*(lamby^(z - x))
  
  prob = const*sum(sum_terms)

  return(prob)
}


pdf_2_ztp(100, 5, 50, 1e6, 150, 31, 1e9)

sum(unlist(lapply(2:170, pdf_2_ztp, i = 5, n = 50, N = 1e9, L = 100, k = 31, G = 1e10)))

(200*G)/L

trials = 1e4
i = 30
j = 11
n = 100
L = 300
k = 31
N = 1.6e9
G = 1e9
theta = 12500

# coverage per genome
C = ((L - k + 1)/L)*((N*L)/(n*G))

# total coverage 
n*C

# number of bp sequenced
N*L

lambx = i*C
lamby = (n - i)*C

x_1 = rpois(trials, lambx)
y_1 = rpois(trials, lamby)

x_2 = rnbinom(trials, mu = lambx, size = theta)
y_2 = rnbinom(trials, mu = lamby, size = theta)

# 95 % confidence intervals for total coverage
upr = qpois(0.0005, n*C, lower.tail = F)
lwr = qpois(0.0005, n*C, lower.tail = T)

sum((lwr < (x_1 + y_1)) & ((x_1 + y_1) < upr)) / trials

# calculate overdispersion ratios
overx = (theta + lambx)/theta

overy = (theta + lamby)/theta

# to get a zero-truncated poisson, drop all zero observations

mean(x_1/(x_1+y_1))

truncate_1 = ( (x_1 != 0) & (y_1 != 0) )

x_1 = x_1[truncate_1]

y_1 = y_1[truncate_1]

truncate_2 = ( (x_2 != 0) & (y_2 != 0) )

x_2 = x_2[truncate_2]

y_2 = y_2[truncate_2]

#
mean(x_1/(x_1+y_1))

mean(x_2/(x_2+y_2))

i/n

p = x_1/(x_1+y_1)

p_hat = p*(1 - (p^(x_1+y_1)) - ((1-p)^(x_1+y_1)))

mean(p)
mean(p_hat)

#
plotdata = data.frame(
  index = 1:(n-1),
  post = posterior(x_1[2], x_1[2] + y_1[2], n, C, 1, 1)
)

ggplot(plotdata, aes(x = index, y = post)) +
  geom_bar(stat = "identity") +
  geom_vline(xintercept = i) +
  theme_classic()

# now calculate proportion
z_1 = sort(x_1/(x_1+y_1))

z_2 = sort(x_2/(x_2+y_2))

# see if beta distribution is a good approximation
# https://math.stackexchange.com/questions/153897/ratio-distribution-poisson-random-variables
# https://www.researchgate.net/publication/4983077_On_the_sum_of_independent_zero-truncated_Poisson_random_variables
w_1 <- ((1:length(z_1))-0.5)/length(z_1)

w_2 <- ((1:length(z_2))-0.5)/length(z_2)

s1_1 = lambx/(1-exp(-lambx))

s2_1 = lamby/(1-exp(-lamby))

s1_2 = lambx/(1-((theta/(lambx+theta))^theta))

s2_2 = lamby/(1-((theta/(lamby+theta))^theta))

qqplot(qbeta(w_1, s1_1, s2_1), z_1,
       xlab = paste("Beta (alpha = ", s1_1, ", beta = ", s2_1, ")", sep=""),
       ylab = paste("X/(X+Y) (Poisson)", sep = ""))
abline(c(0,1),lwd=2,col=2)

qqplot(qbeta(w_2, s1_2, s2_2), z_2,
       xlab = paste("Beta (alpha = ", s1_2, ", beta = ", s2_2, ")", sep=""),
       ylab = paste("X/(X+Y) (NBinom)", sep = ""))
abline(c(0,1),lwd=2,col=2)

qqplot(qnorm(w_1, mean = s1_1/(s1_1+s2_1), sd = sqrt( (s1_1*s2_1)/((s1_1+s2_1)^2 * (s1_1+s2_1+ 1)))), z_1,
       xlab = paste("Normal (mu = ", s1_1/(s1_1+s2_1), ", sd = ", sqrt( (s1_1*s2_1)/((s1_1+s2_1)^2 * (s1_1+s2_1+ 1))), ")", sep=""),
       ylab = paste("X/(X+Y) (Poisson)", sep = ""))
abline(c(0,1),lwd=2,col=2)

qqplot(qnorm(w_2, mean = s1_2/(s1_2+s2_2), sd = sqrt( (s1_2*s2_2)/((s1_2+s2_2)^2 * (s1_2+s2_2+ 1)))), z_2,
       xlab = paste("Normal (mu = ", s1_2/(s1_2+s2_2), ", sd = ", sqrt( (s1_2*s2_2)/((s1_2+s2_2)^2 * (s1_2+s2_2+ 1))), ")", sep=""),
       ylab = paste("X/(X+Y) (NBinom)", sep = ""))
abline(c(0,1),lwd=2,col=2)

# 95 % confidence intervals for allele frequency estimates
s1_1/(s1_1+s2_1) - 1.96*sqrt( (s1_1*s2_1)/((s1_1+s2_1)^2 * (s1_1+s2_1+ 1)))

s1_1/(s1_1+s2_1) + 1.96*sqrt( (s1_1*s2_1)/((s1_1+s2_1)^2 * (s1_1+s2_1+ 1)))

#
posterior = function(x,z,n,C,alpha,beta){
  i=1:(n-1)
  p=i/n
  lambdax = i*C
  lambday = (n-i)*C
  
  likelihood_times_prior = (p^(x+alpha-1))*((1-p)^((z-x)+(beta-1)))/((exp(lambdax)-1)*(exp(lambday)-1))
  
  total_probability = sum(likelihood_times_prior)
  
  post = likelihood_times_prior/total_probability
  
  return(post)
}

# given a discrete posterior probability distribution, calculate center credible interval
center_credible_interval = function(post, prob,n){
  k = which.max(post)
  total = post[k]
  i = 1
  while(total < prob){
    
    if(k-i < 0){
      left_post = 0
    } else{
      left_post = post[(k-i)]
    }
    
    if(k+i>n){
      right_post = 0
    } else{
      right_post = post[(k+i)]
    }
    
    total = total + left_post + right_post 
    i = i + 1
  }
  
  return(c(k-i,k,k+i))
}

results = NULL
for(w in 1:trials){
  post = posterior(x_1[w],(x_1+y_1)[w],n,C,1,1)

  mpe = which.max(post)
  
  #ci = center_credible_interval(post, 0.95, n)
  
  results = rbind(results, mpe)
}

hist(results, breaks = seq(from = 1, to = (n-1), by = 1))
abline(v = i)

table(results)
mean(results)


post = posterior(x_1[1],(x_1+y_1)[1],n,C,1,1)



#
#p_grid <- seq(from=0.001, to=0.999, length.out=10000)

#prior <- dbeta(p_grid, 2, 2)
#prior <- dbeta(p_grid, 1, 1)

x1 = 25
z1 = 183

x2 = 76
z2 = 187

p_grid = (1:(n-1))/(n)
prior = dbeta(p_grid,1,1)

likelihood1 <- dbinom(x1, size=z1, prob=p_grid)/(1 - p_grid^z1 - (1 - p_grid)^z1)
posterior1 <- likelihood1 * prior
posterior1 <- posterior1/sum(posterior1, na.rm = T)

#
x1 = 83
z1 = 197


likelihood = dbinom(x1, size=z1, prob=p_grid)/(1 - p_grid^z1 - (1 - p_grid)^z1)
total =  sum( dbinom(x1, size=z1, prob=p_grid)/(1 - p_grid^z1 - (1 - p_grid)^z1) )

posterior2 = (likelihood*prior)/total

plot(x = p_grid, y = posterior)
points(x = p_grid, y = posterior1, pch = 2)
abline(v = i/n, lty = 1)
abline(v = x1/z1, lty = 2)

p_grid[which.max(posterior)]

post_cdf = ecdf(posterior)

post_cdf(0.01)

quantile(posterior, prob = c(0.01, 0.99))

my_func = approxfun(p_grid, posterior)

plot(x = p_grid, y = my_func(p_grid))

my_func(0.5)

integrate(my_func, 0.01, 0.99)

trapezoid <- function(f, x) {
  if (is.function(f) == FALSE) {
    stop('f must be a function with one parameter (variable)')
  }
  total = 0
  for(i in 1:(length(x)-1)){
    a = x[i]
    b = x[(i + 1)]
    h <- b - a
    fxdx <- (h / 2) * (f(a) + f(b))
    total = total + fxdx
  }
  return(total)
}

trapezoid(my_func, p_grid)

# On average, does the maximum estimate posteriori estimate give the true value
results = c()
for(i in 1:100000){
  x2 = x_1[i]
  z2 = x_1[i] + y_1[i]
  likelihood2 <- dbinom(x2, size=z2, prob=p_grid)/(1 - p_grid^z2 - (1 - p_grid)^z2)
  posterior2 <- likelihood2 * prior
  posterior2 <- posterior2/sum(posterior2, na.rm = T)
  results = c(results, p_grid[which.max(posterior2)])
}

mean(results)
mean(x_1/(x_1+y_1))

posterior3 = dbeta(p_grid, x1, z1)

posterior4 = dbeta(p_grid, x2, z2)

p_grid[which.max(posterior1)]
p_grid[which.max(posterior2)]
p_grid[which.max(posterior3)]
p_grid[which.max(posterior4)]

plot(x = p_grid, y = posterior1)
points(x = p_grid, y = posterior3)

points(x = p_grid, y = posterior2)

plot(x = p_grid, y = posterior1*posterior2)

overlap(list(p1 = posterior1, p2 = posterior2), nbins = 1e6, type = "2")

# likelihood ratio test
p_null = (x1 + x2)/(z1+z2)

p1 = x1/z1
p2 = x2/z2 

null_likelihood = (dbinom(x1, size=z1, prob=p_null)/(1 - p_null^z1 - (1 - p_null)^z1))*(dbinom(x2, size=z2, prob=p_null)/(1 - p_null^z2 - (1 - p_null)^z2))

alt_likelihood = (dbinom(x1, size=z1, prob=p1)/(1 - p1^z1 - (1 - p1)^z1))*(dbinom(x2, size=z2, prob=p2)/(1 - p2^z2 - (1 - p2)^z2))

lk_ratio_test = -2*log(null_likelihood/alt_likelihood)

pchisq(lk_ratio_test, 1, lower.tail = F)

# z-test
ztest = (p1-p2)/sqrt(p_null*(1-p_null)*(1/(z1) + 1/(z2)))

pnorm(ztest)

pt(ztest, z1 + z2 -1)

# see if estimator is unbiased
plotdata = data.frame(p = p, p_hat = p_hat)
plotdata = melt(plotdata)
ggplot(data = plotdata, aes(y = value, color = variable)) +
  geom_boxplot() +
  geom_hline(yintercept = i/n) +
  theme_classic()

hist(rbeta(trials, s1_1, s2_1))

hist(rbeta(trials, s1_1, s2_1) - rbeta(trials, s1_1, s2_1))

# Use difference of proportions test
j = 2*i
C_j = 2*C

lambx_j = j*C_j
lamby_j = (n - j)*C_j

x_3 = rpois(trials, lambx_j)
y_3 = rpois(trials, lamby_j)

truncate_3 = ( (x_3 != 0) & (y_3 != 0) )

x_3 = x_3[truncate_3]

y_3 = y_3[truncate_3]

z_3 = x_3/(x_3+y_3)

hist(z_3 - z_1)

p_hat = (x_1 + x_3)/(x_1 + x_3 + y_1 + y_3)

test_statistics = (z_3-z_1)/sqrt(p_hat*(1-p_hat)*(1/(x_1 + y_1) + 1/(x_3 + y_3)))

hist(test_statistics)

# repeat for negative binomial
x_4 = rnbinom(trials, mu = lambx_j, size = theta)
y_4 = rnbinom(trials, mu = lamby_j, size = theta)

truncate_4 = ( (x_4 != 0) & (y_4 != 0) )

x_4 = x_4[truncate_4]

y_4 = y_4[truncate_4]

z_4 = x_4/(x_4+y_4)

p_hat = (x_2 + x_4)/(x_2 + x_4 + y_2 + y_4)

test_statistics = (z_4-z_2)/sqrt(p_hat*(1-p_hat)*(1/(x_2 + y_2) + 1/(x_4 + y_4)))

hist(test_statistics)



fitdist(z_2, "beta")

fit = fitdistr(z_2, "beta", start = list(shape1 = 100, shape2 = 100))

summary(fit)

plot(fit)

qbeta(0.025, s1_1, s2_1)

i/n

s1_1/(s1_1 + s2_1)

qbeta(0.975, s1_1, s2_1)

#
qbeta(0.025, s1_2, s2_2)

i/n

s1_2/(s1_2 + s2_2)

qbeta(0.975, s1_2, s2_2)

#
mean(x/(x + y))

i/n

#
mean(x)

(lambx)/(1 - exp(-1*lambx))

#
mean(y)

(lamby)/(1 - exp(-1*lamby))




var(x/(x + y))

var(x)/var(x + y)

total_cov = c()
predictions = c()
for(j in 1:trials){
  x1 = rpois(1, i*C)
  
  x2 = rpois(1, (n-i)*C)

  # iterate through all possible values of i (1 to n-1)
  # find value of i that gives highest probability of observing x1 and x2
  probs1 = dpois(x1, (1:(n-1))*C)
  
  probs2 = dpois(x2, ((n-1):1)*C)

  prediction = which.max(probs1*probs2)

  total_cov = c(total_cov, x1 + x2)
  predictions = c(predictions, prediction)
}

plotdata = data.frame(
  total_cov = total_cov,
  predictions = predictions
)

ggplot(plotdata, aes(y = predictions)) +
  geom_boxplot() +
  geom_hline(yintercept = i, lty = 2) +
  theme_classic()

ggplot(plotdata, aes(y = total_cov)) +
  geom_boxplot() +
  geom_hline(yintercept = n, lty = 2) +
  theme_classic()

# 95 % confidence interval for allele frequency (i)
quantile(plotdata$predictions, c(0.025, 0.975))

# 95 % confidence interval for total coverage (x1 + x2)
qpois(c(0.025, 0.975), i*C + (n-i)*C)

# test formulae for sum of two negative binomial random variables
rho = 15

x1 = rnbinom(trials, mu = i*C, size = rho)
x2 = rnbinom(trials, mu = (n-i)*C, size = rho)

var(x1) + var(x2)
var(x1 + x2)

n*C + ( (C^2)*( (n^2) - 2*i*n + 2*(i^2) ) )/rho

mean(x1/(x1 + x2))

i/n


# try to get maximum likelihood estimate
# https://stackoverflow.com/questions/59092976/using-optim-for-maximisation
log_likelihood <- function(par, obs) {
  dnbinom(obs, mu = par["mu"], size = par["size"])
}

op_res = optim(c(mu = 25, size = 5), fn = log_likelihood, obs = 43, lower = c(mu = 1, size = 0.01), upper = c(mu = 1000, size = 10000), control = list(fnscale = -1))

op_res$par

hist(rnbinom(10000, mu = op_res$par["mu"], size = op_res$par["size"]))
abline(v = 15)
```

## kmer fst

```{r}
i1 = 10
i2 = 70
n1 = 70
n2 = 70
NL1 = 2e8
NL2 = 1e8
G = 1e6

num = ( (i1/n1)*(NL1/G) + (i2/n2)*(NL2/G) )/2

denom = ((i1+i2)/(n1+n2))*((NL1+NL2)/G)

i1/n1

i2/n2

(i1+i2)/(n1+n2)

num

denom

1 - num/denom


```

## hetmer fst

```{r}

trials = 1e5
i1 = 15
i2 = 45
j = 30
n = 100
L = 150
k = 31
N = 1.6e9
G = 1e9
theta = 12500

# coverage per genome
C = ((L - k + 1)/L)*((N*L)/(n*G))

# total coverage 
n*C

# number of bp sequenced
N*L

lambx1 = i1*C
lamby1 = (n - i1)*C

x_1 = rpois(trials, lambx1)
y_1 = rpois(trials, lamby1)

lambx2 = i2*C
lamby2 = (n - i2)*C

x_2 = rpois(trials, lambx2)
y_2 = rpois(trials, lamby2)

# truncate
truncate_1 = ( (x_1 != 0) & (y_1 != 0) )

x_1 = x_1[truncate_1]

y_1 = y_1[truncate_1]

truncate_2 = ( (x_2 != 0) & (y_2 != 0) )

x_2 = x_2[truncate_2]

y_2 = y_2[truncate_2]

#
z_1 = x_1 + y_1

z_2 = x_2 + y_2

#
fst = 0.9

theta = (1/fst) - 1

rdirichlet(1, c( theta*j/n, theta*(n-j)/n) )

rbeta(1, theta*j/n, theta*(n-j)/n )

current_allele_freqs = matrix(c(i1/n, (n-i1)/n, i2/n, (n-i2)/n), nrow = 2, ncol = 2, byrow = T)

ancestral_allele_freq = seq(from = 0.001, to = 0.999, by = 0.001)

ddirichlet(current_allele_freqs, c( 10, 10) )

#
pgrid = seq(from = 0.001, to = 0.999, by = 0.001)

L_1 = dbinom(27, 177, pgrid)/(1- (pgrid)^177 - (pgrid)^177)

L_2 = dbinom(81, 202, pgrid)/(1- (pgrid)^202 - (pgrid)^202)

pgrid[which.max(L_1)]

pgrid[which.max(L_2)]
```

## Minimum coverage for minor allele frequency

How high does coverage need to be for 95 % of hetmers at a given minor allele frequency to have a kmer count above some threshold?

Even at 200 x coverage, fewer than 95 % of hetmers with minor kmer frequency < 0.01 will have a k-mer count less than 5

```{r}
i = 2
n = 200
alpha = 0.5
C = 300
L = 150
k = 31

results = expand.grid(
  maf = 1:(n/2),
  cov = seq(from = 50, to = 200, by = 5),
  cutoff = 5
)


results$cdf = qpois(alpha, results$maf/n*results$cov*(L-k+1)/L)

p = plot_ly(
    results, x= ~maf/n, y= ~cov, z= ~cdf,
    type='mesh3d', intensity = ~cdf,
    colors= colorRamp(rainbow(5))
  )

add_mesh(p, x= ~maf/n, y= ~cov, z = ~cutoff, data = results, opacity = 0.3, color = "grey", inherit = F)

```

# zeta distribution
```{r}

my_histos = list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/kmc_histo_results/", pattern = ".histo")


histos_to_plot = c("GCF_000001735.4_TAIR10.1_genomic_31.histo", "GCF_000001735.4_TAIR10.1_genomic_41.histo", "GCF_000001735.4_TAIR10.1_genomic_51.histo", "GCA_000941615.1_P_trichosuri_KNP_genomic_31.histo")

results = NULL
num_skipped = 0

for(one_histo in my_histos){
  # Initialize a flag to indicate if an error occurred in this iteration
  skip_to_next <- FALSE 
  
  histo = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/kmc_histo_results/", one_histo, sep = ""), header = F)
  
  #plot(log10(histo$V1), log10(histo$V2))
  
  data = histo$V2
  
  data = data[1:(pmin( min(which( data == 0 )) - 1, length(data)) )]
  
  p1 = makePSData(n = 1:length(data), count = data, type = "S")

  tryCatch(
    expr = {
      fit <- fitDist(p1)
    },
    error = function(e) {
      # If an error occurs, set the flag to TRUE
      message(paste("Error in iteration:", one_histo, e$message))
      num_skipped <<- num_skipped + 1
      skip_to_next <<- TRUE # Use superassignment to modify the flag in the enclosing environment
    }
  )
  
  if(skip_to_next){
    next
  }
  
  s <- fit$shape
  
  # genome length
  L = sum(data)
  
  prob_mass = data/L
  
  # calculate r-squared
  compare = data.frame(
      x = 1:length(data),
      data = prob_mass,
      zeta = dzeta(1:length(data), shape = s)
  )
    
    sse = sum( (compare$data - compare$zeta)^2 )
    sst = sum( (compare$data - mean(compare$data))^2 )
    
    r2 = 1 - sse/sst
  
  # plot some fits
  if(one_histo %in% histos_to_plot){
    
    print("Plotting...")
    
    compare = melt(compare, id = "x")

    plot_zeta = ggplot(compare, aes(x = log10(x), y = log10(value), group = variable)) +
      geom_point(shape = 16) +
      geom_line(aes(linetype = variable)) +
      theme_minimal() +
      theme(text = element_text(size = 20, 
                                family = "Helvetica", 
                                colour = "black"), 
        axis.text = element_text(colour = "black"),
        legend.title = element_blank()) +
      labs(x = "log10(k-mer frequency)", y = "log10(Probability mass)")
    
    ggsave(paste("../../results/", today, "/", gsub(".histo", "", one_histo), "_zeta_vs_data.pdf", sep = ""), device = cairo_pdf, width = 6, height = 4)
  }

  result = data.frame(
    genome = one_histo,
    perc_uniq_kmers = prob_mass[1],
    shape = s,
    length = L,
    r2 = r2
  )

  results = rbind(results, result)
}

# separate data by k
results$k = NA
results[grepl("_31.histo", results$genome),"k"] <- 31
results[grepl("_41.histo", results$genome),"k"] <- 41
results[grepl("_51.histo", results$genome),"k"] <- 51

# separate plants and animals
plants = read.table("../../data/embryophyta_ncbi_ids.tsv", fill = T, header = T, sep = "\t")

plants$Assembly.Accession = gsub("\\.[0-9]", "", plants$Assembly.Accession)

results$Assembly.Accession = gsub("\\.[0-9]_.*", "", results$genome)

results$kingdom = NA

results$kingdom[(results$Assembly.Accession %in% plants$Assembly.Accession)] = "Plants"

results$kingdom[is.na(results$kingdom)] = "Animals"

table(results$kingdom)

# get three values for sweep sims
summary(results[(results$k == 31), "shape"])
summary(results[(results$k == 41), "shape"])
summary(results[(results$k == 51), "shape"])

# label specific species to orient reader
extra_point_data <- data.frame(
      k = c(31, 31, 31,
            41, 41, 41,
            51, 51, 51,
            31, 31, 31,
            41, 41, 41,
            51, 51, 51),
      shape = c(1.772823, 2.957833, 3.946757,
                1.925577, 3.315291, 4.193795,
                2.066483, 3.577704, 4.382319,
                2.949517, 3.560309, 4.122031,
                3.068094, 3.566018, 4.292060,
                3.159185, 3.585911, 4.447315),
      species = c("maize", "tomato", "arabidopsis",
                  "maize", "tomato", "arabidopsis",
                  "maize", "tomato", "arabidopsis",
                  "human", "fly", "worm",
                  "human", "fly", "worm",
                  "human", "fly", "worm"),
      kingdom = c("Plants", "Plants", "Plants",
                  "Plants", "Plants", "Plants",
                  "Plants", "Plants", "Plants",
                  "Animals", "Animals", "Animals",
                  "Animals", "Animals", "Animals",
                  "Animals", "Animals", "Animals")
)

# distribution of shape parameter
ggplot(results, aes(x = as.factor(k), y = shape)) +
  geom_violin(lwd = 1.1) +
  facet_wrap(~ kingdom, ncol = 1) +
  geom_point(data = extra_point_data, mapping = aes(x = as.factor(k), y = shape, color = species), size = 3) +
  scale_color_manual(values = c("maize" = "yellow3", "tomato" = "pink2", "arabidopsis" = "green1", "human" = "black", "fly" = "grey40", "worm" = "grey90")) +
  theme_minimal() +
  theme(text = element_text(size = 20, family = "Helvetica", colour = "black"), 
        axis.text = element_text(colour = "black"), 
        legend.position = "None",
        axis.title.y = element_text(angle = 0, vjust = 0.5)) +
  labs(x = "k", y = "zeta shape parameter",title = paste("n = ", length(unique(results$Assembly.Accession)), sep = "")) +
  coord_flip()

ggsave(paste("../../results/", today, "/shape_dist.pdf", sep = ""), device = cairo_pdf, width = 6, height = 4)

# percent unique k-mers
ggplot(results, aes(x = as.factor(k), y = perc_uniq_kmers)) +
  geom_violin(lwd = 1.1) +
  facet_wrap(~ kingdom, ncol = 1) +
  theme_minimal() +
  theme(text = element_text(size = 20, family = "Helvetica", colour = "black"), 
        axis.text = element_text(colour = "black"), 
        legend.position = "None",
        axis.title.y = element_text(angle = 0, vjust = 0.5)) +
  labs(x = "k", y = "Percent uni-mers") +
  coord_flip()

# compare plants vs animals
median( results[(results$k == 31) & (results$kingdom == "Plants"), "shape"] )
median( results[(results$k == 31) & (results$kingdom == "Animals"), "shape"] )

wilcox.test(results[(results$k == 31) & (results$kingdom == "Plants"), "shape"], results[(results$k == 31) & (results$kingdom == "Animals"), "shape"])

# zeta model fits
summary(results$r2)

sum(results$r2 > 0.99)/nrow(results) 

```

# process single population metrics
```{r}
params = read.table("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/config/parameters.tsv", header = T)

dones = sort(as.numeric(gsub("_0", "", gsub(".done", "",list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/finished/", pattern = ".done")))))

# if looking at data before workflow is finished, may need to separate out all simulations that are done
#dones = read.table("/global/scratch/users/milesroberts/poolseq-kmers/workflow/dones.txt")
#dones = dones$V1

# empty frame to store results
results = NULL

# get list of simulation ids
# simids = params$ID[(params$simtype == "onepop") | (params$simtype == "sweep")]
# 
# two_pop_ids = expand.grid(params$ID[(params$simtype == "twopop")], c("p1", "p2"))
# 
# two_pop_ids = paste(two_pop_ids$Var1, two_pop_ids$Var2, sep = "_")
# 
# simids = c(two_pop_ids, simids)

#simids = params$ID[(params$simtype == "onepop")]

simids = params$ID[(params$simtype == "onepop") & (params$ID %in% dones)]

k = 31

thresh = 3

G = 2e6

#for(i in dones){
#for(i in params$ID[(params$simtype == "onepop") | (params$simtype == "sweep")]){
#for(i in 127:512){
for(i in simids){
  print(paste("Processing simulation ", i, "...", sep = ""))
  
# get sequencing coverage data
# json_data_R1R2 <- fromJSON(file=paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", i, "_R1R2.json", sep = ""))
# json_data_U1 <- fromJSON(file=paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", i, "_U1.json", sep = ""))
# json_data_U2 <- fromJSON(file=paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", i, "_U2.json", sep = ""))
# 
  n = params[(params$ID == i), "n"] # pool size
# 
# B = json_data_R1R2$summary$after_filtering$total_bases + json_data_U1$summary$after_filtering$total_bases + json_data_U2$summary$after_filtering$total_bases # number of bases sequenced
# 
# G = params[(params$ID == i), "L"] # genome size
# 
# L = (json_data_R1R2$summary$after_filtering$read1_mean_length*json_data_R1R2$summary$after_filtering$total_reads/2 + json_data_R1R2$summary$after_filtering$read2_mean_length*json_data_R1R2$summary$after_filtering$total_reads/2 + json_data_U1$summary$after_filtering$read1_mean_length*json_data_U1$summary$after_filtering$total_reads + json_data_U2$summary$after_filtering$read1_mean_length*json_data_U2$summary$after_filtering$total_reads)/(json_data_R1R2$summary$after_filtering$total_reads + json_data_U1$summary$after_filtering$total_reads + json_data_U2$summary$after_filtering$total_reads) # read length, paired end data, multiply read length by 2
# 
# C = B/(2*n*G) # base coverage per genome
# Ck = ((L-k+1)/L)*C # kmer coverage per genome
# nCk = n*Ck # total kmer coverage
# nC = n*C # total base coverage

# load pairs of k-mers
kmer_pairs = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/smudgeplot_results/", i, "_0_sequences.tsv", sep = ""), col.names = c("k1"))

kmer_covs = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/smudgeplot_results/", i, "_0_coverages.tsv", sep = ""), col.names = c("c1", "c2"))
kmer_pairs = cbind(kmer_pairs, kmer_covs)

# choose kmer count cutoff
# all_counts = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/kmer_counts_1.txt", col.names = c("k1", "c1"))
# 
# ggplot(all_counts, aes(x = c1)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# 
# ggplot(all_counts[(all_counts$c1 >= 4),], aes(x = c1)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()

# calculate minor allele frequency
kmer_pairs$kmer_maf = pmin(kmer_pairs$c1, kmer_pairs$c2)/(kmer_pairs$c1 + kmer_pairs$c2)

# filter kmers
# high_nc = qnorm(0.99995, mean = nC, sd = sqrt(nC))
# low_nc = qnorm(0.00005, mean = nC, sd = sqrt(nC))
# 
# high_nck = qnorm(0.99995, mean = nCk, sd = sqrt(nCk))
# low_nck = qnorm(0.00005, mean = nCk, sd = sqrt(nCk))
# 
# ggplot(kmer_covs, aes(x = c1 + c2)) +
#   geom_density() +
#   geom_vline(xintercept = 2*thresh) +
#   geom_vline(xintercept = low_nc, lty = 2, color = "red") +
#   geom_vline(xintercept = nC, color = "red") +
#   geom_vline(xintercept = high_nc, lty = 2, color = "red") +
#   geom_vline(xintercept = low_nck, lty = 2, color = "black") +
#   geom_vline(xintercept = nCk, color = "black") +
#   geom_vline(xintercept = high_nck, lty = 2, color = "black")

#filter = (kmer_pairs$c1 + kmer_pairs$c2 > low_nck) & (kmer_pairs$c1 + kmer_pairs$c2 < high_nck) & (kmer_pairs$c1 >= thresh) & (kmer_pairs$c2 >= thresh)

#kmer_pairs = kmer_pairs[(kmer_pairs$c1 >= thresh) & (kmer_pairs$c2 >= thresh),]

# use bayes theorem to estimate allele counts
#kmer_pairs$kmer_maf_bayes = mapply(posterior, x = kmer_pairs$c1, z = kmer_pairs$c1 + kmer_pairs$c2, n, C, 1, 1)

# ggplot(kmer_pairs, aes(x = kmer_maf)) +
#   geom_density() +
#   theme_classic()

# load from my hetmers script
hetmers_seqs = read.csv(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/hetmers_results/", i, "_0_seqs.csv", sep = ""), col.names = c("k1", "k2"))

hetmers_counts = read.csv(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/hetmers_results/", i, "_0_counts.csv", sep = ""), col.names = c("c1", "c2"))

hetmers_bayes = read.csv(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/hetmers_results/", i, "_0_bayes_states.csv", sep = ""), col.names = c("bc"))

hetmers = cbind(hetmers_seqs, hetmers_counts, hetmers_bayes)

# calculate minor allele counts/frequencies
hetmers$hetmer_maf = pmin(hetmers$c1, hetmers$c2)/(hetmers$c1 + hetmers$c2)

hetmers$hetmer_maf_bayes = pmin(hetmers$bc, 2*n - hetmers$bc)/(2*n)

# load in actual allele freqs from simulations
true_af = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/slim_results/allele_freqs_", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"))

true_af$ac = as.numeric(true_af$ac)

true_af = true_af[complete.cases(true_af),]

# allele frequency spectrum
# ggplot(true_af, aes(x = ac)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()

# calculate minor allele frequencies
true_af$af = as.numeric(true_af$af)

true_af$snp_maf = pmin(true_af$af, 1 - true_af$af)

true_af$snp_mac = pmin(true_af$ac, 2*n - true_af$ac)

true_af = true_af[(true_af$af < 1) & (true_af$af > 0),]

# load discosnp results, calculate maf
disco = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/disco_results/", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "c1", "c2"))

disco = disco[(disco$c1 >= thresh) & (disco$c2 >= thresh),]

disco$disco_maf = pmin(disco$c1, disco$c2)/(disco$c1 + disco$c2)

# get key for position to k-mers
# if(grepl("_p1|_p2", i)){
#   j = gsub("_p1|_p2", "_p1p2", i)
# } else {
#   j = i
# }

pos_kmer_key = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/seqkit_results/center_kmer_pairs_", i, ".txt", sep = ""), fill = T, col.names = c("pos", "k1", "k2", "k3", "k4"))

pos_kmer_key$leftk = substr(pos_kmer_key$k1, 1, 15)
pos_kmer_key$rightk = substr(pos_kmer_key$k1, 17, 31)

# compare snp and k-mer mafs
kmer_pairs$leftk = substr(kmer_pairs$k1, 1, 15)
kmer_pairs$rightk = substr(kmer_pairs$k1, 17, 31)

hetmers$leftk = substr(hetmers$k1, 1, 15)
hetmers$rightk = substr(hetmers$k1, 17, 31)

pos_kmer_pairs = merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

pos_hetmers = merge(pos_kmer_key, hetmers, by = c("leftk", "rightk"))

# get snps that pair with reverse complements
pos_kmer_key$leftk_rc = chartr("ATGC","TACG",pos_kmer_key$leftk)
pos_kmer_key$rightk_rc = chartr("ATGC","TACG",pos_kmer_key$rightk)

pos_kmer_key$leftk_rc = strReverse(pos_kmer_key$leftk_rc)
pos_kmer_key$rightk_rc = strReverse(pos_kmer_key$rightk_rc)

pos_kmer_pairs_rc = merge(pos_kmer_key, kmer_pairs, by.x = c("leftk_rc", "rightk_rc"), by.y = c("rightk", "leftk"))

pos_hetmers_rc = merge(pos_kmer_key, hetmers, by.x = c("leftk_rc", "rightk_rc"), by.y = c("rightk", "leftk"))

pos_kmer_pairs_rc = subset(pos_kmer_pairs_rc, select=-c(leftk,rightk))

pos_hetmers_rc = subset(pos_hetmers_rc, select=-c(leftk,rightk))

# merge reverse complement and regular sets
names(pos_kmer_pairs_rc)[1:2] = c("leftk", "rightk")
pos_kmer_pairs = rbind(pos_kmer_pairs, pos_kmer_pairs_rc)

names(pos_hetmers_rc)[1:2] = c("leftk", "rightk")
pos_hetmers = rbind(pos_hetmers, pos_hetmers_rc)

# do some filtering
# remove low coverage k-mer pairs
# remove tri-allelic k-mers
#pos_kmer_pairs = pos_kmer_pairs[(pos_kmer_pairs$c1 >=4 & pos_kmer_pairs$c2 >= 4),]
pos_kmer_pairs = pos_kmer_pairs[( (pos_kmer_pairs$k3 == "") | is.na(pos_kmer_pairs$k3)),]
pos_hetmers = pos_hetmers[( (pos_hetmers$k3 == "") | is.na(pos_hetmers$k3)),]

compare_maf_kmer = merge(true_af[,c("pos", "snp_maf", "snp_mac")], pos_kmer_pairs, by = "pos", all.x = T, all.y = T)

compare_maf_hetmer = merge(true_af[,c("pos", "snp_maf", "snp_mac")], pos_hetmers, by = "pos", all.x = T, all.y = T)

# remove all na rows
#compare_maf_kmer = compare_maf_kmer[rowSums(is.na(compare_maf_kmer)) != ncol(compare_maf_kmer), ]

# compare minor allele counts
#compare_maf_kmer_no_na = compare_maf_kmer[!is.na(compare_maf_kmer$c1),]

#compare_maf_kmer_no_na$kmer_mac = unlist(mapply(posterior_min_kmer_count, x = compare_maf_kmer_no_na$c1, z = compare_maf_kmer_no_na$c1 + compare_maf_kmer_no_na$c2, MoreArgs = list(n = 100, C = 2*0.9, c = 5, alpha = 1, beta = 1)))

# include varscan calls
varscan = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/varscan_results/", i, "_0.tsv", sep = ""), header = T)

varscan = varscan[(varscan$Reads1 >= thresh) & (varscan$Reads2 >= thresh),]

varscan$snp_maf_varscan = pmin(varscan$Reads1, varscan$Reads2)/(varscan$Reads1 + varscan$Reads2)

names(varscan)[2] = "pos"

# load poolsnp calls
poolsnp = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/poolsnp_results/", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "c1", "t"), fill = T, na.strings = "")

poolsnp = poolsnp[(poolsnp$c1 >= thresh),]

poolsnp$snp_maf_poolsnp = pmin(poolsnp$c1/poolsnp$t, 1 - poolsnp$c1/poolsnp$t)

# include freebayes calls
freebayes = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freebayes_de_novo_results/", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"), fill = T, na.strings = "")

freebayes$snp_maf_freebayes = pmin(freebayes$af, 1 - freebayes$af)

#freebayes = freebayes[(freebayes$af > 0) & (freebayes$af < 1),]

# compare varscan and ground truth
compare_maf_varscan = merge(true_af[,c("pos", "snp_maf")], varscan, by = "pos", all.x = T, all.y = T)

# compare poolsnp and ground truth
compare_maf_poolsnp = merge(true_af[,c("pos", "snp_maf")], poolsnp, by = "pos", all.x = T, all.y = T)

# compare freebayes to ground truth
compare_maf_freebayes = merge(true_af[,c("pos", "snp_maf")], freebayes, by = "pos", all.x = T, all.y = T)

# compare discosnp and ground truth
compare_maf_disco = merge(true_af[,c("pos", "snp_maf")], disco, by = "pos", all.x = T, all.y = T)

# compare k-mer and varscan
#compare_maf_kmer_varscan = merge(compare_maf_kmer, varscan, by = "pos")

# calculate correlations and performance stats
kmer_vs_truth_maf_cor = cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$kmer_maf, method = "pearson")$estimate

hetmer_vs_truth_maf_cor = cor.test(compare_maf_hetmer$snp_maf, compare_maf_hetmer$hetmer_maf, method = "pearson")$estimate

hetmer_vs_truth_bc_cor = cor.test(compare_maf_hetmer$snp_mac, compare_maf_hetmer$bc, method = "pearson")$estimate

disco_vs_truth_maf_cor = tryCatch(
  cor.test(compare_maf_disco$snp_maf, compare_maf_disco$disco_maf, method = "pearson")$estimate, 
  error=function(err) NA)

varscan_vs_truth_maf_cor = cor.test(compare_maf_varscan$snp_maf, compare_maf_varscan$snp_maf_varscan, method = "pearson")$estimate

poolsnp_vs_truth_maf_cor = tryCatch(
  cor.test(compare_maf_poolsnp$snp_maf, compare_maf_poolsnp$snp_maf_poolsnp, method = "pearson")$estimate, 
  error=function(err) NA)

freebayes_vs_truth_maf_cor = cor.test(compare_maf_freebayes$snp_maf, compare_maf_freebayes$snp_maf_freebayes, method = "pearson")$estimate

varscan_rates = calc_rates(true_af$pos, varscan$pos, 1:G)

poolsnp_rates = calc_rates(true_af$pos, poolsnp$pos, 1:G)

freebayes_rates = calc_rates(true_af$pos, freebayes$pos, 1:G)

kmer_rates = calc_rates(true_af$pos, c(pos_kmer_pairs$pos,rep(1e9, times = nrow(kmer_pairs) - nrow(pos_kmer_pairs))), 1:G)

hetmer_rates = calc_rates(true_af$pos, c(pos_hetmers$pos,rep(1e9, times = nrow(hetmers) - nrow(pos_hetmers))), 1:G)

disco_rates = calc_rates(true_af$pos, disco$pos, 1:G)

# save all results
result = data.frame(
  ID = gsub("_p1|_p2", "", i),
  ID2 = i,
  true_div = sum(diversity(true_af$snp_maf))/G, # average nucleotide diversity
  kmer_div = sum(diversity(kmer_pairs$kmer_maf))/G,
  kmer_vs_truth_maf_cor = kmer_vs_truth_maf_cor,
  kmer_pos = kmer_rates[1],
  kmer_neg = kmer_rates[2],
  kmer_tp = kmer_rates[3],
  kmer_fp = kmer_rates[4],
  kmer_tn = kmer_rates[5],
  kmer_fn = kmer_rates[6],
  kmer_acc = kmer_rates[7],
  kmer_tpr = kmer_rates[8],
  kmer_fpr = kmer_rates[9],
  kmer_tnr = kmer_rates[10],
  kmer_fnr = kmer_rates[11],
  hetmer_div = sum(diversity(hetmers$hetmer_maf))/G,
  hetmer_vs_truth_maf_cor = hetmer_vs_truth_maf_cor,
  hetmer_vs_truth_bc_cor = hetmer_vs_truth_bc_cor,
  hetmer_pos = hetmer_rates[1],
  hetmer_neg = hetmer_rates[2],
  hetmer_tp = hetmer_rates[3],
  hetmer_fp = hetmer_rates[4],
  hetmer_tn = hetmer_rates[5],
  hetmer_fn = hetmer_rates[6],
  hetmer_acc = hetmer_rates[7],
  hetmer_tpr = hetmer_rates[8],
  hetmer_fpr = hetmer_rates[9],
  hetmer_tnr = hetmer_rates[10],
  hetmer_fnr = hetmer_rates[11],
  disco_div = sum(diversity(disco$disco_maf))/G,
  disco_vs_truth_maf_cor = disco_vs_truth_maf_cor,
  disco_pos = disco_rates[1],
  disco_neg = disco_rates[2],
  disco_tp = disco_rates[3],
  disco_fp = disco_rates[4],
  disco_tn = disco_rates[5],
  disco_fn = disco_rates[6],
  disco_acc = disco_rates[7],
  disco_tpr = disco_rates[8],
  disco_fpr = disco_rates[9],
  disco_tnr = disco_rates[10],
  disco_fnr = disco_rates[11],
  varscan_div = sum(diversity(varscan$snp_maf_varscan))/G,
  varscan_vs_truth_maf_cor = varscan_vs_truth_maf_cor,
  varscan_pos = varscan_rates[1],
  varscan_neg = varscan_rates[2],
  varscan_tp = varscan_rates[3],
  varscan_fp = varscan_rates[4],
  varscan_tn = varscan_rates[5],
  varscan_fn = varscan_rates[6],
  varscan_acc = varscan_rates[7],
  varscan_tpr = varscan_rates[8],
  varscan_fpr = varscan_rates[9],
  varscan_tnr = varscan_rates[10],
  varscan_fnr = varscan_rates[11],
  poolsnp_div = sum(diversity(poolsnp$snp_maf_poolsnp))/G,
  poolsnp_vs_truth_maf_cor = poolsnp_vs_truth_maf_cor,
  poolsnp_pos = poolsnp_rates[1],
  poolsnp_neg = poolsnp_rates[2],
  poolsnp_tp = poolsnp_rates[3],
  poolsnp_fp = poolsnp_rates[4],
  poolsnp_tn = poolsnp_rates[5],
  poolsnp_fn = poolsnp_rates[6],
  poolsnp_acc = poolsnp_rates[7],
  poolsnp_tpr = poolsnp_rates[8],
  poolsnp_fpr = poolsnp_rates[9],
  poolsnp_tnr = poolsnp_rates[10],
  poolsnp_fnr = poolsnp_rates[11],
  freebayes_div = sum(diversity(freebayes$snp_maf_freebayes))/G,
  freebayes_vs_truth_maf_cor = freebayes_vs_truth_maf_cor,
  freebayes_pos = freebayes_rates[1],
  freebayes_neg = freebayes_rates[2],
  freebayes_tp = freebayes_rates[3],
  freebayes_fp = freebayes_rates[4],
  freebayes_tn = freebayes_rates[5],
  freebayes_fn = freebayes_rates[6],
  freebayes_acc = freebayes_rates[7],
  freebayes_tpr = freebayes_rates[8],
  freebayes_fpr = freebayes_rates[9],
  freebayes_tnr = freebayes_rates[10],
  freebayes_fnr = freebayes_rates[11]
)

results = rbind(results, result)

rm(compare_maf_kmer, compare_maf_varscan, compare_maf_poolsnp, poolsnp, poolsnp_rates, compare_maf_disco, disco, disco_rates, kmer_covs, kmer_pairs, pos_kmer_key, pos_kmer_pairs, pos_kmer_pairs_rc, true_af, varscan, kmer_rates, kmer_vs_truth_maf_cor, poolsnp_vs_truth_maf_cor, varscan_rates, varscan_vs_truth_maf_cor, disco_vs_truth_maf_cor, result, hetmer_rates, hetmers, hetmers_seqs, hetmers_counts, hetmer_vs_truth_maf_cor, hetmers_bayes)

}

pares = merge(params, results, by = "ID")

write.table(pares, paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/results/", today, "/performance_metrics.txt",sep = ""), sep = " ")
```

# site frequency spectrum
```{r}
# compare accuracy of allele frequency estimates
plotdata_af_cor = melt(pares[,c("ID", "cov", "n", "kmer_vs_truth_maf_cor", "hetmer_vs_truth_maf_cor", "disco_vs_truth_maf_cor", "varscan_vs_truth_maf_cor", "poolsnp_vs_truth_maf_cor")], id = c("ID", "cov", "n"))

plotdata_af_cor$variable = gsub("_vs_truth_maf_cor", "", plotdata_af_cor$variable)

plotdata_af_cor$variable[plotdata_af_cor$variable == "kmer"] = "smudgeplot"

ggplot(plotdata_af_cor, aes(x = as.factor(cov), color = variable, y = value)) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Coverage", y = "Cor. bw true and estimated maf", color = "method")

ggsave(paste("../results/", today, "/maf_cor.png"))

# compare diversity estimates
#melt(pares[,c("ID", "cov", "n", "true_div", "kmer_div", "disco_div", "varscan_div", "poolsnp_div")], id = c("ID", "cov", "n"))

# percent error on diversity
# percent_error = function(truth, measurement){
#   (measurement - truth)/truth * 100
# }
# 
# pares$pe_div_kmer = percent_error(pares$true_div, pares$kmer_div)
# 
# pares$pe_div_hetmer = percent_error(pares$true_div, pares$hetmer_div)
# 
# pares$pe_div_disco = percent_error(pares$true_div, pares$disco_div)
# 
# pares$pe_div_varscan = percent_error(pares$true_div, pares$varscan_div)
# 
# pares$pe_div_poolsnp = percent_error(pares$true_div, pares$poolsnp_div)

plotdata_div = melt(pares[,c("ID", "cov", "n", "true_div","kmer_div", "hetmer_div", "disco_div", "varscan_div", "poolsnp_div")], id = c("ID", "cov", "n", "true_div"))

plotdata_div$variable = gsub("_div", "", plotdata_div$variable)

plotdata_div$variable[plotdata_div$variable == "kmer"] = "smudgeplot"

plotdata_div = plotdata_div[(plotdata_div$variable %in% c("hetmer", "varscan", "disco")),]

# ggplot(plotdata_div, aes(x = as.factor(cov), y = value, color = as.factor(variable))) +
#   geom_boxplot() +
#   theme_classic()

ggplot(plotdata_div, aes(x = log10(true_div), y = log10(value), color = as.factor(cov), shape = as.factor(variable))) +
  geom_point(size = 2) +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  theme(text = element_text(size = 16, family = "Helvetica", color = "black"),
        axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black")) +
  scale_colour_scico_d(palette = "hawaii", name = "Coverage") +
  labs(y = "Estimated diversity", x = "True diversity", color = "Coverage", shape = "Method")

ggsave(paste("../../results/", today, "/estimated_vs_true_diversity.pdf", sep = ""), height = 6, width = 7)

# ggplot(pares, aes(x = log10(true_div), y = log10(hetmer_div), color = as.factor(cov))) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic()
# 
# ggplot(pares, aes(x = log10(true_div), y = log10(disco_div), color = as.factor(cov))) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic()
# 
# ggplot(pares, aes(x = log10(true_div), y = log10(varscan_div), color = as.factor(cov))) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic()
# 
# ggplot(pares, aes(x = log10(true_div), y = log10(poolsnp_div), color = as.factor(cov))) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic()
# 
# # site-frequency spectrum
# all_sfs = merge(compare_maf_kmer[,c("pos", "snp_maf", "kmer_maf")], varscan[,c("pos", "snp_maf_varscan")], by = "pos", all.x = T, all.y = T)
# 
# all_sfs = merge(all_sfs, poolsnp[,c("pos", "snp_maf_poolsnp")], by = "pos", all.x = T, all.y = T)
# 
# all_sfs = merge(all_sfs, disco[,c("pos", "disco_maf")], by = "pos", all.x = T, all.y = T)
# 
# plotdata = melt(all_sfs, id = "pos")
# 
# ggplot(plotdata, aes(x = value, fill = variable)) +
#   geom_density(alpha = 0.5) +
#   theme_classic() +
#   labs(x = "Minor allele frequency", y = "density")
# 
# ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","kmer_vs_snp_sfs.png", sep = ""), height= 7, width = 8)
```

# check if theoretical distributions fit kmer counts
```{r}
n = 200
k = 31
R = 301
G = 1e6
B = 130.641224e6

C = B/(n*G) 
  
Ck = ((R-k+1)/R)*C
nCk = n*Ck
nC = n*C
  
high = qnorm(0.995, mean = nCk, sd = sqrt(nCk))
low = qnorm(0.005, mean = nCk, sd = sqrt(nCk))

ggplot(kmer_covs, aes(x = c1 + c2)) +
  geom_density() +
  geom_vline(xintercept = low, lty = 2) +
  geom_vline(xintercept = nCk) +
  geom_vline(xintercept = high, lty = 2)

high = qnorm(0.995, mean = nC, sd = sqrt(nC))
low = qnorm(0.005, mean = nC, sd = sqrt(nC))

ggplot(kmer_covs, aes(x = c1 + c2)) +
  geom_density() +
  geom_vline(xintercept = low, lty = 2) +
  geom_vline(xintercept = nC) +
  geom_vline(xintercept = high, lty = 2)


thresh = 4

filter = (kmer_covs$c1 + kmer_covs$c2 > low) & (kmer_covs$c1 + kmer_covs$c2 < high) & (kmer_covs$c1 >= thresh) & (kmer_covs$c2 >= thresh)

filter = (kmer_covs$c1 >= thresh) & (kmer_covs$c2 >= thresh)

kmer_covs[filter,]


ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point()

ggplot(kmer_covs[(kmer_covs$c1 + kmer_covs$c2 > low) & (kmer_covs$c1 + kmer_covs$c2 < high),], aes(x = c1/(c1 + c2))) +
  geom_density()

ggplot(compare_maf_kmer, aes(x = kmer_maf)) +
  geom_histogram()

ggplot(true_af, aes(x = snp_maf)) +
  geom_histogram()

  kmer_covs
```

# selective sweeps
```{r}
# Hypothesis: the average distance between SNPs is higher after selective sweep compared to neutral scenario
allele_file_names = list.files(path = "/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_results", pattern = "allele_freqs_.*", full.names = T)

allele_file_names = allele_file_names[!grepl("_p.", allele_file_names)]

allele_file_ids = as.numeric(gsub(".txt", "", gsub("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_results/allele_freqs_", "", allele_file_names)))

allele_files = lapply(allele_file_names, read.table, col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"))

mean_dist_bw_mut = function(x){
  mean(diff(x$pos))
}

max_dist_bw_mut = function(x){
  max(diff(x$pos))
}

# remove invariant sites
invar_check = function(x){
  x[( (x$af < 1) & (x$af > 0) ), ]
}

allele_files = lapply(allele_files, invar_check)

# length of invariant site run at sweep site
dist_at_sweep_site = function(x, L){
  # find where distance to sweep site switches from positive to negative
  top_idx = which(diff(sign(x$pos - L/2)) != 0)
  
  # calculate difference between two positions
  result = x$pos[(top_idx + 1)] - x$pos[(top_idx)]
  
  return(result)
}

plotdata = data.frame(
  ID = allele_file_ids,
  mean_dist = unlist(lapply(allele_files, mean_dist_bw_mut)),
  max_dist = unlist(lapply(allele_files, max_dist_bw_mut)),
  #sweep_site_dist = unlist(lapply(allele_files, dist_at_sweep_site, L = 2e6)),
  S = unlist(lapply(allele_files, nrow))
)

plotdata = merge(params, plotdata, by = "ID")

# expected mean unitig length
plotdata$es = mapply(ES, plotdata$N, plotdata$mu, plotdata$L, plotdata$n)

plotdata$Ubar = plotdata$L/plotdata$es

# Does the expected mean unitig length match reality for neutral simulations
ggplot(plotdata[(plotdata$simtype == "onepop"),], aes(x = es, y = S)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic()

plot_mean_dist_neutral = ggplot(plotdata[(plotdata$simtype == "onepop"),], aes(x = Ubar, y = mean_dist, color = as.factor(mu))) +
  geom_point(aes(shape = as.factor(n))) +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "Expected mean distance between mutations", y = "Observed mean distance between mutations")

#
plotdata$n = as.factor(plotdata$n)
plotdata$cov = as.factor(plotdata$cov)
plotdata$s = as.factor(plotdata$s)

# expected mean unitig length
# plotdata$Ubar = plotdata$L/plotdata$S
# 
# # parameter for geometric distribution
# plotdata$p = 1/plotdata$Ubar
# 
# # 95 percentile for geometric distribution
# plotdata$geom_prob = pgeom(plotdata$max_dist, plotdata$p, lower.tail = F, log.p = T)

#plotdata$simtype[(plotdata$simtype == "onepop")] = "neutral"

# how do selective sweeps change distances between snps?
ggplot(plotdata, aes(x = n, y = mean_dist, color = as.factor(Nes))) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Pool size", y = "Avg. distance between SNPs (bp)", color = "Selection", title = "Ground truth")

plot_max_dist_sweep = ggplot(plotdata, aes(x = n, y = max_dist, color = as.factor(Nes))) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Pool size", y = "Max distance between SNPs (bp)", color = "Selection", title = "Ground truth")
    
# ggplot(plotdata, aes(x = n, y = sweep_site_dist/mean_dist, color = s)) +
#   geom_boxplot() +
#   theme_classic() +
#   geom_hline(yintercept = 1) +
#   labs(x = "Pool size", y = "Distance between snps at sweep/mean distance", color = "Selection", title = "Ground truth")

# Hypothesis: after a selective sweep, the longest unitigs in a population graph are more likely to be near the sweep site, compared to neutral scenario
unitig_files = list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/blast_results", pattern = ".*txt", full.names = T)

#unitig_files = sample(unitig_files, size = 1000)

unitig_files_ids = as.numeric(gsub("_0.txt", "", gsub("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/blast_results/", "", unitig_files)))

unitig_files = lapply(unitig_files, read.table, col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

#unitig_files = lapply(unitig_files, fread, col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

max_len = function(x){
  max(x$alen)
}

mean_len = function(x){
  mean(x$alen)
}

ss_max_len = function(x){
  x$ss[which.max(x$alen)]
}

se_max_len = function(x){
  x$se[which.max(x$alen)]
}

plotdata_unitigs = data.frame(
  ID = unitig_files_ids,
  max_len = unlist(lapply(unitig_files, max_len)),
  mean_len = unlist(lapply(unitig_files, mean_len)),
  ss = unlist(lapply(unitig_files, ss_max_len)),
  se = unlist(lapply(unitig_files, se_max_len))
)

plotdata_unitigs = merge(params, plotdata_unitigs, by = "ID")

# expected mean unitig length
plotdata_unitigs$es = mapply(ES, plotdata_unitigs$N, plotdata_unitigs$mu, plotdata_unitigs$L, plotdata_unitigs$n)

plotdata_unitigs$Ubar = plotdata_unitigs$L/plotdata_unitigs$es

ggplot(plotdata_unitigs[(plotdata_unitigs$simtype == "onepop"),], aes(x = Ubar, y = mean_len)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "Expected mean distance between mutations", y = "Observed mean unitig lengths")

#mod = lm(log10(max_len) ~ n + cov + sequencer + s + h, data = plotdata_unitigs)
#summary(mod)
#plot(mod)

plotdata_unitigs$n = as.factor(plotdata_unitigs$n)
plotdata_unitigs$cov = as.factor(plotdata_unitigs$cov)
plotdata_unitigs$s = as.factor(plotdata_unitigs$s)

#plotdata_unitigs$simtype[(plotdata_unitigs$simtype == "onepop")] = "neutral"

#plotdata = plotdata[( ((plotdata$s == 0.5) & (plotdata$h == 1)) | (plotdata$s == 0) ),]

ggplot(plotdata_unitigs, aes(x = n, y = max_len, color = simtype, fill = cov)) +
  geom_boxplot() +
  #facet_grid(sequencer ~ .) +
  theme_classic() +
  labs(x = "Pool size", y = "Maximum unitig length (bp)", color = "Simulation\ntype", fill = "Coverage")
  #stat_compare_means(aes(group = simtype), label.sep = "\n")

# see if long unitigs appear toward the sweep region
plotdata_unitigs$sc = (plotdata_unitigs$ss + plotdata_unitigs$se)/2

plot_unitigs_all_sims = ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  geom_vline(xintercept = 2e6/2, lty = 2, lwd = 1.) +
  labs(x = "Center of longest unitig", color = "Nes", title = "All")

plot_unitigs_cov_breakdown = ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  facet_grid(cov ~ .) +
  geom_vline(xintercept = 2e6/2, lty = 2, lwd = 1.) +
  labs(x = "Center of longest unitig", color = "Nes", title = "Coverage")

plot_unitigs_n_breakdown = ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  facet_grid(n ~ .) +
  geom_vline(xintercept = 2e6/2, lty = 2, lwd = 1) +
  labs(x = "Center of longest unitig", color = "Nes", title = "Pool size")

plot_unitigs_shape_breakdown = ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density(lwd = 1.25) +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica", color = "black"),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black")) +
  facet_grid(shuffle ~ shape ) +
  scale_color_scico_d(palette = "buda") +
  #facet_grid(shape ~ .) +
  geom_vline(xintercept = 2e6/2, lty = 2, lwd = 1) +
  labs(x = "Center of longest unitig", color = "Nes")

ggsave(paste("../../results/", today, "/unitig_mapping_shape_breakdown.pdf", sep = ""), plot_unitigs_shape_breakdown, dpi = 400, height = 7, width = 11, unit = "in")

plot_unitigs_h_breakdown = ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  facet_grid(shape + h ~ .) +
  geom_vline(xintercept = 2e6/2, lty = 2, lwd = 1) +
  labs(x = "Center of longest unitig", color = "Nes", title = "Shape parameter breakdown")

ggarrange(plot_unitigs_all_sims, plot_unitigs_cov_breakdown, plot_unitigs_n_breakdown, nrow = 1, ncol = 3, labels = c("A", "B", "C"), common.legend = T, widths = c(1.5, 1, 1))

ggsave(paste("../results/", today, "/unitig_mapping.jpg", sep = ""), dpi = 400, height = 7, width = 9, unit = "in")

















check_overlap = function(x, nhits, target, tol){
  # sort by unitig length
  x_sorted = x[order(-x$alen),]
  
  # get top n longest unitigs
  x_top = x_sorted[1:nhits,]
  
  x_top$start = pmin(x_top$ss, x_top$se)
  x_top$end = pmax(x_top$ss, x_top$se)
  
  targ_start = target - tol
  targ_end = target + tol
  
  result = any( (targ_start < x_top$start) & (x_top$start < targ_end) ) | any( (targ_start < x_top$end) &  (x_top$end < targ_end) )
  
  return(result)
}

plotdata = data.frame(
  ID = unitig_files_ids,
  overlap = unlist(lapply(unitig_files, check_overlap, nhits = 10, target = 500e3, tol = 50e3))
)

plotdata = merge(params, plotdata, by = "ID")
plotdata$n = as.factor(plotdata$n)
plotdata$cov = as.factor(plotdata$cov)

aggregate(overlap ~ s, data = plotdata, FUN = mean)

ggplot(plotdata, aes(x = overlap, fill = simtype)) +
  geom_bar(stat = "count", position="dodge") +
  facet_grid(s ~ .) +
  theme_classic()










allele = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_277.txt", col.names = c("chrom", "pos", "ns", "af", "ac"))

mean(diff(allele$pos))

unitigs = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/unitig_alignments_277.txt", col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

ggplot(unitigs, aes(x = alen)) +
  geom_density()

mean(unitigs$alen)

allele = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_5558.txt", col.names = c("chrom", "pos", "ns", "af", "ac"))

mean(diff(allele$pos))

unitigs = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/unitig_alignments_5558.txt", col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

ggplot(unitigs, aes(x = alen)) +
  geom_density()

mean(unitigs$alen)

sam_pos = seq(from = 1000, to = 1e6, by = 1000)

results = NULL
i = 0
while(i <= 1e6){
  result = mean(unitigs[((unitigs$ss <= i) & (i + 1000 <= unitigs$se)),"alen"])
  results = rbind(results, data.frame(start = i, end = i + 1000, avglen = result))
  i = i + 1000
}

ggplot(results, aes(x = (start + end)/2, y = avglen)) +
  geom_point() +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = 5e5) +
  theme_classic()

results = c()
for(pos in sam_pos){
  result = mean(unitigs[((unitigs$ss <= pos) & (pos <= unitigs$se)),"alen"])
  results = c(results, result)
}

plotdata = data.frame(
  pos = sam_pos,
  avglen = results
)

ggplot(plotdata, aes(x = pos, y = avglen)) +
  geom_point() +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = 5e5) +
  theme_classic()

ggplot(unitigs, aes(x = log10(ss), y = alen)) +
  geom_point() +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = log10(5e5)) +
  theme_classic()

```

# process two population metrics (Fst)
```{r}
simids = params$ID[(params$simtype == "twopop")]

fst = function(n1, c1, f1, n2, c2, f2, type){

  if(type == "sample"){
      pi_wn = (1/2)*( (n1/(n1-1))*(c1/(c1-1))*(1-(f1^2+(1-f1)^2)) + (n2/(n2-1))*(c2/(c2-1))*(1-(f2^2+(1-f2)^2)) )
  }
  
  if(type == "population"){
      pi_wn = (1/2)*( (1-(f1^2+(1-f1)^2)) + (1-(f2^2+(1-f2)^2)) )
  }

  pi_bw = 1 - ( f1*f2 + (1-f1)*(1-f2) )
  
  pi_total = (pi_wn + pi_bw)/2
  
  nei = 1 - pi_wn/pi_total
  
  hudson = 1 - pi_wn/pi_bw
  
  return(data.frame(
    pi_wn = pi_wn,
    pi_bw = pi_bw,
    pi_total = pi_total,
    nei = nei,
    hudson = hudson
  ))
  
}

G = 2e6

k = 31

thresh = 5

#for(i in simids[1:5]){
for(i in 137:160){
  # load in actual allele freqs from simulations
  true_af_p1 = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_", i, "_p1.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"))
  
    true_af_p2 = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_", i, "_p2.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"))

  true_af_p1$ac = as.numeric(true_af_p1$ac)
  true_af_p2$ac = as.numeric(true_af_p2$ac)
  
  # calculate minor allele frequencies
  true_af_p1$af = as.numeric(true_af_p1$af)
  true_af_p2$af = as.numeric(true_af_p2$af)
  
  true_af_p1$snp_maf = pmin(true_af_p1$af, 1 - true_af_p1$af)
  true_af_p2$snp_maf = pmin(true_af_p2$af, 1 - true_af_p2$af)
  
  true_af_p1 = true_af_p1[(true_af_p1$af < 1) & (true_af_p1$af > 0),]
  true_af_p2 = true_af_p2[(true_af_p2$af < 1) & (true_af_p2$af > 0),]
  
  true_af_p1 = true_af_p1[complete.cases(true_af_p1),]
  true_af_p2 = true_af_p2[complete.cases(true_af_p2),]
  
  true_af = merge(true_af_p1[,c("pos", "ns", "ac", "af")], true_af_p2[,c("pos", "ns", "ac", "af")], by = "pos")
  
  fst_values = fst(2*true_af$ns.x, true_af$ac.x, true_af$af.x, 2*true_af$ns.y, true_af$ac.y, true_af$af.y, "population")
  
  #fst(true_af$ns.x, true_af$ac.x, true_af$af.x, true_af$ns.y, true_af$ac.y, true_af$af.y, "sample")
  
  true_af = cbind(true_af, fst_values)
  
  # load pairs of k-mers
  kmer_pairs_p1 = read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p1_seqs.csv", sep = ""), col.names = c("k1", "k2"))

  kmer_covs_p1 = read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p1_counts.csv", sep = ""), col.names = c("c1", "c2"))
  
  kmer_hashes_p1 = read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p1_hashes.csv", sep = ""), col.names = c("h"))

  kmer_pairs_p1 = cbind(kmer_pairs_p1, kmer_covs_p1, kmer_hashes_p1)

  kmer_pairs_p2 = read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p2_seqs.csv", sep = ""), col.names = c("k1", "k2"))

  kmer_covs_p2 = read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p2_counts.csv", sep = ""), col.names = c("c1", "c2"))

  kmer_hashes_p2 = read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p2_hashes.csv", sep = ""), col.names = c("h"))

  kmer_pairs_p2 = cbind(kmer_pairs_p2, kmer_covs_p2, kmer_hashes_p2)

  # calculate minor allele frequency
  kmer_pairs_p1$kmer_af = kmer_pairs_p1$c1/(kmer_pairs_p1$c1 + kmer_pairs_p1$c2)
  
  kmer_pairs_p2$kmer_af = kmer_pairs_p2$c1/(kmer_pairs_p2$c1 + kmer_pairs_p2$c2)
  
  kmer_pairs_p1p2 = merge(kmer_pairs_p1, kmer_pairs_p2, by = "h")
  
  # map kmers to snps
  pos_kmer_key = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/center_kmer_pairs_", i, "_p1p2.txt", sep = ""), fill = T, col.names = c("pos", "k1", "k2", "k3", "k4"))

  pos_kmer_key$leftk = substr(pos_kmer_key$k1, 1, 15)
  pos_kmer_key$rightk = substr(pos_kmer_key$k1, 17, 31)

  # compare snp and k-mer mafs
  kmer_pairs_p1p2$leftk = substr(kmer_pairs_p1p2$k1.x, 1, 15)
  kmer_pairs_p1p2$rightk = substr(kmer_pairs_p1p2$k1.x, 17, 31)

  pos_kmer_pairs = merge(pos_kmer_key, kmer_pairs_p1p2, by = c("leftk", "rightk"))

  # get snps that pair with reverse complements
  pos_kmer_key$leftk_rc = chartr("ATGC","TACG",pos_kmer_key$leftk)
  pos_kmer_key$rightk_rc = chartr("ATGC","TACG",pos_kmer_key$rightk)

  pos_kmer_key$leftk_rc = strReverse(pos_kmer_key$leftk_rc)
  pos_kmer_key$rightk_rc = strReverse(pos_kmer_key$rightk_rc)

  pos_kmer_pairs_rc = merge(pos_kmer_key, kmer_pairs_p1p2, by.x = c("leftk_rc", "rightk_rc"), by.y = c("rightk", "leftk"))

  pos_kmer_pairs_rc = subset(pos_kmer_pairs_rc, select=-c(leftk,rightk))

  # merge reverse complement and regular sets
  names(pos_kmer_pairs_rc)[1:2] = c("leftk", "rightk")

  pos_kmer_pairs = rbind(pos_kmer_pairs, pos_kmer_pairs_rc)

  # do some filtering
  # remove low coverage k-mer pairs
  # remove tri-allelic k-mers
  #pos_kmer_pairs = pos_kmer_pairs[(pos_kmer_pairs$c1 >=4 & pos_kmer_pairs$c2 >= 4),]
  pos_kmer_pairs = pos_kmer_pairs[( (pos_kmer_pairs$k3 == "") | is.na(pos_kmer_pairs$k3)),]
  
  # calculate bayesian allele counts
  sample_size = params[params$ID == i,"n"]
  
  coverage = params[params$ID == i,"cov"]
  
  sequencer = params[params$ID == i,"sequencer"]
  
  if(sequencer %in% c("miseq","nextseq")){
    L = 300
  }
  
  if(sequencer == "novaseq"){
    L = 150
  }
  
  if(sequencer == "hiseq"){
    L = 125
  }
  
  mean_kmer_coverage_per_genome = ((L-k+1)/L)*(coverage/sample_size)
  
  kmer_bayes_ac_p1 = mapply(posterior_min_kmer_count, pos_kmer_pairs$c1.x, (pos_kmer_pairs$c1.x + pos_kmer_pairs$c1.y), MoreArgs = list(n = sample_size, C = mean_kmer_coverage_per_genome, c = thresh, alpha = 1, beta = 1))
  
  kmer_bayes_ac_p2 = mapply(posterior_min_kmer_count, pos_kmer_pairs$c1.y, (pos_kmer_pairs$c1.y + pos_kmer_pairs$c1.y), MoreArgs = list(n = sample_size, C = mean_kmer_coverage_per_genome, c = thresh, alpha = 1, beta = 1))
  
  # calculate fst for kmers using empirical frequencies
  kmer_fst_values = fst(2*sample_size, pos_kmer_pairs$c1.x, pos_kmer_pairs$kmer_af.x, 2*sample_size, pos_kmer_pairs$c1.y, pos_kmer_pairs$kmer_af.y, "sample")

  pos_kmer_pairs = cbind(pos_kmer_pairs, kmer_fst_values, kmer_bayes_ac_p1, kmer_bayes_ac_p2)
  
  # add in SNP fst values
  snp_vs_kmer_fst = merge(true_af[,c("pos", "nei", "hudson")], pos_kmer_pairs[,c("pos", "nei", "hudson")], by = "pos")
  
  print(ggplot(snp_vs_kmer_fst, aes(x = nei.x, y = nei.y)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    theme_classic() +
    labs(x = "true SNP Nei Fst", y = "Nei hetmer Fst", title = i))
  
  print(ggplot(snp_vs_kmer_fst, aes(x = hudson.x, y = hudson.y)) +
    geom_point() +
    theme_classic() +
    geom_abline(slope = 1, intercept = 0) +
    labs(x = "true SNP Hudson Fst", y = "Hudson hetmer Fst", title = i))
  
  # mean(snp_vs_kmer_fst$nei.x)
  # mean(snp_vs_kmer_fst$nei.y)
  # 
  # mean(snp_vs_kmer_fst$hudson.x)
  # mean(snp_vs_kmer_fst$hudson.y)
  # 
  # mean(true_af$hudson)
  # mean(kmer_pairs_p1p2$hudson[(kmer_pairs_p1p2$hudson > 0)])
  
}

```

# Allele frequencies of known variants
```{r}

results = NULL

sims_to_plot = c(1, 4, 7, 10, 13, 25, 37, 49)

for(i in params$ID){
  print(i)
  
  kmer_counts = tryCatch({
    read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freqk_results/", i, "_0_freqs.txt", sep = ""), sep = "|")
  }, error = function(e) {
    message(paste("Error at iteration", i, ":", e$message))
    return(NA) # Or some other indicator for error
  })
  
  if (all(is.na(kmer_counts))){
    next
  }

  kmer_counts$kmer_maf = pmin(kmer_counts$V1, kmer_counts$V2)/(kmer_counts$V1 + kmer_counts$V2)
  
  kmer_index = tryCatch({
    read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freqk_ref_dedup/", i, "_0.txt", sep = ""), sep = ",")
  }, error = function(e) {
    message(paste("Error at iteration", i, ":", e$message))
    return(NA) # Or some other indicator for error
  })

  slim = tryCatch({
    read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/slim_results/allele_freqs_", i, "_0.txt", sep = ""), sep = " ")
    }, error = function(e) {
    message(paste("Error at iteration", i, ":", e$message))
    return(NA) # Or some other indicator for error
  })
  
  if (all(is.na(slim))){
    next
  }
  
  names(slim)[1:2] =  c("Chrom", "Position")
  
  slim$slim_maf = pmin(slim$V6, 1 - slim$V6)

  kmer_counts = cbind(kmer_index, kmer_counts)
  
  names(kmer_counts)[2:3] =  c("Chrom", "Position")
  
  # shift position to vcf convention
  kmer_counts$Position = kmer_counts$Position + 2
  
  slim_vs_kmer = merge(slim, kmer_counts, by = c("Chrom", "Position"))
  
  cor_coeff_freqk = signif(cor.test(slim_vs_kmer$kmer_maf, slim_vs_kmer$slim_maf)$estimate, digits = 3)
  
  #print(cor_coeff)
  
  #varscan = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/varscan_results/", i, "_0.tsv", sep = ""), header = T)
  
  #varscan$maf = pmin(varscan$Reads1, varscan$Reads2)/(varscan$Reads1 + varscan$Reads2)
  
  # freebayes = tryCatch({
  #   read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freebayes_a_priori_results/", i, "_0.txt", sep = ""), header = F)
  # }, error = function(e) {
  #   message(paste("Error at iteration", i, ":", e$message))
  #   return(NA) # Or some other indicator for error
  # })
  # 
  # if (all(is.na(freebayes))){
  #   next
  # }
  # 
  # names(freebayes)[c(1:2,6)] = c("Chrom", "Position", "af")
  # 
  # freebayes$maf = pmin(freebayes$af, 1 - freebayes$af)
  # 
  # slim_freebayes = merge(slim, freebayes, by = c("Chrom", "Position"))
  # 
  # cor_coeff_freebayes = signif(cor.test(slim_freebayes$maf.x, slim_freebayes$maf.y)$estimate, digits = 3)
  
  ## vg giraffe + freebayes
  vg = tryCatch({
    read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freebayes_vg_results/", i, "_0.txt", sep = ""), header = F)
  }, error = function(e) {
    message(paste("Error at iteration", i, ":", e$message))
    return(NA) # Or some other indicator for error
  })
  
  if (all(is.na(vg))){
    next
  }
  
  # load benchmarks
  
  names(vg)[c(1:2,6)] = c("Chrom", "Position", "af")
  
  vg$maf = pmin(vg$af, 1 - vg$af)
  
  slim_vg = merge(slim, vg, by = c("Chrom", "Position"))
  
  cor_coeff_vg = signif(cor.test(slim_vg$maf.x, slim_vg$maf.y)$estimate, digits = 3)
  
  ## bwa mem + freebayes
  #bench_bwa_mem = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/bwa_mem/", i, "_0.bench", sep = ""), header = T)
  
  #bench_freebayes_a_priori = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freebayes_a_priori/", i, "_0.bench", sep = ""), header = T)
  
  ## vg giraffe + freebayes
  bench_vg_index = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_autoindex/", i, "_0.bench", sep = ""), header = T)
  
  bench_vg_giraffe = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_giraffe/", i, "_0.bench", sep = ""), header = T)
  
  bench_vg_surject_paired = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_surject/paired_", i, "_0.bam", sep = ""), header = T)
  
  bench_vg_surject_r1 = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_surject/R1_", i, "_0.bam", sep = ""), header = T)
  
  bench_vg_surject_r2 = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_surject/R2_", i, "_0.bam", sep = ""), header = T)
    
  bench_vg_sort_paired = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/samtools_sort/paired_", i, "_0.bam", sep = ""), header = T)
  
  bench_vg_sort_r1 = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/samtools_sort/R1_", i, "_0.bam", sep = ""), header = T)
  
  bench_vg_sort_r2 = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/samtools_sort/R2_", i, "_0.bam", sep = ""), header = T)
  
  bench_vg_merge = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/samtools_merge/", i, "_0.bam", sep = ""), header = T)
  
  bench_freebayes_vg = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freebayes_vg/", i, "_0.bench", sep = ""), header = T)
  
  ## kmer counts
  bench_freqk_index = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freqk_index/", i, "_0.bench", sep = ""), header = T)
  
  bench_freqk_dedup = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freqk_dedup/", i, "_0.bench", sep = ""), header = T)
  
  bench_freqk_count = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freqk_count/", i, "_0.bench", sep = ""), header = T)
  
  # combine all stats in a frame
  result = data.frame(
    ID = i,
    cor_coeff_freqk = cor_coeff_freqk,
    #cor_coeff_freebayes = cor_coeff_freebayes,
    cor_coeff_vg = cor_coeff_vg,
    t_freqk = bench_freqk_count$s,
    mem_freqk = max(bench_freqk_count$max_rss),
    t_freqk_index = bench_freqk_index$s + bench_freqk_dedup$s,
    mem_freqk_index = max(bench_freqk_index$max_rss + bench_freqk_dedup$max_rss),
    #t_freebayes = bench_bwa_mem$s + bench_freebayes_a_priori$s,
    #mem_freebayes = max(bench_bwa_mem$max_rss, bench_freebayes_a_priori$max_rss),
    t_vg = bench_vg_giraffe$s + max(c(bench_vg_surject_paired$s, bench_vg_surject_r1$s, bench_vg_surject_r2$s)) + max(c(bench_vg_sort_paired$s, bench_vg_sort_r1$s, bench_vg_sort_r2$s)) + bench_vg_merge$s + bench_freebayes_vg$s,
    mem_vg = max(bench_vg_giraffe$max_rss, bench_vg_surject_paired$max_rss, bench_vg_surject_r1$max_rss, bench_vg_surject_r2$max_rss, bench_vg_sort_paired$max_rss, bench_vg_sort_r1$max_rss, bench_vg_sort_r2$max_rss, bench_vg_merge$max_rss, bench_freebayes_vg$max_rss),
    t_vg_index = bench_vg_index$s,
    mem_vg_index = bench_vg_index$max_rss
  )
  
  results = rbind(results, result)
  
  # make a plot of some individual simulations
  if(i %in% sims_to_plot){
    plotdata_known_var = data.frame(kmer_maf = kmer_counts$maf, slim_maf = slim$maf)

ggplot(plotdata_known_var, aes(x = slim_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, lwd = 2) +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  labs(x = "True MAF", y = "k-mer counts MAF estiamte", title = paste("r = ", cor_coeff_freqk, sep = ""))

ggsave(paste("../../results/", today, "/", i, ".pdf", sep = ""), width = 7, height = 7)
  }
  
}


# compare mapping approaches to freqk
params_res = merge(params, results, by = "ID")

# ggplot(results, aes(x = cor_coeff_freebayes, y = cor_coeff_freqk)) +
#   geom_point(shape = 16) +
#   geom_abline(slope = 1, intercept = 0, lwd = 2) +
#   theme_minimal() +
#   theme(text = element_text(size = 16, family = "Helvetica")) +
#   labs(x = "Pearson Corr. of MAF\nbwa+Freebayes vs SLiM", y = "Pearson Corr. of MAF\nK-mers vs SLiM")

ggplot(params_res, aes(x = cor_coeff_vg, y = cor_coeff_freqk, color = as.factor(cov))) +
  geom_point(shape = 16) +
  geom_abline(slope = 1, intercept = 0, lwd = 2) +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_colour_scico_d(palette = "hawaii", name = "Coverage") +
  labs(x = "Pearson Corr.\nvg+Freebayes vs SLiM MAF", y = "Pearson Corr.\nk-mers vs SLiM MAF")

ggsave(paste("../../results/", today, "/", "vg_vs_freqk.pdf", sep = ""), height = 6, width = 7, device = cairo_pdf)

# compare run times per sample
time_res = melt(params_res[,c("ID", "cov", "n", "t_freqk", "t_vg")], id = c("ID", "cov", "n"))

mem_res = melt(params_res[,c("ID", "cov", "n", "mem_freqk", "mem_vg")], id = c("ID", "cov", "n"))

names(mem_res)[4] = "tool"
names(mem_res)[5] = "mem"
mem_res$tool = gsub("mem_","", mem_res$tool)

names(time_res)[4] = "tool"
names(time_res)[5] = "time"
time_res$tool = gsub("t_","", time_res$tool)

res_summary = merge(time_res, mem_res, by =  c("ID", "cov", "n", "tool"))

# compare time and memory
plot_time = ggplot(res_summary, aes(x = tool, y = time, fill = as.factor(cov))) +
  geom_boxplot() +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_fill_scico_d(palette = "hawaii", name = "Coverage") +
  #facet_zoom(ylim = c(0,750)) +
  #stat_compare_means(method = "wilcox", paired = TRUE) +
  #ylim(0,1000) +
  labs(x = "Workflow", y = "Wall Time (s)", fill = "Coverage")

plot_mem = ggplot(res_summary, aes(x = tool, y = mem, fill = as.factor(cov))) +
  geom_boxplot() +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_fill_scico_d(palette = "hawaii", name = "Coverage") +
  labs(x = "Workflow", y = "Maximum\nResident Set Size (MB)", fill = "Coverage")

ggarrange(plot_time, plot_mem, nrow = 1, ncol = 2, labels = c("A", "B"), font.label = list(size = 16, family = "Helvetica"), common.legend = T)

ggsave(paste("../../results/", today, "/", "resources_freqk_vs_vg.pdf", sep = ""), height = 4, width = 9)

# compare index times
time_index = melt(params_res[,c("ID", "cov", "n", "t_freqk_index", "t_vg_index")], id = c("ID", "cov", "n"))

mem_index = melt(params_res[,c("ID", "cov", "n", "mem_freqk_index", "mem_vg_index")], id = c("ID", "cov", "n"))

names(mem_index)[4] = "tool"
names(mem_index)[5] = "mem"
mem_index$tool = gsub("mem_","", mem_index$tool)
mem_index$tool = gsub("_index","", mem_index$tool)

names(time_index)[4] = "tool"
names(time_index)[5] = "time"
time_index$tool = gsub("t_","", time_index$tool)
time_index$tool = gsub("_index","", time_index$tool)

res_index = merge(time_index, mem_index, by =  c("ID", "cov", "n", "tool"))

plot_time_index = ggplot(res_index, aes(x = tool, y = time, fill = as.factor(n))) +
  geom_boxplot() +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_fill_scico_d(palette = "acton", name = "Pool size") +
  labs(x = "Workflow", y = "Wall Time (s)", fill = "Pool size")

plot_mem_index = ggplot(res_index, aes(x = tool, y = mem, fill = as.factor(n))) +
  geom_boxplot() +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_fill_scico_d(palette = "acton", name = "Coverage") +
  labs(x = "Workflow", y = "Maximum\nResident Set Size (MB)", fill = "Pool size")

ggarrange(plot_time_index, plot_mem_index, nrow = 1, ncol = 2, labels = c("A", "B"), font.label = list(size = 16, family = "Helvetica"), common.legend = T)

ggsave(paste("../../results/", today, "/", "resources_index_freqk_vs_vg.pdf", sep = ""), height = 4, width = 7)
```

# Empirical data: frequencies of known variants

```{r}
real_counts = read.table("/global/scratch/users/milesroberts/tmp/arabidopsis_thaliana_plantpan_data/counts_by_allele.txt", sep = "|", col.names = c("c1", "c2", "c3", "c4"), fill = T, blank.lines.skip = F)

# alternate allele frequency
real_counts$af = 1 - real_counts$c1/(real_counts$c1 + real_counts$c2)

real_vcf <- read.vcfR("~/tmp/at.genetics.variation/final.vcf.gz")

gt <- apply(real_vcf@gt[,-1], 2, as.numeric)

to_keep = is.na(real_counts$c3)

gt = gt[-nrow(gt), ]

gt = gt[to_keep, ]

real_counts = real_counts[to_keep,]

allele_counts <- rowSums(gt == 1)

real_plot_data = data.frame(
  estim_af = real_counts$af*7,
  estim_rounded_af = round(real_counts$af*7),
  true_af = allele_counts
)


cor_estim <- signif(cor.test(real_counts$af, allele_counts)$estimate, 3)

ggplot(real_plot_data, aes(x = as.factor(true_af), y = estim_af)) +
  geom_boxplot() +
  geom_point(aes(x=1, y=2), colour="red", size = 3, shape = 16) +
  geom_point(aes(x=2, y=3), colour="red", size = 3, shape = 16) +
  geom_point(aes(x=3, y=4), colour="red", size = 3, shape = 16) +
  geom_point(aes(x=4, y=5), colour="red", size = 3, shape = 16) +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  labs(x = "True allele frequency", y = "Estimated allele frequency", title = paste("r = ", cor_estim, sep = ""))

estim_by_counts = table(real_plot_data$true_af, real_plot_data$estim_rounded_af)

estim_by_counts = melt(estim_by_counts)

estim_by_counts$bar_color = ifelse(estim_by_counts$Var.1 == estim_by_counts$Var.2, "red", "grey")

# plot number of estimates for each af class
ggplot(real_plot_data, aes(x = estim_rounded_af)) +
  geom_bar() +
  facet_wrap( ~ true_af) +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  labs(x = "Estimated allele frequency")

ggsave(paste("../../results/", today, "/", "true_vs_estimated_7_genomes.pdf", sep = ""), width = 7, height = 7)
```

# benchmarks
```{r}
bench_file_names = list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks", recursive = T, full.names = T)

#bench_file_names = bench_file_names[!grepl("masked", bench_file_names)] # remove benchmarks from masked experiments for now

bench_files = lapply(bench_file_names, read.table, header = T, sep = "\t")

bench_files = do.call("rbind", bench_files)

#
bench_files$files = bench_file_names

bench_files$tool = gsub("/.*", "", gsub("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/","",bench_files$files))

bench_files$ID = as.numeric(gsub(".*/|.bench|_0.bench", "", gsub("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/","",bench_files$files)))

# merge with workflow parameters
bench_params = merge(params[(params$simtype == "onepop"),c("ID", "cov", "n")], bench_files, by = "ID")

# summarize time and memory by simulation parameters
bench_summary_time = aggregate(s ~ tool + cov, data = bench_params, FUN = "sum")

bench_summary_cpu_time = aggregate(cpu_time ~ tool + cov, data = bench_params, FUN = "sum")

bench_summary_max_rss = aggregate(max_rss ~ tool + cov, data = bench_params, FUN = "sum")

bench_summary_max_vms = aggregate(max_vms ~ tool + cov, data = bench_params, FUN = "sum")

bench_summary_max_uss = aggregate(max_uss ~ tool + cov, data = bench_params, FUN = "sum")

# try to summarize time and memory into snp calling workflow and kmer calling workflow
workflow_summary = function(x, lev_col, resource_col){
  levs = unique(x[,lev_col])
  
  results = NULL
  for(lev in levs){
    
    bwa_mem = x[(x$tool == "bwa_mem") & (x[,lev_col] == lev),resource_col]
    
    varscan = x[(x$tool == "varscan") & (x[,lev_col] == lev),resource_col]
 
    poolsnp = x[(x$tool == "poolsnp") & (x[,lev_col] == lev),resource_col]

    kmc = x[(x$tool == "kmc") & (x[,lev_col] == lev),resource_col]
    
    smudgeplot = x[(x$tool == "smudgeplot") & (x[,lev_col] == lev),resource_col]
    
    hetmers = x[(x$tool == "hetmers") & (x[,lev_col] == lev),resource_col]

    discosnp = x[(x$tool == "discosnp") & (x[,lev_col] == lev),resource_col]
    
    unitig_caller = x[(x$tool == "unitig_caller") & (x[,lev_col] == lev),resource_col]
    
    result = data.frame(
      lev_col = lev,
      bwa_varscan = bwa_mem+varscan,
      bwa_poolsnp = bwa_mem+poolsnp,
      kmc_smudgeplot = kmc+smudgeplot,
      kmc_hetmers=kmc+hetmers,
      kmc_unitig_caller = kmc+unitig_caller,
      discosnp = discosnp
    )
    
    results = rbind(results, result)
    
  }
  
  names(results) = c("lev_col", "bwa+varscan", "bwa+poolsnp", "kmc+smudgeplot", "kmc+hetmers", "kmc+unitig-caller", "discosnp")
  
  return(results)
}

#bench_summary_time_by_flow = workflow_summary(bench_summary_time, "cov", "s")
#bench_summary_time_by_flow = melt(bench_summary_time_by_flow, id = "lev_col")

bench_summary_cpu_time_by_flow = workflow_summary(bench_summary_cpu_time, "cov", "cpu_time")
bench_summary_cpu_time_by_flow = melt(bench_summary_cpu_time_by_flow, id = "lev_col")

#bench_summary_max_rss_by_flow = workflow_summary(bench_summary_max_rss, "cov", "max_rss")
#bench_summary_max_rss_by_flow = melt(bench_summary_max_rss_by_flow, id = "lev_col")
# 
# bench_summary_max_vms_by_flow = workflow_summary(bench_summary_max_vms, "cov", "max_vms")
# bench_summary_max_vms_by_flow = melt(bench_summary_max_vms_by_flow, id = "lev_col")
# 
# bench_summary_max_uss_by_flow = workflow_summary(bench_summary_max_uss, "cov", "max_uss")
# bench_summary_max_uss_by_flow = melt(bench_summary_max_uss_by_flow, id = "lev_col")

# calculate average times
# bench_summary_time_by_flow$value = bench_summary_time_by_flow$value/length(unique(bench_params$ID))

bench_summary_cpu_time_by_flow$value = bench_summary_cpu_time_by_flow$value/length(unique(bench_params$ID))

bench_summary_max_rss$max_rss = bench_summary_max_rss$max_rss/length(unique(bench_params$ID))

# bench_summary_max_vms_by_flow$value = bench_summary_max_vms_by_flow$value/length(unique(bench_params$ID))
# 
# bench_summary_max_uss_by_flow$value = bench_summary_max_uss_by_flow$value/length(unique(bench_params$ID))

# time plots
my_colors = scico(4, palette = "hawaii")
names(my_colors) = unique(bench_summary_cpu_time_by_flow$lev_col)

# ggplot(bench_summary_time_by_flow, aes(x = variable, y = value, fill = as.factor(lev_col))) +
#   geom_bar(stat = "identity", position = "dodge") +
#   theme_classic() +
#   scale_fill_manual(values = my_colors) +
#   labs(x = "Workflow", y = "Average time (s)", fill = "Coverage")

plot_time = ggplot(bench_summary_cpu_time_by_flow, aes(x = variable, y = value, fill = as.factor(lev_col))) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_classic() +
  theme(
    axis.line = element_line(colour = 'black', linewidth = 1),
    axis.ticks = element_line(colour = "black", linewidth = 1),
    axis.text.x = element_text(angle = 45, hjust=1),
    text = element_text(family = "Sans")) +
  scale_fill_manual(values = my_colors) +
  labs(x = "Workflow", y = "Average CPU Time (s)", fill = "Coverage")

# memory plots
bench_summary_max_rss$tool[(bench_summary_max_rss$tool == "unitig_caller")] = "unitig-caller"

bench_summary_max_rss$tool = factor(bench_summary_max_rss$tool, levels = c("bwa_index", "bwa_mem", "varscan", "poolsnp", "freebayes", "kmc", "smudgeplot", "hetmers", "unitig-caller", "discosnp"))

plot_memory = ggplot(bench_summary_max_rss, aes(x = tool, y = max_rss*1.04858, fill = as.factor(cov))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Tool", y = "Average Memory (MB)", fill = "Coverage") +
  theme_classic() +
  theme(
    axis.line = element_line(colour = 'black', linewidth = 1),
    axis.ticks = element_line(colour = "black", linewidth = 1),
    axis.text.x = element_text(angle = 45, hjust=1),
    text = element_text(family = "Sans")) +
  scale_fill_manual(values = my_colors) 

# ggplot(bench_summary_max_vms_by_flow, aes(x = variable, y = value, fill = as.factor(lev_col))) +
#   geom_bar(stat = "identity", position = "dodge") +
#   labs(x = "workflow", y = "Average Maximum Virtual Memory Size (MiB)", fill = "coverage") +
#   theme_classic() +
#   scale_fill_manual(values = my_colors) 
# 
# ggplot(bench_summary_max_uss_by_flow, aes(x = variable, y = value, fill = as.factor(lev_col))) +
#   geom_bar(stat = "identity", position = "dodge") +
#   labs(x = "workflow", y = "Average Maximum Unique Set Size (MiB)", fill = "coverage") +
#   theme_classic() +
#   scale_fill_manual(values = my_colors) 
  
ggarrange(plot_time, plot_memory, common.legend = T, font.label = list(size = 16), labels = c("A", "B"))
ggsave(paste("../results/", today, "/memory_and_time.jpg", sep = ""), dpi = 450)
```

# How does performance vary with sequencing parameters?
```{r}
# compare sensitivities for SNPs represented in reference
tpr_plot_data = melt(pares[,c("ID", "ID2", "cov", "n", "sequencer", "simtype", "kmer_tpr", "hetmer_tpr", "disco_tpr", "varscan_tpr", "poolsnp_tpr")], id = c("ID", "ID2", "cov", "n", "sequencer", "simtype"))

tpr_plot_data$variable = factor(gsub("_tpr", "", tpr_plot_data$variable), levels = c("kmer", "hetmer", "disco", "varscan", "poolsnp"), ordered = T)

tpr_plot_data$cov_per_n = as.factor(tpr_plot_data$cov/(tpr_plot_data$n * 2))

tpr_plot_data$cov = as.factor(tpr_plot_data$cov)
tpr_plot_data$n = as.factor(tpr_plot_data$n)

tpr_vs_coverage_plot = ggplot(tpr_plot_data, aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Total coverage", y = "Sensitivity", color = "Method", title = "All instruments")

#ggsave(paste("../results/", today, "/coverage_vs_sensitivity.png", sep = ""), unit = "in")

# ggplot(tpr_plot_data, aes(x = cov_per_n, y = value, color = simtype)) +
#   geom_boxplot() +
#   theme_classic() +
#   theme(text = element_text(size = 12)) +
#   labs(x = "Coverage per haploid genome", y = "Sensitivity (true positive rate)", color = "Method", title = "All instruments")

# coverage vs false discovery rate
fpr_plot_data = melt(pares[,c("ID", "cov", "n", "sequencer", "kmer_fpr", "hetmer_fpr", "disco_fpr", "varscan_fpr", "poolsnp_fpr")], id = c("ID", "cov", "n", "sequencer"))

fpr_plot_data$variable = factor(gsub("_fpr", "", fpr_plot_data$variable), levels = c("kmer", "hetmer", "disco", "varscan", "poolsnp"), ordered = T)

fpr_plot_data$cov = as.factor(fpr_plot_data$cov)
fpr_plot_data$n = as.factor(fpr_plot_data$n)

fpr_vs_coverage_plot = ggplot(fpr_plot_data, aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Total coverage", y = "False positive rate", color = "Method", title = "All instruments")

#ggsave(paste("../results/", today, "/coverage_vs_fpr.png", sep = ""), unit = "in")

# coverage vs false negative rate
fnr_plot_data = melt(pares[,c("ID", "cov", "n", "sequencer", "kmer_fnr", "hetmer_fnr", "disco_fnr", "varscan_fnr", "poolsnp_fnr")], id = c("ID", "cov", "n", "sequencer"))

fnr_plot_data$variable = factor(gsub("_fnr", "", fnr_plot_data$variable), levels = c("kmer", "hetmer", "disco", "varscan", "poolsnp"), ordered = T)

fnr_plot_data$cov = as.factor(fnr_plot_data$cov)
fnr_plot_data$n = as.factor(fnr_plot_data$n)

fnr_vs_coverage_plot = ggplot(fnr_plot_data, aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Total coverage", y = "False negative rate", color = "Method", title = "All instruments")

# coverage vs specificity
tnr_plot_data = melt(pares[,c("ID", "ID2", "cov", "n", "sequencer", "simtype", "kmer_tnr", "hetmer_tnr", "disco_tnr", "varscan_tnr", "poolsnp_tnr")], id = c("ID", "ID2", "cov", "n", "sequencer", "simtype"))

tnr_plot_data$variable = factor(gsub("_tnr", "", tnr_plot_data$variable), levels = c("kmer", "hetmer", "disco", "varscan", "poolsnp"), ordered = T)

tnr_plot_data$cov = as.factor(tnr_plot_data$cov)
tnr_plot_data$n = as.factor(tnr_plot_data$n)

tnr_vs_coverage_plot = ggplot(tnr_plot_data, aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Total coverage", y = "Specificity", color = "Method", title = "All instruments")

# plot tpr and fnr together
ggarrange(tpr_vs_coverage_plot, tnr_vs_coverage_plot, fpr_vs_coverage_plot, fnr_vs_coverage_plot, common.legend = T, font.label = list(size = 16), labels = c("A", "B", "C", "D"), nrow = 1, ncol = 4)

ggsave(paste("../results/", today, "/coverage_vs_tpr_tnr_fpr_fnr.png", sep = ""), unit = "in", width = 12, height = 4, dpi = 350)











# coverage vs false positives
fp_plot_data = melt(pares[,c("ID", "cov", "n", "sequencer", "kmer_fp", "disco_fp", "varscan_fp", "poolsnp_fp")], id = c("ID", "cov", "n", "sequencer"))

fp_plot_data$variable = factor(gsub("_fp", "", fp_plot_data$variable), levels = c("kmer", "disco", "varscan", "poolsnp"), ordered = T)

fp_plot_data$cov_per_n = as.factor(fp_plot_data$cov/(fp_plot_data$n * 2))

ggplot(fp_plot_data, aes(x = cov_per_n, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Coverage per haploid genome", y = "False positives", color = "Method", title = "All instruments")

#ggsave(paste("../results/", today, "/coverage_vs_fp.png", sep = ""), unit = "in")

#




#
tpr_vs_fpr = merge(tpr_plot_data, fpr_plot_data, by = c("ID", "cov", "n", "sequencer", "variable", "cov_per_n"))

ggplot(tpr_vs_fpr, aes(x = value.y, y = value.x, color = variable)) +
  geom_point() +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "False positive rate", y = "True positive rate", color = "Method", title = "All instruments")

ggplot(tpr_plot_data[(tpr_plot_data$n == 50 & tpr_plot_data$sequencer == "miseq"),], aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "Coverage", y = "Sensitivity (true positive rate)", color = "Method", title = "pool size = 50; MiSeq")

ggsave(paste("../results/", today, "/sensitivity_50_miseq.png", sep = ""), height = 7, width = 8)

# hold pool size constant and increase coverage
ggplot(pares[(pares$n == 5),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 10),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 25),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 50),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 100),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

# hold coverage constant, and add more individuals to pool
ggplot(pares[(pares$cov == 20),], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$cov == 50),], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$cov == 80),], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$cov == 100),], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

# hold pool size constant, increase coverage
ggplot(pares[(pares$n == 5),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 10),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 25),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 50),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 100),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()





ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "True MAF", y = "Predicted MAF from k-mers")

ggplot(compare_maf_kmer, aes(x = snp_maf, y = snp_maf_varscan)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "True MAF", y = "Predicted MAF from VarScan")

# what about for k-mers above some frequency threshold
kmer_test = cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$kmer_maf, method = "spearman")

ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "True MAF", y = "Predicted MAF from k-mers", title = paste("rho =", signif(kmer_test$estimate, 4), "; p =", paste(format(signif(kmer_test$p.value, 2), digits = 2))))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","kmer_maf_vs_truth_pool-seq.png", sep = ""), height= 7, width = 7)

varscan_test = cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$snp_maf_varscan, method = "spearman")

ggplot(compare_maf_kmer, aes(x = snp_maf, y = snp_maf_varscan)) +
  geom_density_2d_filled() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "True MAF", y = "Predicted MAF from VarScan", title = paste("rho =", signif(varscan_test$estimate, 4), "; p =", paste(format(signif(varscan_test$p.value, 2), digits = 2))))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","varscan_maf_vs_truth_pool-seq.png", sep = ""), height= 7, width = 8)

# smudgeplot
ggplot(compare_maf_kmer, aes(x = kmer_maf, y = (c1+c2)/81.670932)) +
  geom_density_2d_filled() +
  #geom_hline(yintercept = 50) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "Minor k-mer frequency", y = "Total pair coverage")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","smudgeplot.png", sep = ""), height= 7, width = 8)



# folded sfs
kmer_pairs = kmer_pairs[(kmer_pairs$c1 >=4 & kmer_pairs$c2 >= 4),]

folded_sfs = data.frame(
    category=factor(c(rep("truth", times = nrow(true_af)), rep("kmer", times = nrow(kmer_pairs)), rep("varscan", times = nrow(varscan)))),
    value=c(true_af$snp_maf,kmer_pairs$kmer_maf, varscan$snp_maf_varscan)
)

ggplot(folded_sfs, aes(x = value, fill = category)) +
  geom_density(alpha = 0.5) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "minor allele frequency")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","kmer_vs_snp_sfs.png", sep = ""), height= 7, width = 8)

ggplot(true_af, aes(x = snp_maf)) +
  geom_density() +
  theme_classic()

ggplot(kmer_pairs, aes(x = kmer_maf)) +
  geom_density() +
  theme_classic()

ggplot(kmer_pairs, aes(x = kmer_maf)) +
  geom_density() +
  theme_classic()

ggplot(true_af[(true_af$snp_maf >= 0.05),], aes(x = snp_maf)) +
  geom_density() +
  theme_classic()

ggplot(compare_maf_kmer, aes(x = kmer_maf)) +
  geom_histogram(binwidth = 0.05) +
  theme_classic()

ggplot(compare_maf_kmer, aes(x = snp_maf)) +
  geom_histogram(breaks = seq(from = 0, to = 0.5, by = 0.02)) +
  theme_classic()

ggplot(compare_maf_kmer, aes(x = kmer_maf)) +
  geom_histogram(breaks = seq(from = 0, to = 0.5, by = 0.02)) +
  theme_classic()

# calculate error
compare_maf_kmer$error = abs(compare_maf_kmer$kmer_maf - compare_maf_kmer$snp_maf)

summary(compare_maf_kmer$error)

sum(compare_maf_kmer$error < 0.1)/nrow(compare_maf_kmer)

ggplot(confident_sites, aes(x = total_cov, y = error)) +
  geom_point() +
  theme_classic()

#
compare_maf_kmer$snp_maf_bins = cut(compare_maf_kmer$snp_maf, breaks = seq(from = 0, to = 0.5, by = 0.05))
compare_maf_kmer$kmer_maf_bins = cut(compare_maf_kmer$kmer_maf, breaks = seq(from = 0, to = 0.5, by = 0.05))

observed <- table(compare_maf_kmer$kmer_maf_bins)  # Observed frequencies
expected <- table(compare_maf_kmer$snp_maf_bins)  # Expected frequencies

observed = observed[(expected != 0)]
expected = expected[(expected != 0)]

# Calculate Chi-Square statistic manually
chi_sq_statistic <- sum((observed - expected)^2 / expected)
df <- length(observed) - 1
p_value <- 1 - pchisq(chi_sq_statistic, df)

# Print results
print(paste("Chi-Square Statistic:", chi_sq_statistic))
print(paste("Degrees of Freedom:", df))
print(paste("P-value:", p_value))

```

# correlations between true and predicted minor allele frequencies
```{r}
# load pairs of k-mers
kmer_pairs = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", "kmerpairs_938_sequences.tsv", sep = ""), col.names = c("k1"))

kmer_covs = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/kmerpairs_938_coverages.tsv", sep = ""), col.names = c("c1", "c2"))

kmer_pairs = cbind(kmer_pairs, kmer_covs)

# calculate minor allele frequency
kmer_pairs$kmer_maf = pmin(kmer_pairs$c1, kmer_pairs$c2)/(kmer_pairs$c1 + kmer_pairs$c2)

# load in actual allele freqs from simulations
true_af = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_938.txt", sep = ""), col.names = c("chrom", "pos", "af", "ac"))

true_af$ac = as.numeric(true_af$ac)

true_af = true_af[complete.cases(true_af),]

# calculate minor allele frequencies
true_af$af = as.numeric(true_af$af)
true_af$snp_maf = pmin(true_af$af, 1 - true_af$af)

# get key for position to k-mers
pos_kmer_key = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/center_kmer_pairs_938.txt", sep = ""), fill = T, col.names = c("pos", "k1", "k2", "k3", "k4"))

pos_kmer_key$leftk = substr(pos_kmer_key$k1, 1, 15)
pos_kmer_key$rightk = substr(pos_kmer_key$k1, 17, 31)

# compare snp and k-mer mafs
kmer_pairs$leftk = substr(kmer_pairs$k1, 1, 15)
kmer_pairs$rightk = substr(kmer_pairs$k1, 17, 31)

pos_kmer_pairs = merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

# get snps that pair with reverse complements
pos_kmer_key$leftk_rc = chartr("ATGC","TACG",pos_kmer_key$leftk)
pos_kmer_key$rightk_rc = chartr("ATGC","TACG",pos_kmer_key$rightk)

pos_kmer_key$leftk_rc = strReverse(pos_kmer_key$leftk_rc)
pos_kmer_key$rightk_rc = strReverse(pos_kmer_key$rightk_rc)

pos_kmer_key$rightk = pos_kmer_key$leftk_rc 
pos_kmer_key$leftk = pos_kmer_key$rightk_rc

pos_kmer_pairs_rc = merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

# merge reverse complement and regular sets
any(duplicated(c(pos_kmer_pairs$pos, pos_kmer_pairs_rc$pos)))

pos_kmer_pairs = rbind(pos_kmer_pairs, pos_kmer_pairs_rc[,c(-8,-9)])

# do some filtering
# remove low coverage k-mer pairs
# remove tri-allelic k-mers
pos_kmer_pairs = pos_kmer_pairs[(pos_kmer_pairs$c1 >=4 & pos_kmer_pairs$c2 >= 4),]
pos_kmer_pairs = pos_kmer_pairs[( (pos_kmer_pairs$k3 == "") | is.na(pos_kmer_pairs$k3)),]

compare_maf_kmer = merge(true_af[,c("pos", "snp_maf")], pos_kmer_pairs, by = "pos")

# include varscan calls
varscan = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/calls_938.tsv", sep = ""), header = T)

varscan$snp_maf_varscan = pmin(varscan$Reads1, varscan$Reads2)/(varscan$Reads1 + varscan$Reads2)

names(varscan)[2] = "pos"

# compare varscan and ground truth
compare_maf_varscan = merge(true_af[,c("pos", "snp_maf")], varscan, by = "pos")

# compare k-mer and varscan
compare_maf_kmer_varscan = merge(compare_maf_kmer, varscan, by = "pos")

# round k-mer mafs to nearest increment 
kmer_test = cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$kmer_maf, method = "pearson")

varscan_test = cor.test(compare_maf_varscan$snp_maf, compare_maf_varscan$snp_maf_varscan, method = "pearson")

# create plots
p1 = ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, lwd = 2, color = "purple") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(x = "True MAF", y = "MAF from k-mers", title = paste("rho =", signif(kmer_test$estimate, 3), "  p =", paste(format(signif(kmer_test$p.value, 3), digits = 3))))

p2 = ggplot(compare_maf_varscan, aes(x = snp_maf, y = snp_maf_varscan)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, lwd = 2, color = "purple") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(x = "True MAF", y = "MAF from varscan", title = paste("rho =", signif(varscan_test$estimate, 3), "  p =", paste(format(signif(varscan_test$p.value, 3), digits = 3))))

ggarrange(p1, p2, nrow = 2, ncol = 1, labels = c("A", "B"), font.label = list(size = 12))

ggsave(paste("../results/", today, "/true_vs_predicted_maf.png", sep = ""), height = 6, width = 3, unit = "in")
```

```{r}
kmer_pairs = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/testpairs_sequences.tsv", col.names = c("k1"))

kmer_counts = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/testpairs_coverages.tsv", col.names = c("c1", "c2"))

kmer_pairs = cbind(kmer_pairs, kmer_counts)

# calculate minor allele frequency
nidv = 5
ploidy = 2

kmer_pairs$kmer_maf = round(pmin(kmer_pairs$c1, kmer_pairs$c2)/(kmer_pairs$c1 + kmer_pairs$c2)*nidv*ploidy)

# filter out k-mers below coverage thresholds
kmer_pairs = kmer_pairs[( (kmer_pairs$c1 > 3) & (kmer_pairs$c2 > 3) ),]



kmer_counts = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/kmer_counts_161.txt", col.names = c("k1", "c1"))

# frequency vs coverage
ggplot(kmer_counts[(kmer_counts$c1 >=4 ),], aes(x = c1)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 4) +
  theme_classic()

# frequency*coverage vs coverage
freq = table(kmer_counts$c1)

cov = as.numeric(names(freq))
freq = as.vector(freq)

trans_spectrum = data.frame(
  freq = freq,
  freq_cov = freq*cov,
  cov = cov
)

ggplot(trans_spectrum, aes(x = cov, y = freq_cov)) +
  geom_bar(stat = "identity") +
  theme_classic()

#
testmers = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/testmers.txt", col.names = c("k1", "c1"))

table(testmers$c1)/nrow(testmers) * 100

# try modeling coverage of a k-mer based on it's frequency in a pool
lamb = 60 # average haploid sequencing coverage
rho = 0.8 # dispersion due to sequencing biases
i = 3 # number of genomes where k-mer is present
n = 60 # number of genomes in pool

plotdata = data.frame(
  x = rnbinom(1000, mu = (i/n)*lamb, size = (i/n)*(lamb/rho))
)

ggplot(plotdata, aes(x)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = (i/n)*lamb)

# try fitting poisson to distribution
fitdistr(kmer_counts$c1, "poisson", method = "L-BFGS-B", lower = 0.1, upper = 0.5)

fitdistr(kmer_counts$c1[], "negative binomial")

ggplot(trans_spectrum, aes(x = cov, y = freq/sum(freq))) +
  geom_bar(stat = "identity") +
  geom_point(aes(y=dpois(cov, 6.7055415264)), colour="red") +
  theme_classic()

ggplot(kmer_counts, aes(x = c1)) +
  geom_histogram(binwidth = 1) +
  geom_point(aes(y=dpois(c1, 1.0344408452)), colour="red") +
  geom_vline(xintercept = 4) +
  theme_classic()

ggplot(data.frame(x=c(0:10)), aes(x)) +
    geom_point(aes(y=dpois(x, 1)), colour="red")
```

#
```{r}
disco = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/discoRes_840_ad.txt", col.names = c("CHROM", "POS", "c1", "c2"))

slim = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_840.txt", col.names = c("CHROM", "POS", "snp_af", "ac"))

disco$kmer_min_af = pmin(disco$c1, disco$c2)/(disco$c1 + disco$c2)

slim$snp_min_af = pmin(slim$snp_af, 1 - slim$snp_af)

all_min_af = merge(slim, disco, by = c("CHROM", "POS"), all.x = T, all.y = T)

cor.test(all_min_af$snp_min_af, all_min_af$kmer_min_af, method = "spearman")

ggplot(all_min_af, aes(x = snp_min_af, y = kmer_min_af)) +
  geom_point() +
  theme_classic() +
  geom_abline(slope = 1, intercept = 0)
```

# Test zeta distribution
```{r}
pA = 0.25
pC = 0.25
pG = 0.25
pT = 0.25
shape_parameter = 2.5
k = 31
L = 1e5

alphabet = c("A", "C", "G", "T")

custom_probs = c(pA, pC, pG, pT)

total_genome = c()

while(k*length(total_genome) < L){

  # generate random k-mer
  kmer = paste(sample(alphabet, size = k, replace = T, prob = custom_probs), collapse = "")

  # detmerine frequency of k-mer
   freq = rzeta(1, shape_parameter)
  #freq = 1

  # copy k-mer 
  kmers = rep(kmer, freq)

  # add k-mers to total_genome
  total_genome = c(total_genome, kmers)
}

# randomly order vector elements
total_genome = sample(total_genome, replace = F)

# concatenate k-mers together
total_genome = paste(total_genome, collapse = "")

# final genome length
nchar(total_genome)

# trim genome to target length
total_genome = substring(total_genome, 1, L)

# final genome length
nchar(total_genome)

# write genome output
#print(total_genome)

# count k-mers in synthetic genome to confirm zipf distribution
start = 1
stop = k

kmer_list = list()

while(stop < L){
  
  kmer = substring(total_genome, start, stop)
  
  if(is.null(kmer_list[[kmer]])){
      kmer_list[[kmer]] = 1
  } else {
    kmer_list[[kmer]] = kmer_list[[kmer]] + 1
  }

  start = start + 1
  stop = stop + 1
}

kmer_freqs = table(unlist(kmer_list))

plotdata = data.frame(
  freqs = as.numeric(names(kmer_freqs)),
  ranks = kmer_freqs
)

#plotdata$ranks.Var1 = as.numeric(plotdata$ranks.Var1)

#plotdata$ranks.Var1 = rank(plotdata$ranks.Freq, ties.method = "min")

ggplot(plotdata, aes(x = log10(freqs), y = log10(ranks.Freq))) +
  geom_point() +
  theme_classic() +
  labs(x = "log10(k-mer count)", y = "log10(Number of k-mers)")

```

# DEPRECATED
```{r}
kmer_counts = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/kmers_poolseq/kmer_counts_1.txt", col.names = c("k1", "c1"))

ggplot(kmer_counts, aes(x = c1)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 5) +
  theme_classic()

# loop over k-mers to find ones that putative snps
alphabet = c("A","T","G","C")

results = list()

j = 1
while(nrow(kmer_counts) > 0){
  # get focal k-mer
  focal_kmer = kmer_counts[1,"k1"]
  focal_counts = kmer_counts[1,"c1"]
  print(focal_kmer)

  # generate variant k-mers
  central_base = substr(focal_kmer, 16, 16)

  var_bases = alphabet[!(alphabet %in% central_base)]

  variant_kmers = rep(focal_kmer, times = 3)

  for(i in 1:3){
    var_base = var_bases[i]
    substr(variant_kmers[i], 16, 16) = var_base
  }

  # add reverse complements
  variant_kmers = c(variant_kmers, chartr("ATGC","TACG",variant_kmers))

  # 
  variant_kmers_index = which(kmer_counts$k1 %in% variant_kmers)
  print(variant_kmers_index)

  # store 
  found_variant_kmers = kmer_counts[variant_kmers_index,"k1"]
  found_variant_counts = kmer_counts[variant_kmers_index,"c1"]
  
  if(!identical(found_variant_kmers, character(0))){
      results[[j]] = data.frame(focal_kmer,
                            focal_counts,
                            found_kmers = paste(found_variant_kmers, sep = ";"),
                            found_counts = paste(found_variant_counts, sep = ";"))

    j = j + 1
  }

  # remove focal and variant k-mers from count table
  kmer_counts = kmer_counts[c(-1, -1*variant_kmers_index),]

}


```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
