---
title: "Testing the utility of allele-specific k-mer counts in pool-seq data"
output: html_notebook
---

# set-up environment

```{r}
# clear env
rm(list = ls())
gc()

# packages
library(vcfR)
library(ggplot2)
library(dplyr)
library(MASS)
library(reshape2)
library(ggpubr)
library(ggforce)
library(VGAM)
library(fitPS)
library(scico)
library(stringr)
library(XML)

# date to mark plots/outputs
today <- Sys.Date()

# global ggplot2 theme
theme_set(
  theme_minimal() +
    theme(
      text = element_text(
        size = 16,
        family = "Helvetica",
        colour = "black"
      ),
      axis.text = element_text(colour = "black"),
      plot.title = element_text(hjust = 0.5),
      strip.text = element_text(colour = "black")
    )
)
```

# common functions
```{r}
# Source - https://stackoverflow.com/a
# Posted by Zheyuan Li, modified by community. See post 'Timeline' for change history
# Retrieved 2025-12-29, License - CC BY-SA 4.0
rsq <- function(x, y) cor(x, y, use = "pairwise.complete.obs")^2

# expected number of segregating sites
ES <- function(N, mu, L, n) {
  4 * N * mu * L * sum(1 / (1:(n - 1)))
}

# reverse a string
strReverse <- function(x) sapply(lapply(strsplit(x, NULL), rev), paste, collapse = "")

# HWE heterozygosity at a single site
diversity <- function(p) {
  1 - (p^2 + (1 - p)^2)
}

# fucntion to calculate rates and accuracy
calc_rates <- function(truth, test, sample_space) {
  negatives <- sample_space[!(sample_space %in% truth)]
  positives <- truth

  true_positives <- sum(test %in% positives)
  false_positives <- sum(!(test %in% positives))

  true_negatives <- sum(!(negatives %in% test))
  false_negatives <- sum(!(positives %in% test))

  # accuracy
  acc <- (true_positives + true_negatives) / (true_positives + true_negatives + false_positives + false_negatives)

  # true positives rate (sensitivity)
  tpr <- true_positives / (true_positives + false_negatives)

  # false positive rate (false alarm)
  fpr <- false_positives / (false_positives + true_negatives)

  # true negative rate (specificity)
  tnr <- true_negatives / (true_negatives + false_positives)

  # false negative rate (miss)
  fnr <- false_negatives / (true_positives + false_negatives)

  return(c(length(positives), length(negatives), true_positives, false_positives, true_negatives, false_negatives, acc, tpr, fpr, tnr, fnr))
}
```

# Get genome size statistics from NCBI
```{r message=FALSE, warning=FALSE, include=FALSE}
assembly_info <- read.table("../../data/ncbi_dataset_2025-12-11.tsv", header = T, sep = "\t", fill = T)

results <- NULL

for (i in 1:nrow(assembly_info)) {
  taxid <- assembly_info$Organism.Taxonomic.ID[i]

  assembly_length <- assembly_info$Assembly.Stats.Total.Sequence.Length[i]

  testfile <- "/global/home/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/notebooks/test.xml"

  url <- paste("https://api.ncbi.nlm.nih.gov/genome/v0/expected_genome_size?species_taxid=", taxid, "&length=", assembly_length, sep = "")

  download.file(url, testfile, mode = "wb") # get data into test

  data <- xmlParse(testfile)

  xml_data <- xmlToList(data)

  expected_ungapped_length <- xml_data$expected_ungapped_length

  if (is.null(expected_ungapped_length)) {
    # minimum_ungapped_length = as.numeric(xml_data$minimum_ungapped_length)
    # maximum_ungapped_length = as.numeric(xml_data$maximum_ungapped_length)
    # expected_ungapped_length = (minimum_ungapped_length + maximum_ungapped_length)/2
    # if(is.null(expected_ungapped_length)){
    expected_ungapped_length <- NA
    # }
  }

  result <- data.frame(
    Organism.Taxonomic.ID = taxid,
    assembly_percent_complete = assembly_length / as.numeric(expected_ungapped_length) * 100,
    expected_ungapped_length = expected_ungapped_length,
    length_status = xml_data$length_status
  )

  results <- rbind(results, result)

  random_wait <- rexp(1, 1 / 3)

  Sys.sleep(random_wait)
}

write.table(results, paste("assembly_completeness_stats_", today, ".tsv", sep = ""), sep = "\t", row.names = F, quote = F)
```

# Fit zeta distributions to k-mer count histograms and compare to genome size estimates
```{r}
my_histos <- list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/kmc_histo_results/", pattern = ".histo")

ref_seq_ids_to_plot <- c("GCF_000001735.4", "GCF_902167145", "GCF_036512215", "GCF_000001405.40", "GCF_000001215.4", "GCF_000002985.6")

histos_to_plot <- my_histos[grepl(paste(ref_seq_ids_to_plot, collapse = "|"), my_histos)]

results <- NULL
num_skipped <- 0

for (one_histo in my_histos) {
  # for(one_histo in histos_to_plot) {
  # Initialize a flag to indicate if an error occurred in this iteration
  skip_to_next <- FALSE

  histo <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/kmc_histo_results/", one_histo, sep = ""), header = F)

  # plot(log10(histo$V1), log10(histo$V2))

  data <- histo$V2

  data <- data[1:(pmin(min(which(data == 0)) - 1, length(data)))]

  p1 <- makePSData(n = 1:length(data), count = data, type = "S")

  tryCatch(
    expr = {
      fit <- fitDist(p1)
    },
    error = function(e) {
      # If an error occurs, set the flag to TRUE
      message(paste("Error in iteration:", one_histo, e$message))
      num_skipped <<- num_skipped + 1
      skip_to_next <<- TRUE # Use superassignment to modify the flag in the enclosing environment
    },
    warning = function(w) {}
  )

  if (skip_to_next) {
    next
  }

  s <- fit$shape

  # genome length
  L <- sum(data)

  prob_mass <- data / L

  # calculate r-squared on log-log scale
  compare <- data.frame(
    x = 1:length(data),
    data = prob_mass,
    zeta = dzeta(1:length(data), shape = s)
  )

  r2 <- rsq(log10(compare$data), log10(compare$zeta))

  # sse = sum( (compare$data - compare$zeta)^2 )
  # sst = sum( (compare$data - mean(compare$data))^2 )

  # r2 = 1 - sse/sst

  # plot some fits
  if (one_histo %in% histos_to_plot) {
    print("Plotting...")

    compare <- melt(compare, id = "x")

    plot_point_zeta_example <- ggplot(compare, aes(x = x, y = value, group = variable)) +
      geom_point(shape = 16) +
      geom_line(aes(linetype = variable)) +
      scale_y_log10() +
      scale_x_log10() +
      theme(
        legend.title = element_blank(),
        legend.position = c(0.8, 0.8),
        plot.title = element_text(hjust = 0.5)
      ) +
      labs(x = "k-mer count in genome", y = "Probability mass", title = bquote(R^2 == .(signif(r2, 3))))

    ggsave(paste("../../results/", today, "/", gsub(".histo", "", one_histo), "_zeta_vs_data.pdf", sep = ""), device = cairo_pdf, width = 6, height = 4)
  }

  result <- data.frame(
    genome = one_histo,
    perc_uniq_kmers = prob_mass[1],
    shape = s,
    length = L,
    r2 = r2
  )

  results <- rbind(results, result)
}

# separate data by k
results$k <- NA
results[grepl("_27.histo", results$genome), "k"] <- 27
results[grepl("_31.histo", results$genome), "k"] <- 31
results[grepl("_41.histo", results$genome), "k"] <- 41
results[grepl("_51.histo", results$genome), "k"] <- 51
results[grepl("_61.histo", results$genome), "k"] <- 61

# separate plants and animals
plants <- read.table("../../data/viridiplantae_ncbi_ids_2025-12-12.tsv", fill = T, header = T, sep = "\t")

animals <- read.table("../../data/metazoa_ncbi_ids_2025-12-13.tsv", fill = T, header = T, sep = "\t")

plants$Assembly.Accession <- gsub("\\.[0-9]", "", plants$Assembly.Accession)

results$Assembly.Accession <- gsub("\\.[0-9]_.*", "", results$genome)

results$kingdom <- NA

results$kingdom[(results$Assembly.Accession %in% plants$Assembly.Accession)] <- "Plants"

results$kingdom[is.na(results$kingdom)] <- "Animals"

table(results$kingdom)

# get three values for sweep sims
summary(results[, "shape"])
summary(results[(results$k == 27), "shape"])
summary(results[(results$k == 31), "shape"])
summary(results[(results$k == 41), "shape"])
summary(results[(results$k == 51), "shape"])
summary(results[(results$k == 61), "shape"])

# label specific species to orient reader
extra_point_data <- data.frame(
  k = c(
    27, 27, 27,
    31, 31, 31,
    41, 41, 41,
    51, 51, 51,
    61, 61, 61,
    27, 27, 27,
    31, 31, 31,
    41, 41, 41,
    51, 51, 51,
    61, 61, 61
  ),
  shape = c(
    1.660557, 2.753124, 3.824493,
    1.727444, 2.940929, 3.946757,
    1.885970, 3.298644, 4.193795,
    2.031053, 3.564381, 4.382319,
    2.168866, 3.767132, 4.546686,
    2.875324, 3.528856, 4.023279,
    2.938262, 3.560309, 4.122031,
    3.061762, 3.550773, 4.292060,
    3.153887, 3.585911, 4.447315,
    3.230458, 3.603063, 4.554442
  ),
  species = c(
    "maize", "tomato", "arabidopsis",
    "maize", "tomato", "arabidopsis",
    "maize", "tomato", "arabidopsis",
    "maize", "tomato", "arabidopsis",
    "maize", "tomato", "arabidopsis",
    "human", "fly", "worm",
    "human", "fly", "worm",
    "human", "fly", "worm",
    "human", "fly", "worm",
    "human", "fly", "worm"
  ),
  kingdom = c(
    "Plants", "Plants", "Plants",
    "Plants", "Plants", "Plants",
    "Plants", "Plants", "Plants",
    "Plants", "Plants", "Plants",
    "Plants", "Plants", "Plants",
    "Animals", "Animals", "Animals",
    "Animals", "Animals", "Animals",
    "Animals", "Animals", "Animals",
    "Animals", "Animals", "Animals",
    "Animals", "Animals", "Animals"
  )
)

# distribution of shape parameter
shape_sample_size <- results %>% dplyr::count(k, kingdom)

plot_violin_boxplot_shape <- ggplot(results, aes(x = as.factor(k), y = shape)) +
  geom_violin(lwd = 1.1) +
  facet_wrap(~kingdom, ncol = 1) +
  geom_point(
    data = extra_point_data,
    mapping = aes(
      x = as.factor(k),
      y = shape,
      color = species
    ),
    size = 3
  ) +
  scale_color_manual(
    values = c(
      "maize" = "yellow3",
      "tomato" = "pink2",
      "arabidopsis" = "green1",
      "human" = "black",
      "fly" = "grey40",
      "worm" = "grey80"
    )
  ) +
  geom_boxplot(width = 0.25, color = "grey", alpha = 0.4) +
  geom_text(data = shape_sample_size, aes(as.factor(k), 17.5, label = paste("n =", n)), vjust = 1.25) +
  theme(legend.position = "None", axis.title.y = element_text(angle = 0, vjust = 0.5)) +
  labs(x = "k", y = "\u03b6 shape parameter") +
  coord_flip()

ggsave(paste("../../results/", today, "/shape_dist.pdf", sep = ""), device = cairo_pdf, width = 5, height = 5)

# percent unique k-mers
# ggplot(results, aes(x = as.factor(k), y = perc_uniq_kmers)) +
#   geom_violin(lwd = 1.1) +
#   facet_wrap(~ kingdom, ncol = 1) +
#   theme_minimal() +
#   theme(text = element_text(size = 20, family = "Helvetica", colour = "black"),
#         axis.text = element_text(colour = "black"),
#         legend.position = "None",
#         axis.title.y = element_text(angle = 0, vjust = 0.5)) +
#   labs(x = "k", y = "Percent uni-mers") +
#   coord_flip()

# compare plants vs animals
median(results[(results$k == 31) & (results$kingdom == "Plants"), "shape"])
median(results[(results$k == 31) & (results$kingdom == "Animals"), "shape"])

wilcox.test(results[(results$k == 31) & (results$kingdom == "Plants"), "shape"], results[(results$k == 31) & (results$kingdom == "Animals"), "shape"])

# zeta model fits
summary(results$r2)
sum(results$r2 > 0.9) / nrow(results)

# compare genome sizes to shape parameter
genome_sizes <- read.table("assembly_completeness_stats_2025-12-11.tsv", header = T)

summary(genome_sizes$expected_ungapped_length)

plant_ids_genome_size <- merge(plants, genome_sizes, by = "Organism.Taxonomic.ID")

results_genome_size <- merge(results, plant_ids_genome_size, by = "Assembly.Accession")

plot_point_size_shape <- ggplot(results_genome_size, aes(x = expected_ungapped_length, y = shape)) +
  geom_point() +
  geom_smooth(color = "red", lwd = 1.5) +
  facet_wrap(. ~ k, ncol = 5, labeller = label_both) +
  theme_minimal() +
  scale_x_log10() +
  theme(
    text = element_text(
      size = 16,
      family = "Helvetica",
      colour = "black"
    ),
    axis.text = element_text(colour = "black")
  ) +
  scale_y_continuous(limits = c(1, 7)) +
  labs(x = "Expected genome size (bp)", y = "\u03b6 shape parameter")

ggsave(paste("../../results/", today, "/shape_vs_genome_size.pdf", sep = ""), device = cairo_pdf, width = 9, height = 3)

# shape vs genome size for only k = 31
r2_size_shape_k31 <- rsq(log10(results_genome_size[(results_genome_size$k == 31), "expected_ungapped_length"]), results_genome_size[(results_genome_size$k == 31), "shape"])

plot_point_size_shape_k31 <- ggplot(results_genome_size[(results_genome_size$k == 31), ], aes(x = expected_ungapped_length, y = shape)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", lwd = 1.5) +
  # facet_wrap(. ~ k, ncol = 5, labeller = label_both) +
  theme_minimal() +
  scale_x_log10() +
  theme(
    text = element_text(
      size = 16,
      family = "Helvetica",
      colour = "black"
    ),
    axis.text = element_text(colour = "black")
  ) +
  scale_y_continuous(limits = c(1, 6)) +
  labs(x = "Expected genome size (bp)", y = "\u03b6 shape parameter", title = bquote(R^2 == .(signif(r2_size_shape_k31, 3))))

ggsave(paste("../../results/", today, "/shape_vs_genome_size_k31.pdf", sep = ""), device = cairo_pdf, width = 9, height = 3)

# demonstrate how zeta distribution sampling creates small genomes with power-law k-mer count distributions
# these will be examples of the frequency distributions for the genomes in the simulations
# except the simulated genomes will have additional unique k-mers because I'm not counting the k-mers at the junctions between chunks
# so the simulated genomes will still probably have lots of unique k-mers than the shape parameter alone would suggest, but the shape parameter still captures relative repeat content
zeta_sampling <- expand.grid(
  k = c(27, 31, 41),
  L = c(1e6, 2e6, 4e6),
  # s = c(3.591, 1e6)
  s = c(2.854, 3.591, 4.772, 1e6)
)

zeta_sampling_results <- NULL

for (i in 1:nrow(zeta_sampling)) {
  print(i)

  k <- zeta_sampling[i, 1]
  L <- zeta_sampling[i, 2]
  s <- zeta_sampling[i, 3]

  zeta_samples <- c()
  total <- 0

  if (s < 1e6) {
    while (total < L) {
      result <- suppressWarnings(rzeta(1, s))
      zeta_samples <- c(zeta_samples, result)
      total <- total + result * k
    }
    result <- data.frame(
      k = k,
      L = L,
      s = s,
      ranks = table(zeta_samples)
    )
    # assume all k-mers at junctions are unique
    result$ranks.Freq[(result$ranks.zeta_samples == 1)] <- result$ranks.Freq[(result$ranks.zeta_samples == 1)] + (sum(result$ranks.Freq) - 1) * (k - 1)
  } else {
    result <- data.frame(
      k = k,
      L = L,
      s = s,
      ranks.zeta_samples = 1,
      ranks.Freq = L - k + 1
    )
  }
  zeta_sampling_results <- rbind(zeta_sampling_results, result)
}

# compare zeta distribution by L and s and k
# first reorder points to plot no-repeats first
zeta_sampling_results <- zeta_sampling_results[order(zeta_sampling_results$s, decreasing = T), ]

plot_point_zeta_sampling <- ggplot(zeta_sampling_results, aes(x = as.numeric(ranks.zeta_samples), y = as.numeric(ranks.Freq), shape = as.factor(k), color = as.factor(s), size = as.factor(s))) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_scico_d(palette = "lajolla", begin = 0, end = 0.9, direction = -1) +
  scale_shape_manual(values = c(16, 17, 18)) +
  scale_size_manual(values = c(1.5, 1.5, 1.5, 3)) + # emphasize no-repeat
  facet_wrap(~L, labeller = label_both) +
  guides(size = "none") +
  theme(plot.title = element_text(hjust = 1)) +
  labs(x = "k-mer count in genome", y = "Number of k-mers", color = "s", shape = "k")

ggsave(paste("../../results/", today, "/zeta_ksL.pdf", sep = ""), device = cairo_pdf, width = 9, height = 3)

# combine zeta plots into one figure
plot_fig1 <- grid.arrange(plot_point_zeta_example, plot_violin_boxplot_shape, plot_point_size_shape_k31, plot_point_zeta_sampling, layout_matrix = matrix(c(1, 4, 2, 4, 3, 4), nrow = 2), heights = c(1, 0.75))

# plot_fig1_patchwork <- ((plot_point_zeta_example + plot_violin_boxplot_shape + plot_point_size_shape_k31) / plot_point_zeta_sampling) + plot_annotation(tag_levels = "A")

ggsave(paste("../../results/", today, "/fig1.pdf", sep = ""), plot_fig1, device = cairo_pdf, width = 10, height = 7)
```

# load simulation parameters

```{r}
params <- read.table("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/config/parameters.tsv", header = T)

# get only simulations that completed
#dones <- sort(as.numeric(gsub("_0", "", gsub(".done", "", list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/finished/", pattern = ".done")))))

dones <- sort(as.numeric(gsub("_0", "", gsub("_calls.txt", "", list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freqk_results/", pattern = "_calls.txt")))))

params <- params[(params$ID %in% dones),]
```

# benchmarks - runtime and memory

```{r}

bench <- read.table("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/bench.txt", sep = "\t", col.names = c("file", "s", "h:m:s", "max_rss", "max_vms", "max_uss", "max_pss", "io_in", "io_out",  "mean_load", "cpu_time"))

bench_file_names <- list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks", recursive = T, full.names = T)

bench_files <- lapply(bench_file_names, read.table, header = T, sep = "\t")

#bench_files <- lapply(bench_file_names, fread, header = T, sep = "\t")

bench_files <- do.call("rbind", bench_files)

#bench_files <- vroom(bench_file_names, id = "files", altrep = F)

#bench_files <- map_dfr(bench_file_names, read.table, header = T, sep = "\t")

# get names
bench_files$files <- bench_file_names

bench_files$tool <- gsub("/.*", "", gsub("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/", "", bench_files$files))

bench_files$ID <- as.numeric(gsub(".*/|.bench|_0.bench|.bam|_0.bam|unpaired_R._|paired_", "", gsub("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/", "", bench_files$files)))

# merge with workflow parameters
bench_params <- merge(params[(params$simtype == "onepop"), c("ID", "cov", "n")], bench_files, by = "ID")

# time for each workflow step
bench_summary <- bench_params %>%
  group_by(ID) %>%
  summarize(
    freqk_index = sum(s[(tool == "freqk_index") | (tool == "freqk_var_dedup")  | (tool == "freqk_ref_dedup")], na.rm = TRUE),
    freqk_count = sum(s[(tool == "freqk_count") | (tool == "freqk_call")], na.rm = TRUE),
    vg_autoindex = max(s[tool == "vg_autoindex"], na.rm = TRUE),
    vg_giraffe_varscan = max(s[tool == "vg_giraffe"], na.rm = TRUE) + max(s[tool == "vg_surject"], na.rm = TRUE) + max(s[tool == "samtools_sort"], na.rm = TRUE) + max(s[tool == "samtools_merge"], na.rm = TRUE) + max(s[tool == "varscan"], na.rm = TRUE),
    vg_giraffe_poolsnp = max(s[tool == "vg_giraffe"], na.rm = TRUE) + max(s[tool == "vg_surject"], na.rm = TRUE) + max(s[tool == "samtools_sort"], na.rm = TRUE) + max(s[tool == "samtools_merge"], na.rm = TRUE) + max(s[tool == "poolsnp"], na.rm = TRUE),
    vg_giraffe_freebayes = max(s[tool == "vg_giraffe"], na.rm = TRUE) + max(s[tool == "vg_surject"], na.rm = TRUE) + max(s[tool == "samtools_sort"], na.rm = TRUE) + max(s[tool == "samtools_merge"], na.rm = TRUE) + max(s[tool == "freebayes_vg"], na.rm = TRUE),
    .groups = 'drop' # Recommended to drop grouping after summarization
  )

bench_summary <- melt(bench_summary, id = "ID")

bench_summary <- merge(params, bench_summary, by = "ID")

bench_summary$index_step <- "Calling"
bench_summary$index_step[(bench_summary$variable == "freqk_index") | ((bench_summary$variable == "vg_autoindex"))] <- "Indexing"

bench_summary$variable <- gsub("_", " ", bench_summary$variable)
bench_summary$variable <- gsub("giraffe ", "giraffe+", bench_summary$variable)

plot_bench_s_call <- ggplot(subset(bench_summary, index_step == "Calling"), aes(x = variable, y = value, color = as.factor(cov))) +
  geom_boxplot() +
  #facet_wrap(index_step~., scales = "free") +
  #theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "top", legend.title = element_text(hjust = 0.5)) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "top",
        legend.title = element_text(hjust = 0.5)) +
  guides(colour = guide_legend(title.position = "top")) +
  scale_color_scico_d(palette = "hawaii") +
  labs(y = "Runtime (s)", color = "Coverage", x = "Workflow")

plot_bench_s_index <- ggplot(subset(bench_summary, index_step == "Indexing"), aes(x = variable, y = value, color = as.factor(n))) +
  geom_boxplot() +
  #facet_wrap(index_step~., scales = "free") +
  #theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "top", legend.title.align=0.5) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "top",
        legend.title = element_text(hjust = 0.5)) +
  guides(colour = guide_legend(title.position = "top")) +
  scale_color_scico_d(palette = "lajolla", begin = 0.05, end = 0.95) +
  labs(y = "Runtime (s)", color = "Pool size", x = "Workflow")

ggarrange(plot_bench_s_index, plot_bench_s_call, nrow = 1, ncol = 2, labels = c("A", "B"), font.label = list(size = 16, family = "Helvetica"))

ggsave(paste("../../results/", today, "/time.pdf", sep = ""), height = 4, width = 8)

## plot memory

bench_memory <- bench_params %>%
  group_by(ID) %>%
  summarize(
    freqk_index = max(max_rss[(tool == "freqk_index") | (tool == "freqk_var_dedup")  | (tool == "freqk_ref_dedup")], na.rm = TRUE),
    freqk_count = max(max_rss[(tool == "freqk_count") | (tool == "freqk_call")], na.rm = TRUE),
    vg_autoindex = max(max_rss[tool == "vg_autoindex"], na.rm = TRUE),
    vg_giraffe_varscan = max(max_rss[(tool == "vg_giraffe") | (tool == "vg_surject") | (tool == "samtools_sort") | (tool == "samtools_merge") | (tool == "varscan")], na.rm = TRUE),
    vg_giraffe_poolsnp = max(max_rss[(tool == "vg_giraffe") | (tool == "vg_surject") | (tool == "samtools_sort") | (tool == "samtools_merge") | (tool == "poolsnp")], na.rm = TRUE),
    vg_giraffe_freebayes = max(max_rss[(tool == "vg_giraffe") | (tool == "vg_surject") | (tool == "samtools_sort") | (tool == "samtools_merge") | (tool == "freebayes_vg")], na.rm = TRUE),
    .groups = "drop"
)

bench_memory <- melt(bench_memory, id = "ID")

bench_memory <- merge(params, bench_memory, by = "ID")

bench_memory$index_step <- "Calling"
bench_memory$index_step[(bench_memory$variable == "freqk_index") | ((bench_memory$variable == "vg_autoindex"))] <- "Indexing"

bench_memory$variable <- gsub("_", " ", bench_memory$variable)
bench_memory$variable <- gsub("giraffe ", "giraffe+", bench_memory$variable)

plot_bench_mem_call <- ggplot(subset(bench_memory, index_step == "Calling"), aes(x = variable, y = value, color = as.factor(cov))) +
  geom_boxplot() +
  facet_zoom(y = value <= 2000, show.area = T, zoom.size = 1.5) +
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "none", legend.title = element_text(hjust = 0.5), zoom = element_rect(colour = "red", linewidth = 0.3, fill = NA)) +
  scale_color_scico_d(palette = "hawaii") +
  labs(y = "Memory (Mb)", color = "Coverage", x = "Workflow")

plot_bench_mem_index <- ggplot(subset(bench_memory, index_step == "Indexing"), aes(x = variable, y = value, color = as.factor(n))) +
  geom_boxplot() +
  #facet_wrap(index_step~., scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "none") +
  scale_color_scico_d(palette = "lajolla", begin = 0.05, end = 0.95) +
  labs(y = "Memory (Mb)", color = "Pool size", x = "Workflow")

ggarrange(plot_bench_mem_index, plot_bench_mem_call, nrow = 1, ncol = 2, labels = c("C", "D"), font.label = list(size = 16, family = "Helvetica"))

ggsave(paste("../../results/", today, "/memory.pdf", sep = ""), height = 4, width = 8)

## bring it all together

ggarrange(plot_bench_s_index, plot_bench_s_call, plot_bench_mem_index, plot_bench_mem_call, nrow = 2, ncol = 2, labels = c("A", "B", "C", "D"), font.label = list(size = 16, family = "Helvetica"))

ggsave(paste("../../results/", today, "/time_memory.pdf", sep = ""), height = 7, width = 9)
```

# benchmarks - accuracy

```{r}

results <- NULL

to_plot <- c(10:12)

for (i in params$ID) {
  print(paste("Processing simulation ", i, "...", sep = ""))
  
  # load in actual allele freqs from simulations
  true_af <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/slim_results/allele_freqs_", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"))

  true_af$ac <- as.numeric(true_af$ac)

  true_af <- true_af[complete.cases(true_af), ]
  
  # include varscan calls
  varscan <- tryCatch({
    read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/varscan_results/", i, "_0.tsv", sep = ""), header = T)
  }, error=function(e){})
  
  if(is.null(varscan)){
    next
  }

  #varscan <- varscan[(varscan$Reads1 >= thresh) & (varscan$Reads2 >= thresh), ]

  #varscan$snp_maf_varscan <- pmin(varscan$Reads1, varscan$Reads2) / (varscan$Reads1 + varscan$Reads2)

  names(varscan)[1] <- "chrom"
  names(varscan)[2] <- "pos"

  varscan$varscan_af <- as.numeric(gsub("%", "", varscan$VarFreq))/100
  
  # load poolsnp calls
  poolsnp <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/poolsnp_results/", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ac", "an"), fill = T, na.strings = "")

  poolsnp$poolsnp_af <- poolsnp$ac/poolsnp$an

  #poolsnp$snp_maf_poolsnp <- pmin(poolsnp$c1 / poolsnp$t, 1 - poolsnp$c1 / poolsnp$t)

  # include freebayes calls
  freebayes <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freebayes_vg_results/", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "freebayes_af", "ac"), fill = T, na.strings = "")

  #freebayes$snp_maf_freebayes <- pmin(freebayes$af, 1 - freebayes$af)
  
  #
  freqk_freq <- tryCatch({
    read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freqk_results/", i, "_0_calls.txt", sep = ""), sep = "|", header = F, col.names = c("rf", "freqk_af"))}
  , error = function(e){}  
  )
  
  if(is.null(freqk_freq)){
    next
  }
  
  freqk_index <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freqk_ref_dedup/", i, "_0.txt", sep = ""), sep = ",", header = F)
  
  freqk <- cbind(freqk_index, freqk_freq)
  
  names(freqk)[2:3] <- c("chrom", "pos")
  
  freqk$pos <- freqk$pos + 2
  
  # merge all results with ground truth
  true_af <- merge(true_af[,c("chrom", "pos", "af")], freqk[,c("chrom", "pos", "freqk_af")], by = c("chrom", "pos"), all.x = T)
    
  true_af <- merge(true_af, freebayes[,c("chrom", "pos", "freebayes_af")], by = c("chrom", "pos"), all.x = T)
  
  true_af <- merge(true_af, varscan[,c("chrom", "pos", "varscan_af")], by = c("chrom", "pos"), all.x = T)
    
  true_af <- merge(true_af, poolsnp[,c("chrom", "pos", "poolsnp_af")], by = c("chrom", "pos"), all.x = T)
  
  plotdata <- melt(true_af, id = c("chrom", "pos", "af"))
  
  plotdata$variable <- gsub("_af", "", plotdata$variable)
  
  plotdata$variable <- as.factor(plotdata$variable)
  
  levels(plotdata$variable) <- c("vg giraffe+freebayes", "freqk", "vg giraffe+poolsnp", "vg giraffe+varscan")
  
  plotdata$variable <- factor(plotdata$variable, levels = c("freqk", "vg giraffe+freebayes", "vg giraffe+poolsnp", "vg giraffe+varscan"))
  
  
  cor_by_sim <- plotdata %>% 
    dplyr::group_by(variable) %>% 
    summarise(correlation = cor.test(af, value)$estimate, n = n())
  
  cor_by_sim$ID = i
  
  results <- rbind(results, cor_by_sim)
  
  if (i %in% to_plot){
    ggplot(plotdata, aes(x = af, y = value)) +
     geom_point() +
     geom_abline(slope = 1, intercept = 0) +
     facet_wrap(~variable, nrow = 1, ncol = 4) +
     stat_cor(method = "pearson") +
     theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
     labs(x = "True allele frequency", y = "Predicted allele frequency")
   
    ggsave(paste("../../results/", today, "/", i, "_true_vs_pred_af",".pdf", sep = ""), height = 4, width = 10)
  }
}

params_corr <- merge(params, results, by = "ID")

params_corr$variable <- as.factor(params_corr$variable)

ggplot(params_corr, aes(x = variable, y = correlation, color = as.factor(cov))) +
    geom_boxplot() +
    scale_color_scico_d(palette = "hawaii") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "workflow", y = "Correlation\ntrue vs predicted AF", color = "coverage")

ggsave(paste("../../results/", today, "/true_vs_pred_by_cov",".pdf", sep = ""), height = 5, width = 6)

ggplot(params_corr, aes(x = variable, y = correlation, color = as.factor(n.x))) +
    geom_boxplot() +
    scale_color_scico_d(palette = "lajolla", begin = 0.05, end = 0.95) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "workflow", y = "Correlation\ntrue vs predicted AF", color = "Pool size")

ggsave(paste("../../results/", today, "/true_vs_pred_by_n",".pdf", sep = ""), height = 5, width = 6)
```

# single population simulations

```{r}

# empty frame to store results
results <- NULL

# get list of simulation ids
# simids = params$ID[(params$simtype == "onepop") | (params$simtype == "sweep")]
#
# two_pop_ids = expand.grid(params$ID[(params$simtype == "twopop")], c("p1", "p2"))
#
# two_pop_ids = paste(two_pop_ids$Var1, two_pop_ids$Var2, sep = "_")
#
# simids = c(two_pop_ids, simids)

# simids = params$ID[(params$simtype == "onepop")]

simids <- params$ID[(params$simtype == "onepop") & (params$ID %in% dones)]

k <- 31

thresh <- 3

G <- 2e6

# for(i in dones){
# for(i in params$ID[(params$simtype == "onepop") | (params$simtype == "sweep")]){
# for(i in 127:512){
for (i in simids) {
  print(paste("Processing simulation ", i, "...", sep = ""))

  # get sequencing coverage data
  # json_data_R1R2 <- fromJSON(file=paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", i, "_R1R2.json", sep = ""))
  # json_data_U1 <- fromJSON(file=paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", i, "_U1.json", sep = ""))
  # json_data_U2 <- fromJSON(file=paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", i, "_U2.json", sep = ""))
  #
  n <- params[(params$ID == i), "n"] # pool size
  #
  # B = json_data_R1R2$summary$after_filtering$total_bases + json_data_U1$summary$after_filtering$total_bases + json_data_U2$summary$after_filtering$total_bases # number of bases sequenced
  #
  # G = params[(params$ID == i), "L"] # genome size
  #
  # L = (json_data_R1R2$summary$after_filtering$read1_mean_length*json_data_R1R2$summary$after_filtering$total_reads/2 + json_data_R1R2$summary$after_filtering$read2_mean_length*json_data_R1R2$summary$after_filtering$total_reads/2 + json_data_U1$summary$after_filtering$read1_mean_length*json_data_U1$summary$after_filtering$total_reads + json_data_U2$summary$after_filtering$read1_mean_length*json_data_U2$summary$after_filtering$total_reads)/(json_data_R1R2$summary$after_filtering$total_reads + json_data_U1$summary$after_filtering$total_reads + json_data_U2$summary$after_filtering$total_reads) # read length, paired end data, multiply read length by 2
  #
  # C = B/(2*n*G) # base coverage per genome
  # Ck = ((L-k+1)/L)*C # kmer coverage per genome
  # nCk = n*Ck # total kmer coverage
  # nC = n*C # total base coverage

  # load pairs of k-mers
  kmer_pairs <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/smudgeplot_results/", i, "_0_sequences.tsv", sep = ""), col.names = c("k1"))

  kmer_covs <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/smudgeplot_results/", i, "_0_coverages.tsv", sep = ""), col.names = c("c1", "c2"))
  kmer_pairs <- cbind(kmer_pairs, kmer_covs)

  # choose kmer count cutoff
  # all_counts = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/kmer_counts_1.txt", col.names = c("k1", "c1"))
  #
  # ggplot(all_counts, aes(x = c1)) +
  #   geom_histogram(binwidth = 1) +
  #   theme_classic()
  #
  # ggplot(all_counts[(all_counts$c1 >= 4),], aes(x = c1)) +
  #   geom_histogram(binwidth = 1) +
  #   theme_classic()

  # calculate minor allele frequency
  kmer_pairs$kmer_maf <- pmin(kmer_pairs$c1, kmer_pairs$c2) / (kmer_pairs$c1 + kmer_pairs$c2)

  # filter kmers
  # high_nc = qnorm(0.99995, mean = nC, sd = sqrt(nC))
  # low_nc = qnorm(0.00005, mean = nC, sd = sqrt(nC))
  #
  # high_nck = qnorm(0.99995, mean = nCk, sd = sqrt(nCk))
  # low_nck = qnorm(0.00005, mean = nCk, sd = sqrt(nCk))
  #
  # ggplot(kmer_covs, aes(x = c1 + c2)) +
  #   geom_density() +
  #   geom_vline(xintercept = 2*thresh) +
  #   geom_vline(xintercept = low_nc, lty = 2, color = "red") +
  #   geom_vline(xintercept = nC, color = "red") +
  #   geom_vline(xintercept = high_nc, lty = 2, color = "red") +
  #   geom_vline(xintercept = low_nck, lty = 2, color = "black") +
  #   geom_vline(xintercept = nCk, color = "black") +
  #   geom_vline(xintercept = high_nck, lty = 2, color = "black")

  # filter = (kmer_pairs$c1 + kmer_pairs$c2 > low_nck) & (kmer_pairs$c1 + kmer_pairs$c2 < high_nck) & (kmer_pairs$c1 >= thresh) & (kmer_pairs$c2 >= thresh)

  # kmer_pairs = kmer_pairs[(kmer_pairs$c1 >= thresh) & (kmer_pairs$c2 >= thresh),]

  # use bayes theorem to estimate allele counts
  # kmer_pairs$kmer_maf_bayes = mapply(posterior, x = kmer_pairs$c1, z = kmer_pairs$c1 + kmer_pairs$c2, n, C, 1, 1)

  # ggplot(kmer_pairs, aes(x = kmer_maf)) +
  #   geom_density() +
  #   theme_classic()

  # load from my hetmers script
  hetmers_seqs <- read.csv(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/hetmers_results/", i, "_0_seqs.csv", sep = ""), col.names = c("k1", "k2"))

  hetmers_counts <- read.csv(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/hetmers_results/", i, "_0_counts.csv", sep = ""), col.names = c("c1", "c2"))

  hetmers_bayes <- read.csv(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/hetmers_results/", i, "_0_bayes_states.csv", sep = ""), col.names = c("bc"))

  hetmers <- cbind(hetmers_seqs, hetmers_counts, hetmers_bayes)

  # calculate minor allele counts/frequencies
  hetmers$hetmer_maf <- pmin(hetmers$c1, hetmers$c2) / (hetmers$c1 + hetmers$c2)

  hetmers$hetmer_maf_bayes <- pmin(hetmers$bc, 2 * n - hetmers$bc) / (2 * n)

  # load in actual allele freqs from simulations
  true_af <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/slim_results/allele_freqs_", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"))

  true_af$ac <- as.numeric(true_af$ac)

  true_af <- true_af[complete.cases(true_af), ]

  # allele frequency spectrum
  # ggplot(true_af, aes(x = ac)) +
  #   geom_histogram(binwidth = 1) +
  #   theme_classic()

  # calculate minor allele frequencies
  true_af$af <- as.numeric(true_af$af)

  true_af$snp_maf <- pmin(true_af$af, 1 - true_af$af)

  true_af$snp_mac <- pmin(true_af$ac, 2 * n - true_af$ac)

  true_af <- true_af[(true_af$af < 1) & (true_af$af > 0), ]

  # load discosnp results, calculate maf
  disco <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/disco_results/", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "c1", "c2"))

  disco <- disco[(disco$c1 >= thresh) & (disco$c2 >= thresh), ]

  disco$disco_maf <- pmin(disco$c1, disco$c2) / (disco$c1 + disco$c2)

  # get key for position to k-mers
  # if(grepl("_p1|_p2", i)){
  #   j = gsub("_p1|_p2", "_p1p2", i)
  # } else {
  #   j = i
  # }

  pos_kmer_key <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/seqkit_results/center_kmer_pairs_", i, ".txt", sep = ""), fill = T, col.names = c("pos", "k1", "k2", "k3", "k4"))

  pos_kmer_key$leftk <- substr(pos_kmer_key$k1, 1, 15)
  pos_kmer_key$rightk <- substr(pos_kmer_key$k1, 17, 31)

  # compare snp and k-mer mafs
  kmer_pairs$leftk <- substr(kmer_pairs$k1, 1, 15)
  kmer_pairs$rightk <- substr(kmer_pairs$k1, 17, 31)

  hetmers$leftk <- substr(hetmers$k1, 1, 15)
  hetmers$rightk <- substr(hetmers$k1, 17, 31)

  pos_kmer_pairs <- merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

  pos_hetmers <- merge(pos_kmer_key, hetmers, by = c("leftk", "rightk"))

  # get snps that pair with reverse complements
  pos_kmer_key$leftk_rc <- chartr("ATGC", "TACG", pos_kmer_key$leftk)
  pos_kmer_key$rightk_rc <- chartr("ATGC", "TACG", pos_kmer_key$rightk)

  pos_kmer_key$leftk_rc <- strReverse(pos_kmer_key$leftk_rc)
  pos_kmer_key$rightk_rc <- strReverse(pos_kmer_key$rightk_rc)

  pos_kmer_pairs_rc <- merge(pos_kmer_key, kmer_pairs, by.x = c("leftk_rc", "rightk_rc"), by.y = c("rightk", "leftk"))

  pos_hetmers_rc <- merge(pos_kmer_key, hetmers, by.x = c("leftk_rc", "rightk_rc"), by.y = c("rightk", "leftk"))

  pos_kmer_pairs_rc <- subset(pos_kmer_pairs_rc, select = -c(leftk, rightk))

  pos_hetmers_rc <- subset(pos_hetmers_rc, select = -c(leftk, rightk))

  # merge reverse complement and regular sets
  names(pos_kmer_pairs_rc)[1:2] <- c("leftk", "rightk")
  pos_kmer_pairs <- rbind(pos_kmer_pairs, pos_kmer_pairs_rc)

  names(pos_hetmers_rc)[1:2] <- c("leftk", "rightk")
  pos_hetmers <- rbind(pos_hetmers, pos_hetmers_rc)

  # do some filtering
  # remove low coverage k-mer pairs
  # remove tri-allelic k-mers
  # pos_kmer_pairs = pos_kmer_pairs[(pos_kmer_pairs$c1 >=4 & pos_kmer_pairs$c2 >= 4),]
  pos_kmer_pairs <- pos_kmer_pairs[((pos_kmer_pairs$k3 == "") | is.na(pos_kmer_pairs$k3)), ]
  pos_hetmers <- pos_hetmers[((pos_hetmers$k3 == "") | is.na(pos_hetmers$k3)), ]

  compare_maf_kmer <- merge(true_af[, c("pos", "snp_maf", "snp_mac")], pos_kmer_pairs, by = "pos", all.x = T, all.y = T)

  compare_maf_hetmer <- merge(true_af[, c("pos", "snp_maf", "snp_mac")], pos_hetmers, by = "pos", all.x = T, all.y = T)

  # remove all na rows
  # compare_maf_kmer = compare_maf_kmer[rowSums(is.na(compare_maf_kmer)) != ncol(compare_maf_kmer), ]

  # compare minor allele counts
  # compare_maf_kmer_no_na = compare_maf_kmer[!is.na(compare_maf_kmer$c1),]

  # compare_maf_kmer_no_na$kmer_mac = unlist(mapply(posterior_min_kmer_count, x = compare_maf_kmer_no_na$c1, z = compare_maf_kmer_no_na$c1 + compare_maf_kmer_no_na$c2, MoreArgs = list(n = 100, C = 2*0.9, c = 5, alpha = 1, beta = 1)))

  # include varscan calls
  varscan <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/varscan_results/", i, "_0.tsv", sep = ""), header = T)

  varscan <- varscan[(varscan$Reads1 >= thresh) & (varscan$Reads2 >= thresh), ]

  varscan$snp_maf_varscan <- pmin(varscan$Reads1, varscan$Reads2) / (varscan$Reads1 + varscan$Reads2)

  names(varscan)[2] <- "pos"

  # load poolsnp calls
  poolsnp <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/poolsnp_results/", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "c1", "t"), fill = T, na.strings = "")

  poolsnp <- poolsnp[(poolsnp$c1 >= thresh), ]

  poolsnp$snp_maf_poolsnp <- pmin(poolsnp$c1 / poolsnp$t, 1 - poolsnp$c1 / poolsnp$t)

  # include freebayes calls
  freebayes <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freebayes_de_novo_results/", i, "_0.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"), fill = T, na.strings = "")

  freebayes$snp_maf_freebayes <- pmin(freebayes$af, 1 - freebayes$af)

  # freebayes = freebayes[(freebayes$af > 0) & (freebayes$af < 1),]

  # compare varscan and ground truth
  compare_maf_varscan <- merge(true_af[, c("pos", "snp_maf")], varscan, by = "pos", all.x = T, all.y = T)

  # compare poolsnp and ground truth
  compare_maf_poolsnp <- merge(true_af[, c("pos", "snp_maf")], poolsnp, by = "pos", all.x = T, all.y = T)

  # compare freebayes to ground truth
  compare_maf_freebayes <- merge(true_af[, c("pos", "snp_maf")], freebayes, by = "pos", all.x = T, all.y = T)

  # compare discosnp and ground truth
  compare_maf_disco <- merge(true_af[, c("pos", "snp_maf")], disco, by = "pos", all.x = T, all.y = T)

  # compare k-mer and varscan
  # compare_maf_kmer_varscan = merge(compare_maf_kmer, varscan, by = "pos")

  # calculate correlations and performance stats
  kmer_vs_truth_maf_cor <- cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$kmer_maf, method = "pearson")$estimate

  hetmer_vs_truth_maf_cor <- cor.test(compare_maf_hetmer$snp_maf, compare_maf_hetmer$hetmer_maf, method = "pearson")$estimate

  hetmer_vs_truth_bc_cor <- cor.test(compare_maf_hetmer$snp_mac, compare_maf_hetmer$bc, method = "pearson")$estimate

  disco_vs_truth_maf_cor <- tryCatch(
    cor.test(compare_maf_disco$snp_maf, compare_maf_disco$disco_maf, method = "pearson")$estimate,
    error = function(err) NA
  )

  varscan_vs_truth_maf_cor <- cor.test(compare_maf_varscan$snp_maf, compare_maf_varscan$snp_maf_varscan, method = "pearson")$estimate

  poolsnp_vs_truth_maf_cor <- tryCatch(
    cor.test(compare_maf_poolsnp$snp_maf, compare_maf_poolsnp$snp_maf_poolsnp, method = "pearson")$estimate,
    error = function(err) NA
  )

  freebayes_vs_truth_maf_cor <- cor.test(compare_maf_freebayes$snp_maf, compare_maf_freebayes$snp_maf_freebayes, method = "pearson")$estimate

  varscan_rates <- calc_rates(true_af$pos, varscan$pos, 1:G)

  poolsnp_rates <- calc_rates(true_af$pos, poolsnp$pos, 1:G)

  freebayes_rates <- calc_rates(true_af$pos, freebayes$pos, 1:G)

  kmer_rates <- calc_rates(true_af$pos, c(pos_kmer_pairs$pos, rep(1e9, times = nrow(kmer_pairs) - nrow(pos_kmer_pairs))), 1:G)

  hetmer_rates <- calc_rates(true_af$pos, c(pos_hetmers$pos, rep(1e9, times = nrow(hetmers) - nrow(pos_hetmers))), 1:G)

  disco_rates <- calc_rates(true_af$pos, disco$pos, 1:G)

  # save all results
  result <- data.frame(
    ID = gsub("_p1|_p2", "", i),
    ID2 = i,
    true_div = sum(diversity(true_af$snp_maf)) / G, # average nucleotide diversity
    kmer_div = sum(diversity(kmer_pairs$kmer_maf)) / G,
    kmer_vs_truth_maf_cor = kmer_vs_truth_maf_cor,
    kmer_pos = kmer_rates[1],
    kmer_neg = kmer_rates[2],
    kmer_tp = kmer_rates[3],
    kmer_fp = kmer_rates[4],
    kmer_tn = kmer_rates[5],
    kmer_fn = kmer_rates[6],
    kmer_acc = kmer_rates[7],
    kmer_tpr = kmer_rates[8],
    kmer_fpr = kmer_rates[9],
    kmer_tnr = kmer_rates[10],
    kmer_fnr = kmer_rates[11],
    hetmer_div = sum(diversity(hetmers$hetmer_maf)) / G,
    hetmer_vs_truth_maf_cor = hetmer_vs_truth_maf_cor,
    hetmer_vs_truth_bc_cor = hetmer_vs_truth_bc_cor,
    hetmer_pos = hetmer_rates[1],
    hetmer_neg = hetmer_rates[2],
    hetmer_tp = hetmer_rates[3],
    hetmer_fp = hetmer_rates[4],
    hetmer_tn = hetmer_rates[5],
    hetmer_fn = hetmer_rates[6],
    hetmer_acc = hetmer_rates[7],
    hetmer_tpr = hetmer_rates[8],
    hetmer_fpr = hetmer_rates[9],
    hetmer_tnr = hetmer_rates[10],
    hetmer_fnr = hetmer_rates[11],
    disco_div = sum(diversity(disco$disco_maf)) / G,
    disco_vs_truth_maf_cor = disco_vs_truth_maf_cor,
    disco_pos = disco_rates[1],
    disco_neg = disco_rates[2],
    disco_tp = disco_rates[3],
    disco_fp = disco_rates[4],
    disco_tn = disco_rates[5],
    disco_fn = disco_rates[6],
    disco_acc = disco_rates[7],
    disco_tpr = disco_rates[8],
    disco_fpr = disco_rates[9],
    disco_tnr = disco_rates[10],
    disco_fnr = disco_rates[11],
    varscan_div = sum(diversity(varscan$snp_maf_varscan)) / G,
    varscan_vs_truth_maf_cor = varscan_vs_truth_maf_cor,
    varscan_pos = varscan_rates[1],
    varscan_neg = varscan_rates[2],
    varscan_tp = varscan_rates[3],
    varscan_fp = varscan_rates[4],
    varscan_tn = varscan_rates[5],
    varscan_fn = varscan_rates[6],
    varscan_acc = varscan_rates[7],
    varscan_tpr = varscan_rates[8],
    varscan_fpr = varscan_rates[9],
    varscan_tnr = varscan_rates[10],
    varscan_fnr = varscan_rates[11],
    poolsnp_div = sum(diversity(poolsnp$snp_maf_poolsnp)) / G,
    poolsnp_vs_truth_maf_cor = poolsnp_vs_truth_maf_cor,
    poolsnp_pos = poolsnp_rates[1],
    poolsnp_neg = poolsnp_rates[2],
    poolsnp_tp = poolsnp_rates[3],
    poolsnp_fp = poolsnp_rates[4],
    poolsnp_tn = poolsnp_rates[5],
    poolsnp_fn = poolsnp_rates[6],
    poolsnp_acc = poolsnp_rates[7],
    poolsnp_tpr = poolsnp_rates[8],
    poolsnp_fpr = poolsnp_rates[9],
    poolsnp_tnr = poolsnp_rates[10],
    poolsnp_fnr = poolsnp_rates[11],
    freebayes_div = sum(diversity(freebayes$snp_maf_freebayes)) / G,
    freebayes_vs_truth_maf_cor = freebayes_vs_truth_maf_cor,
    freebayes_pos = freebayes_rates[1],
    freebayes_neg = freebayes_rates[2],
    freebayes_tp = freebayes_rates[3],
    freebayes_fp = freebayes_rates[4],
    freebayes_tn = freebayes_rates[5],
    freebayes_fn = freebayes_rates[6],
    freebayes_acc = freebayes_rates[7],
    freebayes_tpr = freebayes_rates[8],
    freebayes_fpr = freebayes_rates[9],
    freebayes_tnr = freebayes_rates[10],
    freebayes_fnr = freebayes_rates[11]
  )

  results <- rbind(results, result)

  rm(compare_maf_kmer, compare_maf_varscan, compare_maf_poolsnp, poolsnp, poolsnp_rates, compare_maf_disco, disco, disco_rates, kmer_covs, kmer_pairs, pos_kmer_key, pos_kmer_pairs, pos_kmer_pairs_rc, true_af, varscan, kmer_rates, kmer_vs_truth_maf_cor, poolsnp_vs_truth_maf_cor, varscan_rates, varscan_vs_truth_maf_cor, disco_vs_truth_maf_cor, result, hetmer_rates, hetmers, hetmers_seqs, hetmers_counts, hetmer_vs_truth_maf_cor, hetmers_bayes)
}

pares <- merge(params, results, by = "ID")

write.table(pares, paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/results/", today, "/performance_metrics.txt", sep = ""), sep = " ")
```

# site frequency spectrum
```{r}
# compare accuracy of allele frequency estimates
plotdata_af_cor <- melt(pares[, c("ID", "cov", "n", "kmer_vs_truth_maf_cor", "hetmer_vs_truth_maf_cor", "disco_vs_truth_maf_cor", "varscan_vs_truth_maf_cor", "poolsnp_vs_truth_maf_cor")], id = c("ID", "cov", "n"))

plotdata_af_cor$variable <- gsub("_vs_truth_maf_cor", "", plotdata_af_cor$variable)

plotdata_af_cor$variable[plotdata_af_cor$variable == "kmer"] <- "smudgeplot"

ggplot(plotdata_af_cor, aes(x = as.factor(cov), color = variable, y = value)) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Coverage", y = "Cor. bw true and estimated maf", color = "method")

ggsave(paste("../results/", today, "/maf_cor.png"))

# compare diversity estimates
# melt(pares[,c("ID", "cov", "n", "true_div", "kmer_div", "disco_div", "varscan_div", "poolsnp_div")], id = c("ID", "cov", "n"))

# percent error on diversity
# percent_error = function(truth, measurement){
#   (measurement - truth)/truth * 100
# }
#
# pares$pe_div_kmer = percent_error(pares$true_div, pares$kmer_div)
#
# pares$pe_div_hetmer = percent_error(pares$true_div, pares$hetmer_div)
#
# pares$pe_div_disco = percent_error(pares$true_div, pares$disco_div)
#
# pares$pe_div_varscan = percent_error(pares$true_div, pares$varscan_div)
#
# pares$pe_div_poolsnp = percent_error(pares$true_div, pares$poolsnp_div)

plotdata_div <- melt(pares[, c("ID", "cov", "n", "true_div", "kmer_div", "hetmer_div", "disco_div", "varscan_div", "poolsnp_div")], id = c("ID", "cov", "n", "true_div"))

plotdata_div$variable <- gsub("_div", "", plotdata_div$variable)

plotdata_div$variable[plotdata_div$variable == "kmer"] <- "smudgeplot"

plotdata_div <- plotdata_div[(plotdata_div$variable %in% c("hetmer", "varscan", "disco")), ]

# ggplot(plotdata_div, aes(x = as.factor(cov), y = value, color = as.factor(variable))) +
#   geom_boxplot() +
#   theme_classic()

ggplot(plotdata_div, aes(x = log10(true_div), y = log10(value), color = as.factor(cov), shape = as.factor(variable))) +
  geom_point(size = 2) +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  theme(
    text = element_text(size = 16, family = "Helvetica", color = "black"),
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(color = "black")
  ) +
  scale_colour_scico_d(palette = "hawaii", name = "Coverage") +
  labs(y = "Estimated diversity", x = "True diversity", color = "Coverage", shape = "Method")

ggsave(paste("../../results/", today, "/estimated_vs_true_diversity.pdf", sep = ""), height = 6, width = 7)

# ggplot(pares, aes(x = log10(true_div), y = log10(hetmer_div), color = as.factor(cov))) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic()
#
# ggplot(pares, aes(x = log10(true_div), y = log10(disco_div), color = as.factor(cov))) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic()
#
# ggplot(pares, aes(x = log10(true_div), y = log10(varscan_div), color = as.factor(cov))) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic()
#
# ggplot(pares, aes(x = log10(true_div), y = log10(poolsnp_div), color = as.factor(cov))) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic()
#
# # site-frequency spectrum
# all_sfs = merge(compare_maf_kmer[,c("pos", "snp_maf", "kmer_maf")], varscan[,c("pos", "snp_maf_varscan")], by = "pos", all.x = T, all.y = T)
#
# all_sfs = merge(all_sfs, poolsnp[,c("pos", "snp_maf_poolsnp")], by = "pos", all.x = T, all.y = T)
#
# all_sfs = merge(all_sfs, disco[,c("pos", "disco_maf")], by = "pos", all.x = T, all.y = T)
#
# plotdata = melt(all_sfs, id = "pos")
#
# ggplot(plotdata, aes(x = value, fill = variable)) +
#   geom_density(alpha = 0.5) +
#   theme_classic() +
#   labs(x = "Minor allele frequency", y = "density")
#
# ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","kmer_vs_snp_sfs.png", sep = ""), height= 7, width = 8)
```

# check if theoretical distributions fit kmer counts
```{r}
n <- 200
k <- 31
R <- 301
G <- 1e6
B <- 130.641224e6

C <- B / (n * G)

Ck <- ((R - k + 1) / R) * C
nCk <- n * Ck
nC <- n * C

high <- qnorm(0.995, mean = nCk, sd = sqrt(nCk))
low <- qnorm(0.005, mean = nCk, sd = sqrt(nCk))

ggplot(kmer_covs, aes(x = c1 + c2)) +
  geom_density() +
  geom_vline(xintercept = low, lty = 2) +
  geom_vline(xintercept = nCk) +
  geom_vline(xintercept = high, lty = 2)

high <- qnorm(0.995, mean = nC, sd = sqrt(nC))
low <- qnorm(0.005, mean = nC, sd = sqrt(nC))

ggplot(kmer_covs, aes(x = c1 + c2)) +
  geom_density() +
  geom_vline(xintercept = low, lty = 2) +
  geom_vline(xintercept = nC) +
  geom_vline(xintercept = high, lty = 2)


thresh <- 4

filter <- (kmer_covs$c1 + kmer_covs$c2 > low) & (kmer_covs$c1 + kmer_covs$c2 < high) & (kmer_covs$c1 >= thresh) & (kmer_covs$c2 >= thresh)

filter <- (kmer_covs$c1 >= thresh) & (kmer_covs$c2 >= thresh)

kmer_covs[filter, ]


ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point()

ggplot(kmer_covs[(kmer_covs$c1 + kmer_covs$c2 > low) & (kmer_covs$c1 + kmer_covs$c2 < high), ], aes(x = c1 / (c1 + c2))) +
  geom_density()

ggplot(compare_maf_kmer, aes(x = kmer_maf)) +
  geom_histogram()

ggplot(true_af, aes(x = snp_maf)) +
  geom_histogram()

kmer_covs
```

# selective sweeps
```{r}
# Hypothesis: the average distance between SNPs is higher after selective sweep compared to neutral scenario
allele_file_names <- list.files(path = "/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_results", pattern = "allele_freqs_.*", full.names = T)

allele_file_names <- allele_file_names[!grepl("_p.", allele_file_names)]

allele_file_ids <- as.numeric(gsub(".txt", "", gsub("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_results/allele_freqs_", "", allele_file_names)))

allele_files <- lapply(allele_file_names, read.table, col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"))

mean_dist_bw_mut <- function(x) {
  mean(diff(x$pos))
}

max_dist_bw_mut <- function(x) {
  max(diff(x$pos))
}

# remove invariant sites
invar_check <- function(x) {
  x[((x$af < 1) & (x$af > 0)), ]
}

allele_files <- lapply(allele_files, invar_check)

# length of invariant site run at sweep site
dist_at_sweep_site <- function(x, L) {
  # find where distance to sweep site switches from positive to negative
  top_idx <- which(diff(sign(x$pos - L / 2)) != 0)

  # calculate difference between two positions
  result <- x$pos[(top_idx + 1)] - x$pos[(top_idx)]

  return(result)
}

plotdata <- data.frame(
  ID = allele_file_ids,
  mean_dist = unlist(lapply(allele_files, mean_dist_bw_mut)),
  max_dist = unlist(lapply(allele_files, max_dist_bw_mut)),
  # sweep_site_dist = unlist(lapply(allele_files, dist_at_sweep_site, L = 2e6)),
  S = unlist(lapply(allele_files, nrow))
)

plotdata <- merge(params, plotdata, by = "ID")

# expected mean unitig length
plotdata$es <- mapply(ES, plotdata$N, plotdata$mu, plotdata$L, plotdata$n)

plotdata$Ubar <- plotdata$L / plotdata$es

# Does the expected mean unitig length match reality for neutral simulations
ggplot(plotdata[(plotdata$simtype == "onepop"), ], aes(x = es, y = S)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic()

plot_mean_dist_neutral <- ggplot(plotdata[(plotdata$simtype == "onepop"), ], aes(x = Ubar, y = mean_dist, color = as.factor(mu))) +
  geom_point(aes(shape = as.factor(n))) +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "Expected mean distance between mutations", y = "Observed mean distance between mutations")

#
plotdata$n <- as.factor(plotdata$n)
plotdata$cov <- as.factor(plotdata$cov)
plotdata$s <- as.factor(plotdata$s)

# expected mean unitig length
# plotdata$Ubar = plotdata$L/plotdata$S
#
# # parameter for geometric distribution
# plotdata$p = 1/plotdata$Ubar
#
# # 95 percentile for geometric distribution
# plotdata$geom_prob = pgeom(plotdata$max_dist, plotdata$p, lower.tail = F, log.p = T)

# plotdata$simtype[(plotdata$simtype == "onepop")] = "neutral"

# how do selective sweeps change distances between snps?
ggplot(plotdata, aes(x = n, y = mean_dist, color = as.factor(Nes))) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Pool size", y = "Avg. distance between SNPs (bp)", color = "Selection", title = "Ground truth")

plot_max_dist_sweep <- ggplot(plotdata, aes(x = n, y = max_dist, color = as.factor(Nes))) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Pool size", y = "Max distance between SNPs (bp)", color = "Selection", title = "Ground truth")

# ggplot(plotdata, aes(x = n, y = sweep_site_dist/mean_dist, color = s)) +
#   geom_boxplot() +
#   theme_classic() +
#   geom_hline(yintercept = 1) +
#   labs(x = "Pool size", y = "Distance between snps at sweep/mean distance", color = "Selection", title = "Ground truth")

# Hypothesis: after a selective sweep, the longest unitigs in a population graph are more likely to be near the sweep site, compared to neutral scenario
unitig_files <- list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/blast_results", pattern = ".*txt", full.names = T)

# unitig_files = sample(unitig_files, size = 1000)

unitig_files_ids <- as.numeric(gsub("_0.txt", "", gsub("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/blast_results/", "", unitig_files)))

unitig_files <- lapply(unitig_files, read.table, col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

# unitig_files = lapply(unitig_files, fread, col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

max_len <- function(x) {
  max(x$alen)
}

mean_len <- function(x) {
  mean(x$alen)
}

ss_max_len <- function(x) {
  x$ss[which.max(x$alen)]
}

se_max_len <- function(x) {
  x$se[which.max(x$alen)]
}

plotdata_unitigs <- data.frame(
  ID = unitig_files_ids,
  max_len = unlist(lapply(unitig_files, max_len)),
  mean_len = unlist(lapply(unitig_files, mean_len)),
  ss = unlist(lapply(unitig_files, ss_max_len)),
  se = unlist(lapply(unitig_files, se_max_len))
)

plotdata_unitigs <- merge(params, plotdata_unitigs, by = "ID")

# expected mean unitig length
plotdata_unitigs$es <- mapply(ES, plotdata_unitigs$N, plotdata_unitigs$mu, plotdata_unitigs$L, plotdata_unitigs$n)

plotdata_unitigs$Ubar <- plotdata_unitigs$L / plotdata_unitigs$es

ggplot(plotdata_unitigs[(plotdata_unitigs$simtype == "onepop"), ], aes(x = Ubar, y = mean_len)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "Expected mean distance between mutations", y = "Observed mean unitig lengths")

# mod = lm(log10(max_len) ~ n + cov + sequencer + s + h, data = plotdata_unitigs)
# summary(mod)
# plot(mod)

plotdata_unitigs$n <- as.factor(plotdata_unitigs$n)
plotdata_unitigs$cov <- as.factor(plotdata_unitigs$cov)
plotdata_unitigs$s <- as.factor(plotdata_unitigs$s)

# plotdata_unitigs$simtype[(plotdata_unitigs$simtype == "onepop")] = "neutral"

# plotdata = plotdata[( ((plotdata$s == 0.5) & (plotdata$h == 1)) | (plotdata$s == 0) ),]

ggplot(plotdata_unitigs, aes(x = n, y = max_len, color = simtype, fill = cov)) +
  geom_boxplot() +
  # facet_grid(sequencer ~ .) +
  theme_classic() +
  labs(x = "Pool size", y = "Maximum unitig length (bp)", color = "Simulation\ntype", fill = "Coverage")
# stat_compare_means(aes(group = simtype), label.sep = "\n")

# see if long unitigs appear toward the sweep region
plotdata_unitigs$sc <- (plotdata_unitigs$ss + plotdata_unitigs$se) / 2

plot_unitigs_all_sims <- ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  geom_vline(xintercept = 2e6 / 2, lty = 2, lwd = 1.) +
  labs(x = "Center of longest unitig", color = "Nes", title = "All")

plot_unitigs_cov_breakdown <- ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  facet_grid(cov ~ .) +
  geom_vline(xintercept = 2e6 / 2, lty = 2, lwd = 1.) +
  labs(x = "Center of longest unitig", color = "Nes", title = "Coverage")

plot_unitigs_n_breakdown <- ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  facet_grid(n ~ .) +
  geom_vline(xintercept = 2e6 / 2, lty = 2, lwd = 1) +
  labs(x = "Center of longest unitig", color = "Nes", title = "Pool size")

plot_unitigs_shape_breakdown <- ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density(lwd = 1.25) +
  theme_minimal() +
  theme(
    text = element_text(size = 16, family = "Helvetica", color = "black"),
    axis.text.x = element_blank(),
    axis.text.y = element_text(color = "black")
  ) +
  facet_grid(shuffle ~ shape) +
  scale_color_scico_d(palette = "buda") +
  # facet_grid( ~ shape) +
  geom_vline(xintercept = 2e6 / 2, lty = 2, lwd = 1) +
  labs(x = "Center of longest unitig", color = "Nes")

ggsave(paste("../../results/", today, "/unitig_mapping_shape_breakdown.pdf", sep = ""), plot_unitigs_shape_breakdown, dpi = 400, height = 7, width = 11, unit = "in")

plot_unitigs_h_breakdown <- ggplot(plotdata_unitigs, aes(x = sc, color = as.factor(Nes))) +
  geom_density() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  facet_grid(shape + h ~ .) +
  geom_vline(xintercept = 2e6 / 2, lty = 2, lwd = 1) +
  labs(x = "Center of longest unitig", color = "Nes", title = "Shape parameter breakdown")

ggarrange(plot_unitigs_all_sims, plot_unitigs_cov_breakdown, plot_unitigs_n_breakdown, nrow = 1, ncol = 3, labels = c("A", "B", "C"), common.legend = T, widths = c(1.5, 1, 1))

ggsave(paste("../results/", today, "/unitig_mapping.jpg", sep = ""), dpi = 400, height = 7, width = 9, unit = "in")


check_overlap <- function(x, nhits, target, tol) {
  # sort by unitig length
  x_sorted <- x[order(-x$alen), ]

  # get top n longest unitigs
  x_top <- x_sorted[1:nhits, ]

  x_top$start <- pmin(x_top$ss, x_top$se)
  x_top$end <- pmax(x_top$ss, x_top$se)

  targ_start <- target - tol
  targ_end <- target + tol

  result <- any((targ_start < x_top$start) & (x_top$start < targ_end)) | any((targ_start < x_top$end) & (x_top$end < targ_end))

  return(result)
}

plotdata <- data.frame(
  ID = unitig_files_ids,
  overlap = unlist(lapply(unitig_files, check_overlap, nhits = 10, target = 500e3, tol = 50e3))
)

plotdata <- merge(params, plotdata, by = "ID")
plotdata$n <- as.factor(plotdata$n)
plotdata$cov <- as.factor(plotdata$cov)

aggregate(overlap ~ s, data = plotdata, FUN = mean)

ggplot(plotdata, aes(x = overlap, fill = simtype)) +
  geom_bar(stat = "count", position = "dodge") +
  facet_grid(s ~ .) +
  theme_classic()


allele <- read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_277.txt", col.names = c("chrom", "pos", "ns", "af", "ac"))

mean(diff(allele$pos))

unitigs <- read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/unitig_alignments_277.txt", col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

ggplot(unitigs, aes(x = alen)) +
  geom_density()

mean(unitigs$alen)

allele <- read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_5558.txt", col.names = c("chrom", "pos", "ns", "af", "ac"))

mean(diff(allele$pos))

unitigs <- read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/unitig_alignments_5558.txt", col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

ggplot(unitigs, aes(x = alen)) +
  geom_density()

mean(unitigs$alen)

sam_pos <- seq(from = 1000, to = 1e6, by = 1000)

results <- NULL
i <- 0
while (i <= 1e6) {
  result <- mean(unitigs[((unitigs$ss <= i) & (i + 1000 <= unitigs$se)), "alen"])
  results <- rbind(results, data.frame(start = i, end = i + 1000, avglen = result))
  i <- i + 1000
}

ggplot(results, aes(x = (start + end) / 2, y = avglen)) +
  geom_point() +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = 5e5) +
  theme_classic()

results <- c()
for (pos in sam_pos) {
  result <- mean(unitigs[((unitigs$ss <= pos) & (pos <= unitigs$se)), "alen"])
  results <- c(results, result)
}

plotdata <- data.frame(
  pos = sam_pos,
  avglen = results
)

ggplot(plotdata, aes(x = pos, y = avglen)) +
  geom_point() +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = 5e5) +
  theme_classic()

ggplot(unitigs, aes(x = log10(ss), y = alen)) +
  geom_point() +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = log10(5e5)) +
  theme_classic()
```

# process two population metrics (Fst)
```{r}
simids <- params$ID[(params$simtype == "twopop")]

fst <- function(n1, c1, f1, n2, c2, f2, type) {
  if (type == "sample") {
    pi_wn <- (1 / 2) * ((n1 / (n1 - 1)) * (c1 / (c1 - 1)) * (1 - (f1^2 + (1 - f1)^2)) + (n2 / (n2 - 1)) * (c2 / (c2 - 1)) * (1 - (f2^2 + (1 - f2)^2)))
  }

  if (type == "population") {
    pi_wn <- (1 / 2) * ((1 - (f1^2 + (1 - f1)^2)) + (1 - (f2^2 + (1 - f2)^2)))
  }

  pi_bw <- 1 - (f1 * f2 + (1 - f1) * (1 - f2))

  pi_total <- (pi_wn + pi_bw) / 2

  nei <- 1 - pi_wn / pi_total

  hudson <- 1 - pi_wn / pi_bw

  return(data.frame(
    pi_wn = pi_wn,
    pi_bw = pi_bw,
    pi_total = pi_total,
    nei = nei,
    hudson = hudson
  ))
}

G <- 2e6

k <- 31

thresh <- 5

# for(i in simids[1:5]){
for (i in 137:160) {
  # load in actual allele freqs from simulations
  true_af_p1 <- read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_", i, "_p1.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"))

  true_af_p2 <- read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_", i, "_p2.txt", sep = ""), col.names = c("chrom", "pos", "ref", "alt", "ns", "af", "ac"))

  true_af_p1$ac <- as.numeric(true_af_p1$ac)
  true_af_p2$ac <- as.numeric(true_af_p2$ac)

  # calculate minor allele frequencies
  true_af_p1$af <- as.numeric(true_af_p1$af)
  true_af_p2$af <- as.numeric(true_af_p2$af)

  true_af_p1$snp_maf <- pmin(true_af_p1$af, 1 - true_af_p1$af)
  true_af_p2$snp_maf <- pmin(true_af_p2$af, 1 - true_af_p2$af)

  true_af_p1 <- true_af_p1[(true_af_p1$af < 1) & (true_af_p1$af > 0), ]
  true_af_p2 <- true_af_p2[(true_af_p2$af < 1) & (true_af_p2$af > 0), ]

  true_af_p1 <- true_af_p1[complete.cases(true_af_p1), ]
  true_af_p2 <- true_af_p2[complete.cases(true_af_p2), ]

  true_af <- merge(true_af_p1[, c("pos", "ns", "ac", "af")], true_af_p2[, c("pos", "ns", "ac", "af")], by = "pos")

  fst_values <- fst(2 * true_af$ns.x, true_af$ac.x, true_af$af.x, 2 * true_af$ns.y, true_af$ac.y, true_af$af.y, "population")

  # fst(true_af$ns.x, true_af$ac.x, true_af$af.x, true_af$ns.y, true_af$ac.y, true_af$af.y, "sample")

  true_af <- cbind(true_af, fst_values)

  # load pairs of k-mers
  kmer_pairs_p1 <- read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p1_seqs.csv", sep = ""), col.names = c("k1", "k2"))

  kmer_covs_p1 <- read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p1_counts.csv", sep = ""), col.names = c("c1", "c2"))

  kmer_hashes_p1 <- read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p1_hashes.csv", sep = ""), col.names = c("h"))

  kmer_pairs_p1 <- cbind(kmer_pairs_p1, kmer_covs_p1, kmer_hashes_p1)

  kmer_pairs_p2 <- read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p2_seqs.csv", sep = ""), col.names = c("k1", "k2"))

  kmer_covs_p2 <- read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p2_counts.csv", sep = ""), col.names = c("c1", "c2"))

  kmer_hashes_p2 <- read.csv(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/hetmers_", i, "_p2_hashes.csv", sep = ""), col.names = c("h"))

  kmer_pairs_p2 <- cbind(kmer_pairs_p2, kmer_covs_p2, kmer_hashes_p2)

  # calculate minor allele frequency
  kmer_pairs_p1$kmer_af <- kmer_pairs_p1$c1 / (kmer_pairs_p1$c1 + kmer_pairs_p1$c2)

  kmer_pairs_p2$kmer_af <- kmer_pairs_p2$c1 / (kmer_pairs_p2$c1 + kmer_pairs_p2$c2)

  kmer_pairs_p1p2 <- merge(kmer_pairs_p1, kmer_pairs_p2, by = "h")

  # map kmers to snps
  pos_kmer_key <- read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/center_kmer_pairs_", i, "_p1p2.txt", sep = ""), fill = T, col.names = c("pos", "k1", "k2", "k3", "k4"))

  pos_kmer_key$leftk <- substr(pos_kmer_key$k1, 1, 15)
  pos_kmer_key$rightk <- substr(pos_kmer_key$k1, 17, 31)

  # compare snp and k-mer mafs
  kmer_pairs_p1p2$leftk <- substr(kmer_pairs_p1p2$k1.x, 1, 15)
  kmer_pairs_p1p2$rightk <- substr(kmer_pairs_p1p2$k1.x, 17, 31)

  pos_kmer_pairs <- merge(pos_kmer_key, kmer_pairs_p1p2, by = c("leftk", "rightk"))

  # get snps that pair with reverse complements
  pos_kmer_key$leftk_rc <- chartr("ATGC", "TACG", pos_kmer_key$leftk)
  pos_kmer_key$rightk_rc <- chartr("ATGC", "TACG", pos_kmer_key$rightk)

  pos_kmer_key$leftk_rc <- strReverse(pos_kmer_key$leftk_rc)
  pos_kmer_key$rightk_rc <- strReverse(pos_kmer_key$rightk_rc)

  pos_kmer_pairs_rc <- merge(pos_kmer_key, kmer_pairs_p1p2, by.x = c("leftk_rc", "rightk_rc"), by.y = c("rightk", "leftk"))

  pos_kmer_pairs_rc <- subset(pos_kmer_pairs_rc, select = -c(leftk, rightk))

  # merge reverse complement and regular sets
  names(pos_kmer_pairs_rc)[1:2] <- c("leftk", "rightk")

  pos_kmer_pairs <- rbind(pos_kmer_pairs, pos_kmer_pairs_rc)

  # do some filtering
  # remove low coverage k-mer pairs
  # remove tri-allelic k-mers
  # pos_kmer_pairs = pos_kmer_pairs[(pos_kmer_pairs$c1 >=4 & pos_kmer_pairs$c2 >= 4),]
  pos_kmer_pairs <- pos_kmer_pairs[((pos_kmer_pairs$k3 == "") | is.na(pos_kmer_pairs$k3)), ]

  # calculate bayesian allele counts
  sample_size <- params[params$ID == i, "n"]

  coverage <- params[params$ID == i, "cov"]

  sequencer <- params[params$ID == i, "sequencer"]

  if (sequencer %in% c("miseq", "nextseq")) {
    L <- 300
  }

  if (sequencer == "novaseq") {
    L <- 150
  }

  if (sequencer == "hiseq") {
    L <- 125
  }

  mean_kmer_coverage_per_genome <- ((L - k + 1) / L) * (coverage / sample_size)

  kmer_bayes_ac_p1 <- mapply(posterior_min_kmer_count, pos_kmer_pairs$c1.x, (pos_kmer_pairs$c1.x + pos_kmer_pairs$c1.y), MoreArgs = list(n = sample_size, C = mean_kmer_coverage_per_genome, c = thresh, alpha = 1, beta = 1))

  kmer_bayes_ac_p2 <- mapply(posterior_min_kmer_count, pos_kmer_pairs$c1.y, (pos_kmer_pairs$c1.y + pos_kmer_pairs$c1.y), MoreArgs = list(n = sample_size, C = mean_kmer_coverage_per_genome, c = thresh, alpha = 1, beta = 1))

  # calculate fst for kmers using empirical frequencies
  kmer_fst_values <- fst(2 * sample_size, pos_kmer_pairs$c1.x, pos_kmer_pairs$kmer_af.x, 2 * sample_size, pos_kmer_pairs$c1.y, pos_kmer_pairs$kmer_af.y, "sample")

  pos_kmer_pairs <- cbind(pos_kmer_pairs, kmer_fst_values, kmer_bayes_ac_p1, kmer_bayes_ac_p2)

  # add in SNP fst values
  snp_vs_kmer_fst <- merge(true_af[, c("pos", "nei", "hudson")], pos_kmer_pairs[, c("pos", "nei", "hudson")], by = "pos")

  print(ggplot(snp_vs_kmer_fst, aes(x = nei.x, y = nei.y)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    theme_classic() +
    labs(x = "true SNP Nei Fst", y = "Nei hetmer Fst", title = i))

  print(ggplot(snp_vs_kmer_fst, aes(x = hudson.x, y = hudson.y)) +
    geom_point() +
    theme_classic() +
    geom_abline(slope = 1, intercept = 0) +
    labs(x = "true SNP Hudson Fst", y = "Hudson hetmer Fst", title = i))

  # mean(snp_vs_kmer_fst$nei.x)
  # mean(snp_vs_kmer_fst$nei.y)
  #
  # mean(snp_vs_kmer_fst$hudson.x)
  # mean(snp_vs_kmer_fst$hudson.y)
  #
  # mean(true_af$hudson)
  # mean(kmer_pairs_p1p2$hudson[(kmer_pairs_p1p2$hudson > 0)])
}
```

# Allele frequencies of known variants

## Simulations
```{r}
results <- NULL

sims_to_plot <- c(1, 4, 7, 10, 13, 25, 37, 49)

for (i in params$ID) {
  print(i)

  kmer_counts <- tryCatch(
    {
      read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freqk_results/", i, "_0_freqs.txt", sep = ""), sep = "|")
    },
    error = function(e) {
      message(paste("Error at iteration", i, ":", e$message))
      return(NA) # Or some other indicator for error
    }
  )

  if (all(is.na(kmer_counts))) {
    next
  }

  kmer_counts$kmer_maf <- pmin(kmer_counts$V1, kmer_counts$V2) / (kmer_counts$V1 + kmer_counts$V2)

  kmer_index <- tryCatch(
    {
      read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freqk_ref_dedup/", i, "_0.txt", sep = ""), sep = ",")
    },
    error = function(e) {
      message(paste("Error at iteration", i, ":", e$message))
      return(NA) # Or some other indicator for error
    }
  )

  slim <- tryCatch(
    {
      read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/slim_results/allele_freqs_", i, "_0.txt", sep = ""), sep = " ")
    },
    error = function(e) {
      message(paste("Error at iteration", i, ":", e$message))
      return(NA) # Or some other indicator for error
    }
  )

  if (all(is.na(slim))) {
    next
  }

  names(slim)[1:2] <- c("Chrom", "Position")

  slim$slim_maf <- pmin(slim$V6, 1 - slim$V6)

  kmer_counts <- cbind(kmer_index, kmer_counts)

  names(kmer_counts)[2:3] <- c("Chrom", "Position")

  # shift position to vcf convention
  kmer_counts$Position <- kmer_counts$Position + 2

  slim_vs_kmer <- merge(slim, kmer_counts, by = c("Chrom", "Position"))

  cor_coeff_freqk <- signif(cor.test(slim_vs_kmer$kmer_maf, slim_vs_kmer$slim_maf)$estimate, digits = 3)

  # print(cor_coeff)

  # varscan = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/varscan_results/", i, "_0.tsv", sep = ""), header = T)

  # varscan$maf = pmin(varscan$Reads1, varscan$Reads2)/(varscan$Reads1 + varscan$Reads2)

  # freebayes = tryCatch({
  #   read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freebayes_a_priori_results/", i, "_0.txt", sep = ""), header = F)
  # }, error = function(e) {
  #   message(paste("Error at iteration", i, ":", e$message))
  #   return(NA) # Or some other indicator for error
  # })
  #
  # if (all(is.na(freebayes))){
  #   next
  # }
  #
  # names(freebayes)[c(1:2,6)] = c("Chrom", "Position", "af")
  #
  # freebayes$maf = pmin(freebayes$af, 1 - freebayes$af)
  #
  # slim_freebayes = merge(slim, freebayes, by = c("Chrom", "Position"))
  #
  # cor_coeff_freebayes = signif(cor.test(slim_freebayes$maf.x, slim_freebayes$maf.y)$estimate, digits = 3)

  ## vg giraffe + freebayes
  vg <- tryCatch(
    {
      read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/freebayes_vg_results/", i, "_0.txt", sep = ""), header = F)
    },
    error = function(e) {
      message(paste("Error at iteration", i, ":", e$message))
      return(NA) # Or some other indicator for error
    }
  )

  if (all(is.na(vg))) {
    next
  }

  # load benchmarks

  names(vg)[c(1:2, 6)] <- c("Chrom", "Position", "af")

  vg$vg_maf <- pmin(vg$af, 1 - vg$af)

  slim_vg <- merge(slim, vg, by = c("Chrom", "Position"))

  cor_coeff_vg <- signif(cor.test(slim_vg$slim_maf, slim_vg$vg_maf)$estimate, digits = 3)

  ## bwa mem + freebayes
  # bench_bwa_mem = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/bwa_mem/", i, "_0.bench", sep = ""), header = T)

  # bench_freebayes_a_priori = read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freebayes_a_priori/", i, "_0.bench", sep = ""), header = T)

  ## vg giraffe + freebayes
  bench_vg_index <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_autoindex/", i, "_0.bench", sep = ""), header = T)

  bench_vg_giraffe <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_giraffe/", i, "_0.bench", sep = ""), header = T)

  bench_vg_surject_paired <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_surject/paired_", i, "_0.bam", sep = ""), header = T)

  bench_vg_surject_r1 <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_surject/R1_", i, "_0.bam", sep = ""), header = T)

  bench_vg_surject_r2 <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/vg_surject/R2_", i, "_0.bam", sep = ""), header = T)

  bench_vg_sort_paired <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/samtools_sort/paired_", i, "_0.bam", sep = ""), header = T)

  bench_vg_sort_r1 <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/samtools_sort/R1_", i, "_0.bam", sep = ""), header = T)

  bench_vg_sort_r2 <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/samtools_sort/R2_", i, "_0.bam", sep = ""), header = T)

  bench_vg_merge <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/samtools_merge/", i, "_0.bam", sep = ""), header = T)

  bench_freebayes_vg <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freebayes_vg/", i, "_0.bench", sep = ""), header = T)

  ## kmer counts
  bench_freqk_index <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freqk_index/", i, "_0.bench", sep = ""), header = T)

  bench_freqk_var_dedup <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freqk_var_dedup/", i, "_0.bench", sep = ""), header = T)

  bench_freqk_ref_dedup <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freqk_ref_dedup/", i, "_0.bench", sep = ""), header = T)

  bench_freqk_count <- read.table(paste("/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/benchmarks/freqk_count/", i, "_0.bench", sep = ""), header = T)

  # combine all stats in a frame
  result <- data.frame(
    ID = i,
    cor_coeff_freqk = cor_coeff_freqk,
    # cor_coeff_freebayes = cor_coeff_freebayes,
    cor_coeff_vg = cor_coeff_vg,
    t_freqk = bench_freqk_count$s,
    mem_freqk = max(bench_freqk_count$max_rss),
    t_freqk_index = bench_freqk_index$s + bench_freqk_var_dedup$s + bench_freqk_ref_dedup$s,
    mem_freqk_index = max(bench_freqk_index$max_rss, bench_freqk_var_dedup$max_rss, bench_freqk_ref_dedup$max_rss),
    # t_freebayes = bench_bwa_mem$s + bench_freebayes_a_priori$s,
    # mem_freebayes = max(bench_bwa_mem$max_rss, bench_freebayes_a_priori$max_rss),
    t_vg = bench_vg_giraffe$s + max(c(bench_vg_surject_paired$s, bench_vg_surject_r1$s, bench_vg_surject_r2$s)) + max(c(bench_vg_sort_paired$s, bench_vg_sort_r1$s, bench_vg_sort_r2$s)) + bench_vg_merge$s + bench_freebayes_vg$s,
    mem_vg = max(bench_vg_giraffe$max_rss, bench_vg_surject_paired$max_rss, bench_vg_surject_r1$max_rss, bench_vg_surject_r2$max_rss, bench_vg_sort_paired$max_rss, bench_vg_sort_r1$max_rss, bench_vg_sort_r2$max_rss, bench_vg_merge$max_rss, bench_freebayes_vg$max_rss),
    t_vg_index = bench_vg_index$s,
    mem_vg_index = bench_vg_index$max_rss
  )

  results <- rbind(results, result)

  # make a plot of some individual simulations
  if (i %in% sims_to_plot) {
    plotdata_known_var <- data.frame(kmer_maf = kmer_counts$maf, slim_maf = slim$maf)

    ggplot(plotdata_known_var, aes(x = slim_maf, y = kmer_maf)) +
      geom_point() +
      geom_abline(slope = 1, intercept = 0, lwd = 2) +
      theme_minimal() +
      theme(text = element_text(size = 16, family = "Helvetica")) +
      labs(x = "True MAF", y = "k-mer counts MAF estiamte", title = paste("r = ", cor_coeff_freqk, sep = ""))

    ggsave(paste("../../results/", today, "/", i, ".pdf", sep = ""), width = 7, height = 7)
  }
}


# compare mapping approaches to freqk
params_res <- merge(params, results, by = "ID")

# ggplot(results, aes(x = cor_coeff_freebayes, y = cor_coeff_freqk)) +
#   geom_point(shape = 16) +
#   geom_abline(slope = 1, intercept = 0, lwd = 2) +
#   theme_minimal() +
#   theme(text = element_text(size = 16, family = "Helvetica")) +
#   labs(x = "Pearson Corr. of MAF\nbwa+Freebayes vs SLiM", y = "Pearson Corr. of MAF\nK-mers vs SLiM")

ggplot(params_res, aes(x = cor_coeff_vg, y = cor_coeff_freqk, color = as.factor(cov))) +
  geom_point(shape = 16) +
  geom_abline(slope = 1, intercept = 0, lwd = 2) +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_colour_scico_d(palette = "hawaii", name = "Coverage") +
  labs(x = "Pearson Corr.\nvg+Freebayes vs SLiM MAF", y = "Pearson Corr.\nk-mers vs SLiM MAF")

ggsave(paste("../../results/", today, "/", "vg_vs_freqk.pdf", sep = ""), height = 6, width = 7, device = cairo_pdf)

# compare run times per sample
time_res <- melt(params_res[, c("ID", "cov", "n", "t_freqk", "t_vg")], id = c("ID", "cov", "n"))

mem_res <- melt(params_res[, c("ID", "cov", "n", "mem_freqk", "mem_vg")], id = c("ID", "cov", "n"))

names(mem_res)[4] <- "tool"
names(mem_res)[5] <- "mem"
mem_res$tool <- gsub("mem_", "", mem_res$tool)

names(time_res)[4] <- "tool"
names(time_res)[5] <- "time"
time_res$tool <- gsub("t_", "", time_res$tool)

res_summary <- merge(time_res, mem_res, by = c("ID", "cov", "n", "tool"))

# compare time and memory
plot_time <- ggplot(res_summary, aes(x = tool, y = time, fill = as.factor(cov))) +
  geom_boxplot() +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_fill_scico_d(palette = "hawaii", name = "Coverage") +
  # facet_zoom(ylim = c(0,750)) +
  # stat_compare_means(method = "wilcox", paired = TRUE) +
  # ylim(0,1000) +
  labs(x = "Workflow", y = "Wall Time (s)", fill = "Coverage")

plot_mem <- ggplot(res_summary, aes(x = tool, y = mem, fill = as.factor(cov))) +
  geom_boxplot() +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_fill_scico_d(palette = "hawaii", name = "Coverage") +
  labs(x = "Workflow", y = "Maximum\nResident Set Size (MB)", fill = "Coverage")

ggarrange(plot_time, plot_mem, nrow = 1, ncol = 2, labels = c("A", "B"), font.label = list(size = 16, family = "Helvetica"), common.legend = T)

ggsave(paste("../../results/", today, "/", "resources_freqk_vs_vg.pdf", sep = ""), height = 4, width = 9)

# compare index times
time_index <- melt(params_res[, c("ID", "cov", "n", "t_freqk_index", "t_vg_index")], id = c("ID", "cov", "n"))

mem_index <- melt(params_res[, c("ID", "cov", "n", "mem_freqk_index", "mem_vg_index")], id = c("ID", "cov", "n"))

names(mem_index)[4] <- "tool"
names(mem_index)[5] <- "mem"
mem_index$tool <- gsub("mem_", "", mem_index$tool)
mem_index$tool <- gsub("_index", "", mem_index$tool)

names(time_index)[4] <- "tool"
names(time_index)[5] <- "time"
time_index$tool <- gsub("t_", "", time_index$tool)
time_index$tool <- gsub("_index", "", time_index$tool)

res_index <- merge(time_index, mem_index, by = c("ID", "cov", "n", "tool"))

plot_time_index <- ggplot(res_index, aes(x = tool, y = time, fill = as.factor(n))) +
  geom_boxplot() +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_fill_scico_d(palette = "acton", name = "Pool size") +
  labs(x = "Workflow", y = "Wall Time (s)", fill = "Pool size")

plot_mem_index <- ggplot(res_index, aes(x = tool, y = mem, fill = as.factor(n))) +
  geom_boxplot() +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  scale_fill_scico_d(palette = "acton", name = "Coverage") +
  labs(x = "Workflow", y = "Maximum\nResident Set Size (MB)", fill = "Pool size")

ggarrange(plot_time_index, plot_mem_index, nrow = 1, ncol = 2, labels = c("A", "B"), font.label = list(size = 16, family = "Helvetica"), common.legend = T)

ggsave(paste("../../results/", today, "/", "resources_index_freqk_vs_vg.pdf", sep = ""), height = 4, width = 7)
```

## Empirical data
```{r}
# real_counts = read.table("/global/scratch/users/milesroberts/tmp/arabidopsis_thaliana_plantpan_data/freqs.txt", sep = " ")
# real_counts = read.table("/global/scratch/users/milesroberts/tmp/brassica_rapa_plantpan_data/calls.txt", sep = " ")
real_counts <- read.table("/global/scratch/users/milesroberts/tmp/arabidopsis_thaliana_plantpan_data/calls.txt", sep = " ")

real_counts <- str_split(real_counts$V1, "\\|")

# real_counts = lapply(real_counts, as.numeric)

num_cols <- max(unlist(lapply(real_counts, length)))

real_col_names <- paste("f", 1:num_cols, sep = "")

# real_counts = read.table("/global/scratch/users/milesroberts/tmp/arabidopsis_thaliana_plantpan_data/freqs.txt", sep = "|", col.names =real_col_names, fill = T)
# real_counts = read.table("/global/scratch/users/milesroberts/tmp/brassica_rapa_plantpan_data/calls.txt", sep = "|", col.names =real_col_names, fill = T)
real_counts <- read.table("/global/scratch/users/milesroberts/tmp/arabidopsis_thaliana_plantpan_data/calls.txt", sep = "|", col.names = real_col_names, fill = T)

# real_index = read.table("/global/scratch/users/milesroberts/tmp/arabidopsis_thaliana_plantpan_data/at_ref_index.txt", sep = ",", fill = T)
# real_index = read.table("/global/scratch/users/milesroberts/tmp/brassica_rapa_plantpan_data/ref_index.txt", sep = ",", fill = T)
real_index <- read.table("/global/scratch/users/milesroberts/tmp/arabidopsis_thaliana_plantpan_data/ref_index.txt", sep = ",", fill = T)

real_counts <- cbind(real_index, real_counts)

real_counts$V3 <- real_counts$V3 + 2

# divide counts by number of k-mers
# num_kmers = str_split(real_counts$V7, "\\|")

# num_kmers = lapply(num_kmers, as.numeric)

# alternate allele frequency
# real_counts$af = 1 - real_counts$c1/rowSums(real_counts[,real_col_names], na.rm = T)

# load in vcf
# real_vcf <- read.vcfR("/global/scratch/users/milesroberts/tmp/arabidopsis_thaliana_plantpan_data/at.genetics.variation/final.vcf.gz")
# real_vcf <- read.vcfR("/global/scratch/users/milesroberts/tmp/brassica_rapa_plantpan_data/br.genetics.variation/final.vcf.gz")
real_vcf <- read.vcfR("/global/scratch/users/milesroberts/tmp/arabidopsis_thaliana_plantpan_data/at.genetics.variation/all_nonoverlap_norm.vcf.gz")

gt <- apply(real_vcf@gt[, -1], 2, as.numeric)

pos <- real_vcf@fix

# chrom_pos = paste(pos[,"CHROM"], pos[,"POS"], sep = " ")

# to_keep = paste(pos[,"CHROM"], pos[,"POS"], sep = " ") %in% paste(real_counts$V2, real_counts$V3, sep = " ")

# gt = gt[to_keep,]

# pos = pos[to_keep,]

# real_counts = real_counts[to_keep,]

# remove duplicates
# to_keep = !duplicated(paste(pos[,"CHROM"], pos[,"POS"], sep = " "))

# gt = gt[to_keep,]

# pos = pos[to_keep,]

allele_counts <- rowSums(gt == 0)

pos <- cbind(pos, allele_counts)

names(real_counts)[2:3] <- c("CHROM", "POS")

real_plot_data <- merge(real_counts, pos, by = c("CHROM", "POS"))

real_plot_data$allele_counts <- as.numeric(real_plot_data$allele_counts)

# round here based on number of genomes in pool
real_plot_data$estim_af <- real_plot_data$f1 * 7
real_plot_data$estim_rounded_af <- round(real_plot_data$f1 * 7)

# real_plot_data = data.frame(
#  estim_af = real_counts$f1*7,
#  estim_rounded_af = round(real_counts$f1*7),
#  true_af = allele_counts
# )

# type of variant
real_plot_data$var_type <- "foobar"
real_plot_data$var_type[(rowSums(sapply(c("SNP", "INS", "DEL"), grepl, real_plot_data$ID)) > 1)] <- "MIXED"
real_plot_data$var_type[grepl("SNP", real_plot_data$ID) & (real_plot_data$var_type != "MIXED")] <- "SNP"
real_plot_data$var_type[grepl("INS", real_plot_data$ID) & (real_plot_data$var_type != "MIXED")] <- "INS"
real_plot_data$var_type[grepl("DEL", real_plot_data$ID) & (real_plot_data$var_type != "MIXED")] <- "DEL"
real_plot_data$var_type[real_plot_data$var_type == "foobar"] <- "HDR"

table(real_plot_data$var_type)


# number of alleles
num_alleles <- str_split(real_plot_data$V5, "\\|")
num_alleles <- unlist(lapply(num_alleles, length))

real_plot_data$num_alleles <- num_alleles

# remove cases where the number of allele-specific k-mer for at least 1 allele is 0
all_greater_than_zero <- function(x) {
  all(x > 0)
}

real_plot_data <- real_plot_data[unlist(lapply(lapply(str_split(real_plot_data$V7, "\\|"), as.numeric), all_greater_than_zero)), ]

# whether estimated = true frequency
real_plot_data$match <- (real_plot_data$estim_rounded_af == real_plot_data$allele_counts)

cor_estim <- signif(cor.test(real_plot_data$estim_rounded_af, real_plot_data$allele_counts, method = "spearman")$estimate, 3)

# ggplot(real_plot_data, aes(x = as.factor(true_af), y = estim_af)) +
#   geom_boxplot() +
#   geom_point(aes(x=1, y=2), colour="red", size = 3, shape = 16) +
#   geom_point(aes(x=2, y=3), colour="red", size = 3, shape = 16) +
#   geom_point(aes(x=3, y=4), colour="red", size = 3, shape = 16) +
#   geom_point(aes(x=4, y=5), colour="red", size = 3, shape = 16) +
#   theme_minimal() +
#   theme(text = element_text(size = 16, family = "Helvetica")) +
#   labs(x = "True allele frequency", y = "Estimated allele frequency", title = paste("r = ", cor_estim, sep = ""))
#
# estim_by_counts = table(real_plot_data$true_af, real_plot_data$estim_rounded_af)
#
# estim_by_counts = melt(estim_by_counts)
#
# estim_by_counts$bar_color = ifelse(estim_by_counts$Var.1 == estim_by_counts$Var.2, "red", "grey")

# plot number of estimates for each af class
ggplot(real_plot_data, aes(x = estim_rounded_af, fill = match)) +
  geom_bar() +
  facet_wrap(var_type ~ allele_counts, scales = "free", ncol = 7) +
  scale_fill_manual(values = c("black", "grey")) +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica", colour = "black"), axis.text = element_text(color = "black"), legend.position = "None") +
  labs(x = "Estimated REF allele frequency", y = "Number of variants")

ggsave(paste("../../results/", today, "/", "true_vs_estimated_arabidopsis_thaliana_7_genomes.pdf", sep = ""), width = 9, height = 9)

ggplot(real_plot_data[(real_plot_data$num_alleles == 2), ], aes(x = estim_rounded_af, fill = match)) +
  geom_bar() +
  facet_wrap(var_type ~ allele_counts, scales = "free", ncol = 7) +
  scale_fill_manual(values = c("black", "grey")) +
  theme_minimal() +
  theme(text = element_text(size = 16, family = "Helvetica", colour = "black"), axis.text = element_text(color = "black"), legend.position = "None") +
  labs(x = "Estimated REF allele frequency", y = "Number of variants")

ggsave(paste("../../results/", today, "/", "true_vs_estimated_biallelic_arabidopsis_thaliana_7_genomes.pdf", sep = ""), height = 9, width = 9)

# vg performance
vg_results <- list.files(path = "/global/scratch/users/milesroberts/moi_lab_projects/poolseq-kmers/workflow/", full.names = T, pattern = "calls_A.*.vcf")

vg_results <- lapply(vg_results, read.vcfR)

get_allele_counts <- function(vcf) {
  gt <- vcf@gt[, -1]

  gt <- gsub(":.*", "", gt)

  print(head(gt))

  pos <- vcf@fix

  allele_counts <- str_count(gt, pattern = "0")

  pos <- cbind(pos, allele_counts)

  return(pos)
}

vg_allele_counts <- lapply(vg_results, get_allele_counts)

vg_allele_counts <- do.call("rbind", vg_allele_counts)

# vg_allele_counts[,"allele_counts"] = as.numeric(vg_allele_counts[,"allele_counts"])

# pos[,"allele_counts"] = 2*as.numeric(pos[,"allele_counts"])

vg_vs_real <- merge(pos, vg_allele_counts, by = c("CHROM", "POS"))

vg_vs_real$allele_counts.x <- 2 * as.numeric(vg_vs_real$allele_counts.x)

# vg_vs_real$allele_counts.y = round(as.numeric(vg_vs_real$allele_counts.y)/2)
vg_vs_real$allele_counts.y <- as.numeric(vg_vs_real$allele_counts.y)

cor.test(vg_vs_real$allele_counts.x, vg_vs_real$allele_counts.y, method = "spearman")

vg_vs_real$MATCH <- (vg_vs_real$allele_counts.x == vg_vs_real$allele_counts.y)

ggplot(vg_vs_real, aes(x = allele_counts.y, fill = MATCH)) +
  geom_bar() +
  theme_minimal() +
  scale_fill_manual(values = c("black", "grey")) +
  facet_wrap(~allele_counts.x, scales = "free") +
  theme(text = element_text(size = 16, family = "Helvetica", colour = "black"), axis.text = element_text(color = "black"), legend.position = "None") +
  labs(x = "Estimated REF allele count", y = "Number of variants")

ggsave(paste("../../results/", today, "/", "true_vs_vg_brassica_napus.pdf", sep = ""), height = 7, width = 9)
```



# How does performance vary with sequencing parameters?
```{r}
# compare sensitivities for SNPs represented in reference
tpr_plot_data <- melt(pares[, c("ID", "ID2", "cov", "n", "sequencer", "simtype", "kmer_tpr", "hetmer_tpr", "disco_tpr", "varscan_tpr", "poolsnp_tpr")], id = c("ID", "ID2", "cov", "n", "sequencer", "simtype"))

tpr_plot_data$variable <- factor(gsub("_tpr", "", tpr_plot_data$variable), levels = c("kmer", "hetmer", "disco", "varscan", "poolsnp"), ordered = T)

tpr_plot_data$cov_per_n <- as.factor(tpr_plot_data$cov / (tpr_plot_data$n * 2))

tpr_plot_data$cov <- as.factor(tpr_plot_data$cov)
tpr_plot_data$n <- as.factor(tpr_plot_data$n)

tpr_vs_coverage_plot <- ggplot(tpr_plot_data, aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Total coverage", y = "Sensitivity", color = "Method", title = "All instruments")

# ggsave(paste("../results/", today, "/coverage_vs_sensitivity.png", sep = ""), unit = "in")

# ggplot(tpr_plot_data, aes(x = cov_per_n, y = value, color = simtype)) +
#   geom_boxplot() +
#   theme_classic() +
#   theme(text = element_text(size = 12)) +
#   labs(x = "Coverage per haploid genome", y = "Sensitivity (true positive rate)", color = "Method", title = "All instruments")

# coverage vs false discovery rate
fpr_plot_data <- melt(pares[, c("ID", "cov", "n", "sequencer", "kmer_fpr", "hetmer_fpr", "disco_fpr", "varscan_fpr", "poolsnp_fpr")], id = c("ID", "cov", "n", "sequencer"))

fpr_plot_data$variable <- factor(gsub("_fpr", "", fpr_plot_data$variable), levels = c("kmer", "hetmer", "disco", "varscan", "poolsnp"), ordered = T)

fpr_plot_data$cov <- as.factor(fpr_plot_data$cov)
fpr_plot_data$n <- as.factor(fpr_plot_data$n)

fpr_vs_coverage_plot <- ggplot(fpr_plot_data, aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Total coverage", y = "False positive rate", color = "Method", title = "All instruments")

# ggsave(paste("../results/", today, "/coverage_vs_fpr.png", sep = ""), unit = "in")

# coverage vs false negative rate
fnr_plot_data <- melt(pares[, c("ID", "cov", "n", "sequencer", "kmer_fnr", "hetmer_fnr", "disco_fnr", "varscan_fnr", "poolsnp_fnr")], id = c("ID", "cov", "n", "sequencer"))

fnr_plot_data$variable <- factor(gsub("_fnr", "", fnr_plot_data$variable), levels = c("kmer", "hetmer", "disco", "varscan", "poolsnp"), ordered = T)

fnr_plot_data$cov <- as.factor(fnr_plot_data$cov)
fnr_plot_data$n <- as.factor(fnr_plot_data$n)

fnr_vs_coverage_plot <- ggplot(fnr_plot_data, aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Total coverage", y = "False negative rate", color = "Method", title = "All instruments")

# coverage vs specificity
tnr_plot_data <- melt(pares[, c("ID", "ID2", "cov", "n", "sequencer", "simtype", "kmer_tnr", "hetmer_tnr", "disco_tnr", "varscan_tnr", "poolsnp_tnr")], id = c("ID", "ID2", "cov", "n", "sequencer", "simtype"))

tnr_plot_data$variable <- factor(gsub("_tnr", "", tnr_plot_data$variable), levels = c("kmer", "hetmer", "disco", "varscan", "poolsnp"), ordered = T)

tnr_plot_data$cov <- as.factor(tnr_plot_data$cov)
tnr_plot_data$n <- as.factor(tnr_plot_data$n)

tnr_vs_coverage_plot <- ggplot(tnr_plot_data, aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Total coverage", y = "Specificity", color = "Method", title = "All instruments")

# plot tpr and fnr together
ggarrange(tpr_vs_coverage_plot, tnr_vs_coverage_plot, fpr_vs_coverage_plot, fnr_vs_coverage_plot, common.legend = T, font.label = list(size = 16), labels = c("A", "B", "C", "D"), nrow = 1, ncol = 4)

ggsave(paste("../results/", today, "/coverage_vs_tpr_tnr_fpr_fnr.png", sep = ""), unit = "in", width = 12, height = 4, dpi = 350)


# coverage vs false positives
fp_plot_data <- melt(pares[, c("ID", "cov", "n", "sequencer", "kmer_fp", "disco_fp", "varscan_fp", "poolsnp_fp")], id = c("ID", "cov", "n", "sequencer"))

fp_plot_data$variable <- factor(gsub("_fp", "", fp_plot_data$variable), levels = c("kmer", "disco", "varscan", "poolsnp"), ordered = T)

fp_plot_data$cov_per_n <- as.factor(fp_plot_data$cov / (fp_plot_data$n * 2))

ggplot(fp_plot_data, aes(x = cov_per_n, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Coverage per haploid genome", y = "False positives", color = "Method", title = "All instruments")

# ggsave(paste("../results/", today, "/coverage_vs_fp.png", sep = ""), unit = "in")

#


#
tpr_vs_fpr <- merge(tpr_plot_data, fpr_plot_data, by = c("ID", "cov", "n", "sequencer", "variable", "cov_per_n"))

ggplot(tpr_vs_fpr, aes(x = value.y, y = value.x, color = variable)) +
  geom_point() +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "False positive rate", y = "True positive rate", color = "Method", title = "All instruments")

ggplot(tpr_plot_data[(tpr_plot_data$n == 50 & tpr_plot_data$sequencer == "miseq"), ], aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "Coverage", y = "Sensitivity (true positive rate)", color = "Method", title = "pool size = 50; MiSeq")

ggsave(paste("../results/", today, "/sensitivity_50_miseq.png", sep = ""), height = 7, width = 8)

# hold pool size constant and increase coverage
ggplot(pares[(pares$n == 5), ], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 10), ], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 25), ], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 50), ], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 100), ], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

# hold coverage constant, and add more individuals to pool
ggplot(pares[(pares$cov == 20), ], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$cov == 50), ], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$cov == 80), ], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$cov == 100), ], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

# hold pool size constant, increase coverage
ggplot(pares[(pares$n == 5), ], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 10), ], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 25), ], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 50), ], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 100), ], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()


ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "True MAF", y = "Predicted MAF from k-mers")

ggplot(compare_maf_kmer, aes(x = snp_maf, y = snp_maf_varscan)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "True MAF", y = "Predicted MAF from VarScan")

# what about for k-mers above some frequency threshold
kmer_test <- cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$kmer_maf, method = "spearman")

ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "True MAF", y = "Predicted MAF from k-mers", title = paste("rho =", signif(kmer_test$estimate, 4), "; p =", paste(format(signif(kmer_test$p.value, 2), digits = 2))))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today, "/", "kmer_maf_vs_truth_pool-seq.png", sep = ""), height = 7, width = 7)

varscan_test <- cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$snp_maf_varscan, method = "spearman")

ggplot(compare_maf_kmer, aes(x = snp_maf, y = snp_maf_varscan)) +
  geom_density_2d_filled() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "True MAF", y = "Predicted MAF from VarScan", title = paste("rho =", signif(varscan_test$estimate, 4), "; p =", paste(format(signif(varscan_test$p.value, 2), digits = 2))))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today, "/", "varscan_maf_vs_truth_pool-seq.png", sep = ""), height = 7, width = 8)

# smudgeplot
ggplot(compare_maf_kmer, aes(x = kmer_maf, y = (c1 + c2) / 81.670932)) +
  geom_density_2d_filled() +
  # geom_hline(yintercept = 50) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "Minor k-mer frequency", y = "Total pair coverage")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today, "/", "smudgeplot.png", sep = ""), height = 7, width = 8)


# folded sfs
kmer_pairs <- kmer_pairs[(kmer_pairs$c1 >= 4 & kmer_pairs$c2 >= 4), ]

folded_sfs <- data.frame(
  category = factor(c(rep("truth", times = nrow(true_af)), rep("kmer", times = nrow(kmer_pairs)), rep("varscan", times = nrow(varscan)))),
  value = c(true_af$snp_maf, kmer_pairs$kmer_maf, varscan$snp_maf_varscan)
)

ggplot(folded_sfs, aes(x = value, fill = category)) +
  geom_density(alpha = 0.5) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "minor allele frequency")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today, "/", "kmer_vs_snp_sfs.png", sep = ""), height = 7, width = 8)

ggplot(true_af, aes(x = snp_maf)) +
  geom_density() +
  theme_classic()

ggplot(kmer_pairs, aes(x = kmer_maf)) +
  geom_density() +
  theme_classic()

ggplot(kmer_pairs, aes(x = kmer_maf)) +
  geom_density() +
  theme_classic()

ggplot(true_af[(true_af$snp_maf >= 0.05), ], aes(x = snp_maf)) +
  geom_density() +
  theme_classic()

ggplot(compare_maf_kmer, aes(x = kmer_maf)) +
  geom_histogram(binwidth = 0.05) +
  theme_classic()

ggplot(compare_maf_kmer, aes(x = snp_maf)) +
  geom_histogram(breaks = seq(from = 0, to = 0.5, by = 0.02)) +
  theme_classic()

ggplot(compare_maf_kmer, aes(x = kmer_maf)) +
  geom_histogram(breaks = seq(from = 0, to = 0.5, by = 0.02)) +
  theme_classic()

# calculate error
compare_maf_kmer$error <- abs(compare_maf_kmer$kmer_maf - compare_maf_kmer$snp_maf)

summary(compare_maf_kmer$error)

sum(compare_maf_kmer$error < 0.1) / nrow(compare_maf_kmer)

ggplot(confident_sites, aes(x = total_cov, y = error)) +
  geom_point() +
  theme_classic()

#
compare_maf_kmer$snp_maf_bins <- cut(compare_maf_kmer$snp_maf, breaks = seq(from = 0, to = 0.5, by = 0.05))
compare_maf_kmer$kmer_maf_bins <- cut(compare_maf_kmer$kmer_maf, breaks = seq(from = 0, to = 0.5, by = 0.05))

observed <- table(compare_maf_kmer$kmer_maf_bins) # Observed frequencies
expected <- table(compare_maf_kmer$snp_maf_bins) # Expected frequencies

observed <- observed[(expected != 0)]
expected <- expected[(expected != 0)]

# Calculate Chi-Square statistic manually
chi_sq_statistic <- sum((observed - expected)^2 / expected)
df <- length(observed) - 1
p_value <- 1 - pchisq(chi_sq_statistic, df)

# Print results
print(paste("Chi-Square Statistic:", chi_sq_statistic))
print(paste("Degrees of Freedom:", df))
print(paste("P-value:", p_value))
```

# DEPRECATED: correlations between true and predicted minor allele frequencies

```{r}
# load pairs of k-mers
kmer_pairs <- read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", "kmerpairs_938_sequences.tsv", sep = ""), col.names = c("k1"))

kmer_covs <- read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/kmerpairs_938_coverages.tsv", sep = ""), col.names = c("c1", "c2"))

kmer_pairs <- cbind(kmer_pairs, kmer_covs)

# calculate minor allele frequency
kmer_pairs$kmer_maf <- pmin(kmer_pairs$c1, kmer_pairs$c2) / (kmer_pairs$c1 + kmer_pairs$c2)

# load in actual allele freqs from simulations
true_af <- read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_938.txt", sep = ""), col.names = c("chrom", "pos", "af", "ac"))

true_af$ac <- as.numeric(true_af$ac)

true_af <- true_af[complete.cases(true_af), ]

# calculate minor allele frequencies
true_af$af <- as.numeric(true_af$af)
true_af$snp_maf <- pmin(true_af$af, 1 - true_af$af)

# get key for position to k-mers
pos_kmer_key <- read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/center_kmer_pairs_938.txt", sep = ""), fill = T, col.names = c("pos", "k1", "k2", "k3", "k4"))

pos_kmer_key$leftk <- substr(pos_kmer_key$k1, 1, 15)
pos_kmer_key$rightk <- substr(pos_kmer_key$k1, 17, 31)

# compare snp and k-mer mafs
kmer_pairs$leftk <- substr(kmer_pairs$k1, 1, 15)
kmer_pairs$rightk <- substr(kmer_pairs$k1, 17, 31)

pos_kmer_pairs <- merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

# get snps that pair with reverse complements
pos_kmer_key$leftk_rc <- chartr("ATGC", "TACG", pos_kmer_key$leftk)
pos_kmer_key$rightk_rc <- chartr("ATGC", "TACG", pos_kmer_key$rightk)

pos_kmer_key$leftk_rc <- strReverse(pos_kmer_key$leftk_rc)
pos_kmer_key$rightk_rc <- strReverse(pos_kmer_key$rightk_rc)

pos_kmer_key$rightk <- pos_kmer_key$leftk_rc
pos_kmer_key$leftk <- pos_kmer_key$rightk_rc

pos_kmer_pairs_rc <- merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

# merge reverse complement and regular sets
any(duplicated(c(pos_kmer_pairs$pos, pos_kmer_pairs_rc$pos)))

pos_kmer_pairs <- rbind(pos_kmer_pairs, pos_kmer_pairs_rc[, c(-8, -9)])

# do some filtering
# remove low coverage k-mer pairs
# remove tri-allelic k-mers
pos_kmer_pairs <- pos_kmer_pairs[(pos_kmer_pairs$c1 >= 4 & pos_kmer_pairs$c2 >= 4), ]
pos_kmer_pairs <- pos_kmer_pairs[((pos_kmer_pairs$k3 == "") | is.na(pos_kmer_pairs$k3)), ]

compare_maf_kmer <- merge(true_af[, c("pos", "snp_maf")], pos_kmer_pairs, by = "pos")

# include varscan calls
varscan <- read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/calls_938.tsv", sep = ""), header = T)

varscan$snp_maf_varscan <- pmin(varscan$Reads1, varscan$Reads2) / (varscan$Reads1 + varscan$Reads2)

names(varscan)[2] <- "pos"

# compare varscan and ground truth
compare_maf_varscan <- merge(true_af[, c("pos", "snp_maf")], varscan, by = "pos")

# compare k-mer and varscan
compare_maf_kmer_varscan <- merge(compare_maf_kmer, varscan, by = "pos")

# round k-mer mafs to nearest increment
kmer_test <- cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$kmer_maf, method = "pearson")

varscan_test <- cor.test(compare_maf_varscan$snp_maf, compare_maf_varscan$snp_maf_varscan, method = "pearson")

# create plots
p1 <- ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, lwd = 2, color = "purple") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(x = "True MAF", y = "MAF from k-mers", title = paste("rho =", signif(kmer_test$estimate, 3), "  p =", paste(format(signif(kmer_test$p.value, 3), digits = 3))))

p2 <- ggplot(compare_maf_varscan, aes(x = snp_maf, y = snp_maf_varscan)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, lwd = 2, color = "purple") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(x = "True MAF", y = "MAF from varscan", title = paste("rho =", signif(varscan_test$estimate, 3), "  p =", paste(format(signif(varscan_test$p.value, 3), digits = 3))))

ggarrange(p1, p2, nrow = 2, ncol = 1, labels = c("A", "B"), font.label = list(size = 12))

ggsave(paste("../results/", today, "/true_vs_predicted_maf.png", sep = ""), height = 6, width = 3, unit = "in")
```
