---
title: "R Notebook"
output: html_notebook
---

# packages

```{r}
rm(list = ls())

library(ggplot2)
library(plyr)
library(MASS)
library(reshape)
library(ggpubr)

today = Sys.Date()
```

# functions

```{r}
# reverse a string
strReverse <- function(x) sapply(lapply(strsplit(x, NULL), rev), paste, collapse="")

# HWE heterozygosity at a single site
diversity = function(p){
  1 - (p^2 + (1-p)^2)
}

# fucntion to calculate rates and accuracy
calc_rates = function(truth, test, sample_space){
  
  negatives = sample_space[!(sample_space %in% truth)]
  positives = truth
  
  true_positives = sum(test %in% positives)
  false_positives = sum(!(test %in% positives))
  
  true_negatives = sum(!(negatives %in% test))
  false_negatives = sum(!(positives %in% test))
  
  # accuracy
  acc = (true_positives + true_negatives)/(true_positives + true_negatives + false_positives + false_negatives)
  
  # true positives rate (sensitivity)
  tpr = true_positives/(true_positives + false_negatives)

  # false positive rate (false alarm)
  fpr = false_positives/(false_positives + true_negatives)

  # true negative rate (specificity)
  tnr = true_negatives/(true_negatives + false_positives)

  # false negative rate (miss)
  fnr = false_negatives/(true_positives + false_negatives)

  return(c(length(positives), length(negatives), true_positives, false_positives, true_negatives, false_negatives, acc, tpr, fpr, tnr, fnr))
}

#calc_rates(true_af$pos, varscan$pos, 1:1e6)

#calc_rates(true_af$pos, compare_maf_kmer$pos, 1:1e6)
```

# process single population metrics
```{r}
params = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/config/parameters.tsv", header = T)

#dones = sort(as.numeric(gsub(".txt", "", gsub("discoRes_ad_", "",list.files(path = "/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow", pattern = "discoRes_ad_")))))

results = NULL

# get list of simulation ids
simids = params$ID[(params$simtype == "onepop") | (params$simtype == "sweep")]

two_pop_ids = expand.grid(params$ID[(params$simtype == "twopop")], c("p1", "p2"))

two_pop_ids = paste(two_pop_ids$Var1, two_pop_ids$Var2, sep = "_")

simids = c(two_pop_ids, simids)

#for(i in dones){
#for(i in params$ID[(params$simtype == "onepop") | (params$simtype == "sweep")]){
#for(i in 923:1000){
for(i in simids){
  print(paste("Processing simulation ", i, "...", sep = ""))
  
# load pairs of k-mers
kmer_pairs = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", "kmerpairs_", i, "_sequences.tsv", sep = ""), col.names = c("k1"))

kmer_covs = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/kmerpairs_", i, "_coverages.tsv", sep = ""), col.names = c("c1", "c2"))

kmer_pairs = cbind(kmer_pairs, kmer_covs)

# choose kmer count cutoff
# all_counts = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/kmer_counts_1.txt", col.names = c("k1", "c1"))
# 
# ggplot(all_counts, aes(x = c1)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()
# 
# ggplot(all_counts[(all_counts$c1 >= 4),], aes(x = c1)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()

# calculate minor allele frequency
kmer_pairs$kmer_maf = pmin(kmer_pairs$c1, kmer_pairs$c2)/(kmer_pairs$c1 + kmer_pairs$c2)

# ggplot(kmer_pairs, aes(x = kmer_maf)) +
#   geom_density() +
#   theme_classic()

# load in actual allele freqs from simulations
true_af = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_", i, ".txt", sep = ""), col.names = c("chrom", "pos", "ns", "af", "ac"))

true_af$ac = as.numeric(true_af$ac)

true_af = true_af[complete.cases(true_af),]

# allele frequency spectrum
# ggplot(true_af, aes(x = ac)) +
#   geom_histogram(binwidth = 1) +
#   theme_classic()

# calculate minor allele frequencies
true_af$af = as.numeric(true_af$af)
true_af$snp_maf = pmin(true_af$af, 1 - true_af$af)

# load discosnp results, calculate maf
disco = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/discoRes_ad_", i, ".txt", sep = ""), col.names = c("chrom", "pos", "c1", "c2"))

disco$disco_maf = pmin(disco$c1, disco$c2)/(disco$c1 + disco$c2)

# get key for position to k-mers
if(grepl("_p1|_p2", i)){
  j = gsub("_p1|_p2", "_p1p2", i)
} else {
  j = i
}

pos_kmer_key = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/center_kmer_pairs_", j, ".txt", sep = ""), fill = T, col.names = c("pos", "k1", "k2", "k3", "k4"))

pos_kmer_key$leftk = substr(pos_kmer_key$k1, 1, 15)
pos_kmer_key$rightk = substr(pos_kmer_key$k1, 17, 31)

# compare snp and k-mer mafs
kmer_pairs$leftk = substr(kmer_pairs$k1, 1, 15)
kmer_pairs$rightk = substr(kmer_pairs$k1, 17, 31)

pos_kmer_pairs = merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

# get snps that pair with reverse complements
pos_kmer_key$leftk_rc = chartr("ATGC","TACG",pos_kmer_key$leftk)
pos_kmer_key$rightk_rc = chartr("ATGC","TACG",pos_kmer_key$rightk)

pos_kmer_key$leftk_rc = strReverse(pos_kmer_key$leftk_rc)
pos_kmer_key$rightk_rc = strReverse(pos_kmer_key$rightk_rc)

pos_kmer_key$rightk = pos_kmer_key$leftk_rc 
pos_kmer_key$leftk = pos_kmer_key$rightk_rc

pos_kmer_pairs_rc = merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

# merge reverse complement and regular sets
any(duplicated(c(pos_kmer_pairs$pos, pos_kmer_pairs_rc$pos)))

pos_kmer_pairs = rbind(pos_kmer_pairs, pos_kmer_pairs_rc[,c(-8,-9)])

# do some filtering
# remove low coverage k-mer pairs
# remove tri-allelic k-mers
pos_kmer_pairs = pos_kmer_pairs[(pos_kmer_pairs$c1 >=3 & pos_kmer_pairs$c2 >= 3),]
pos_kmer_pairs = pos_kmer_pairs[( (pos_kmer_pairs$k3 == "") | is.na(pos_kmer_pairs$k3)),]

compare_maf_kmer = merge(true_af[,c("pos", "snp_maf")], pos_kmer_pairs, by = "pos")

# include varscan calls
varscan = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/calls_", i, ".tsv", sep = ""), header = T)

varscan$snp_maf_varscan = pmin(varscan$Reads1, varscan$Reads2)/(varscan$Reads1 + varscan$Reads2)

names(varscan)[2] = "pos"

# load poolsnp calls
poolsnp = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/poolsnp_final_", i, ".txt", sep = ""), col.names = c("chrom", "pos", "c1", "t"), fill = T, na.strings = "")

poolsnp$snp_maf_poolsnp = pmin(poolsnp$c1/poolsnp$t, 1 - poolsnp$c1/poolsnp$t)

# compare varscan and ground truth
compare_maf_varscan = merge(true_af[,c("pos", "snp_maf")], varscan, by = "pos")

# compare poolsnp and ground truth
compare_maf_poolsnp = merge(true_af[,c("pos", "snp_maf")], poolsnp, by = "pos")

# compare discosnp and ground truth
compare_maf_disco = merge(true_af[,c("pos", "snp_maf")], disco, by = "pos")

# compare k-mer and varscan
compare_maf_kmer_varscan = merge(compare_maf_kmer, varscan, by = "pos")

# calculate correlations and performance stats
kmer_vs_truth_maf_cor = cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$kmer_maf, method = "pearson")$estimate

disco_vs_truth_maf_cor = tryCatch(
  cor.test(compare_maf_disco$snp_maf, compare_maf_disco$disco_maf, method = "pearson")$estimate, 
  error=function(err) NA)

varscan_vs_truth_maf_cor = cor.test(compare_maf_varscan$snp_maf, compare_maf_varscan$snp_maf_varscan, method = "pearson")$estimate

poolsnp_vs_truth_maf_cor = tryCatch(
  cor.test(compare_maf_poolsnp$snp_maf, compare_maf_poolsnp$snp_maf_poolsnp, method = "pearson")$estimate, 
  error=function(err) NA)

varscan_rates = calc_rates(true_af$pos, varscan$pos, 1:1e6)

poolsnp_rates = calc_rates(true_af$pos, poolsnp$pos, 1:1e6)

kmer_rates = calc_rates(true_af$pos, pos_kmer_pairs$pos, 1:1e6)

disco_rates = calc_rates(true_af$pos, disco$pos, 1:1e6)

# save all results
result = data.frame(
  ID = gsub("_p1|_p2", "", i),
  ID2 = i,
  kmer_vs_truth_maf_cor = kmer_vs_truth_maf_cor,
  kmer_pos = kmer_rates[1],
  kmer_neg = kmer_rates[2],
  kmer_tp = kmer_rates[3],
  kmer_fp = kmer_rates[4],
  kmer_tn = kmer_rates[5],
  kmer_fn = kmer_rates[6],
  kmer_acc = kmer_rates[7],
  kmer_tpr = kmer_rates[8],
  kmer_fpr = kmer_rates[9],
  kmer_tnr = kmer_rates[10],
  kmer_fnr = kmer_rates[11],
  disco_vs_truth_maf_cor = disco_vs_truth_maf_cor,
  disco_pos = disco_rates[1],
  disco_neg = disco_rates[2],
  disco_tp = disco_rates[3],
  disco_fp = disco_rates[4],
  disco_tn = disco_rates[5],
  disco_fn = disco_rates[6],
  disco_acc = disco_rates[7],
  disco_tpr = disco_rates[8],
  disco_fpr = disco_rates[9],
  disco_tnr = disco_rates[10],
  disco_fnr = disco_rates[11],
  varscan_vs_truth_maf_cor = varscan_vs_truth_maf_cor,
  varscan_pos = varscan_rates[1],
  varscan_neg = varscan_rates[2],
  varscan_tp = varscan_rates[3],
  varscan_fp = varscan_rates[4],
  varscan_tn = varscan_rates[5],
  varscan_fn = varscan_rates[6],
  varscan_acc = varscan_rates[7],
  varscan_tpr = varscan_rates[8],
  varscan_fpr = varscan_rates[9],
  varscan_tnr = varscan_rates[10],
  varscan_fnr = varscan_rates[11],
  poolsnp_vs_truth_maf_cor = poolsnp_vs_truth_maf_cor,
  poolsnp_pos = poolsnp_rates[1],
  poolsnp_neg = poolsnp_rates[2],
  poolsnp_tp = poolsnp_rates[3],
  poolsnp_fp = poolsnp_rates[4],
  poolsnp_tn = poolsnp_rates[5],
  poolsnp_fn = poolsnp_rates[6],
  poolsnp_acc = poolsnp_rates[7],
  poolsnp_tpr = poolsnp_rates[8],
  poolsnp_fpr = poolsnp_rates[9],
  poolsnp_tnr = poolsnp_rates[10],
  poolsnp_fnr = poolsnp_rates[11]
)

results = rbind(results, result)

rm(compare_maf_kmer, compare_maf_kmer_varscan, compare_maf_varscan, compare_maf_poolsnp, poolsnp, poolsnp_rates, compare_maf_disco, disco, disco_rates, kmer_covs, kmer_pairs, pos_kmer_key, pos_kmer_pairs, pos_kmer_pairs_rc, true_af, varscan, kmer_rates, kmer_vs_truth_maf_cor, poolsnp_vs_truth_maf_cor, varscan_rates, varscan_vs_truth_maf_cor, disco_vs_truth_maf_cor, result)

}

pares = merge(params, results, by = "ID")

write.table(pares, paste("/mnt/ufs18/home-010/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today, "/performance_metrics.txt",sep = ""), sep = " ")
```

# selective sweeps
```{r}
# Hypothesis: the average distance between SNPs is higher after selective sweep compared to neutral scenario
allele_files = list.files(path = "/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow", pattern = "slim_allele_freqs_.*", full.names = T)

allele_files = allele_files[!grepl("_p.", allele_files)]

allele_file_ids = as.numeric(gsub(".txt", "", gsub("slim_allele_freqs_", "", gsub("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", "", allele_files))))

allele_files = lapply(allele_files, read.table, col.names = c("chrom", "pos", "ns", "af", "ac"))

mean_dist_bw_mut = function(x){
  mean(diff(x$pos))
}

plotdata = data.frame(
  ID = allele_file_ids,
  mean_dist = unlist(lapply(allele_files, mean_dist_bw_mut))
)

plotdata = merge(params, plotdata, by = "ID")
plotdata$n = as.factor(plotdata$n)
plotdata$cov = as.factor(plotdata$cov)

ggplot(plotdata, aes(x = n, y = mean_dist, color = simtype)) +
  geom_boxplot() +
  theme_classic() +
  stat_compare_means(aes(group = simtype), label.sep = "\n")
    
# Hypothesis: after a selective sweep, the longest unitigs in a population graph are more likely to be near the sweep site, compared to neutral scenario
unitig_files = list.files(path = "/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow", pattern = "unitig_alignments_.*", full.names = T)

unitig_files_ids = as.numeric(gsub(".txt", "", gsub("unitig_alignments_", "", gsub("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", "", unitig_files))))

unitig_files = lapply(unitig_files, read.table, col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

mean_len = function(x){
  mean(x$alen)
}

plotdata = data.frame(
  ID = unitig_files_ids,
  avg_uni_len = unlist(lapply(unitig_files, mean_len))
)


plotdata = merge(params, plotdata, by = "ID")
plotdata$n = as.factor(plotdata$n)
plotdata$cov = as.factor(plotdata$cov)

plotdata = plotdata[( ((plotdata$s == 0.5) & (plotdata$h == 1)) | (plotdata$s == 0) ),]

ggplot(plotdata, aes(x = n, y = avg_uni_len, color = simtype, fill = cov)) +
  geom_boxplot() +
  facet_grid(sequencer ~ .) +
  theme_classic()
  #stat_compare_means(aes(group = simtype), label.sep = "\n")







check_overlap = function(x, nhits, target, tol){
  # sort by unitig length
  x_sorted = x[order(-x$alen),]
  
  # get top n longest unitigs
  x_top = x_sorted[1:nhits,]
  
  x_top$start = pmin(x_top$ss, x_top$se)
  x_top$end = pmax(x_top$ss, x_top$se)
  
  targ_start = target - tol
  targ_end = target + tol
  
  result = any( (targ_start < x_top$start) & (x_top$start < targ_end) ) | any( (targ_start < x_top$end) &  (x_top$end < targ_end) )
  
  return(result)
}

plotdata = data.frame(
  ID = unitig_files_ids,
  overlap = unlist(lapply(unitig_files, check_overlap, nhits = 10, target = 500e3, tol = 50e3))
)

plotdata = merge(params, plotdata, by = "ID")
plotdata$n = as.factor(plotdata$n)
plotdata$cov = as.factor(plotdata$cov)

aggregate(overlap ~ s, data = plotdata, FUN = mean)

ggplot(plotdata, aes(x = overlap, fill = simtype)) +
  geom_bar(stat = "count", position="dodge") +
  facet_grid(s ~ .) +
  theme_classic()










allele = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_277.txt", col.names = c("chrom", "pos", "ns", "af", "ac"))

mean(diff(allele$pos))

unitigs = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/unitig_alignments_277.txt", col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

ggplot(unitigs, aes(x = alen)) +
  geom_density()

mean(unitigs$alen)

allele = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_5558.txt", col.names = c("chrom", "pos", "ns", "af", "ac"))

mean(diff(allele$pos))

unitigs = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/unitig_alignments_5558.txt", col.names = c("qid", "sid", "pid", "alen", "mm", "go", "qs", "qe", "ss", "se", "ev", "bs"))

ggplot(unitigs, aes(x = alen)) +
  geom_density()

mean(unitigs$alen)

sam_pos = seq(from = 1000, to = 1e6, by = 1000)

results = NULL
i = 0
while(i <= 1e6){
  result = mean(unitigs[((unitigs$ss <= i) & (i + 1000 <= unitigs$se)),"alen"])
  results = rbind(results, data.frame(start = i, end = i + 1000, avglen = result))
  i = i + 1000
}

ggplot(results, aes(x = (start + end)/2, y = avglen)) +
  geom_point() +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = 5e5) +
  theme_classic()

results = c()
for(pos in sam_pos){
  result = mean(unitigs[((unitigs$ss <= pos) & (pos <= unitigs$se)),"alen"])
  results = c(results, result)
}

plotdata = data.frame(
  pos = sam_pos,
  avglen = results
)

ggplot(plotdata, aes(x = pos, y = avglen)) +
  geom_point() +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = 5e5) +
  theme_classic()

ggplot(unitigs, aes(x = log10(ss), y = alen)) +
  geom_point() +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = log10(5e5)) +
  theme_classic()

```

# process two population metrics (Fst)
```{r}

```

# How does performance vary with sequencing parameters?
```{r}
# compare sensitivities for SNPs represented in reference
tpr_plot_data = melt(pares[,c("ID", "ID2", "cov", "n", "sequencer", "simtype", "kmer_tpr", "disco_tpr", "varscan_tpr", "poolsnp_tpr")], id = c("ID", "ID2", "cov", "n", "sequencer", "simtype"))

tpr_plot_data$variable = factor(gsub("_tpr", "", tpr_plot_data$variable), levels = c("kmer", "disco", "varscan", "poolsnp"), ordered = T)

tpr_plot_data$cov_per_n = as.factor(tpr_plot_data$cov/(tpr_plot_data$n * 2))

tpr_plot_data$cov = as.factor(tpr_plot_data$cov)
tpr_plot_data$n = as.factor(tpr_plot_data$n)

ggplot(tpr_plot_data, aes(x = cov_per_n, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Coverage per haploid genome", y = "Sensitivity (true positive rate)", color = "Method", title = "All instruments")

ggsave(paste("../results/", today, "/coverage_vs_sensitivity.png", sep = ""), unit = "in")

ggplot(tpr_plot_data, aes(x = cov_per_n, y = value, color = simtype)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Coverage per haploid genome", y = "Sensitivity (true positive rate)", color = "Method", title = "All instruments")

# coverage vs false discovery rate
fpr_plot_data = melt(pares[,c("ID", "cov", "n", "sequencer", "kmer_fpr", "disco_fpr", "varscan_fpr", "poolsnp_fpr")], id = c("ID", "cov", "n", "sequencer"))

fpr_plot_data$variable = factor(gsub("_fpr", "", fpr_plot_data$variable), levels = c("kmer", "disco", "varscan", "poolsnp"), ordered = T)

fpr_plot_data$cov_per_n = as.factor(fpr_plot_data$cov/(fpr_plot_data$n * 2))

ggplot(fpr_plot_data, aes(x = cov_per_n, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Coverage per haploid genome", y = "False positive rate", color = "Method", title = "All instruments")

ggsave(paste("../results/", today, "/coverage_vs_fpr.png", sep = ""), unit = "in")

# coverage vs false positives
fp_plot_data = melt(pares[,c("ID", "cov", "n", "sequencer", "kmer_fp", "disco_fp", "varscan_fp", "poolsnp_fp")], id = c("ID", "cov", "n", "sequencer"))

fp_plot_data$variable = factor(gsub("_fp", "", fp_plot_data$variable), levels = c("kmer", "disco", "varscan", "poolsnp"), ordered = T)

fp_plot_data$cov_per_n = as.factor(fp_plot_data$cov/(fp_plot_data$n * 2))

ggplot(fp_plot_data, aes(x = cov_per_n, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 12)) +
  labs(x = "Coverage per haploid genome", y = "False positives", color = "Method", title = "All instruments")

ggsave(paste("../results/", today, "/coverage_vs_fp.png", sep = ""), unit = "in")

#
tpr_vs_fpr = merge(tpr_plot_data, fpr_plot_data, by = c("ID", "cov", "n", "sequencer", "variable", "cov_per_n"))

ggplot(tpr_vs_fpr, aes(x = value.y, y = value.x, color = variable)) +
  geom_point() +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "False positive rate", y = "True positive rate", color = "Method", title = "All instruments")


ggplot(tpr_plot_data[(tpr_plot_data$n == 50 & tpr_plot_data$sequencer == "miseq"),], aes(x = cov, y = value, color = variable)) +
  geom_boxplot() +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "Coverage", y = "Sensitivity (true positive rate)", color = "Method", title = "pool size = 50; MiSeq")

ggsave(paste("../results/", today, "/sensitivity_50_miseq.png", sep = ""), height = 7, width = 8)

# hold pool size constant and increase coverage
ggplot(pares[(pares$n == 5),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 10),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 25),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 50),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 100),], aes(x = cov, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

# hold coverage constant, and add more individuals to pool
ggplot(pares[(pares$cov == 20),], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$cov == 50),], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$cov == 80),], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$cov == 100),], aes(x = n, y = kmer_vs_truth_maf_cor)) +
  geom_boxplot() +
  geom_point()

# hold pool size constant, increase coverage
ggplot(pares[(pares$n == 5),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 10),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 25),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 50),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()

ggplot(pares[(pares$n == 100),], aes(x = cov, y = kmer_tpr)) +
  geom_boxplot() +
  geom_point()





ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "True MAF", y = "Predicted MAF from k-mers")

ggplot(compare_maf_kmer, aes(x = snp_maf, y = snp_maf_varscan)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "True MAF", y = "Predicted MAF from VarScan")

# what about for k-mers above some frequency threshold
kmer_test = cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$kmer_maf, method = "spearman")

ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "True MAF", y = "Predicted MAF from k-mers", title = paste("rho =", signif(kmer_test$estimate, 4), "; p =", paste(format(signif(kmer_test$p.value, 2), digits = 2))))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","kmer_maf_vs_truth_pool-seq.png", sep = ""), height= 7, width = 7)

varscan_test = cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$snp_maf_varscan, method = "spearman")

ggplot(compare_maf_kmer, aes(x = snp_maf, y = snp_maf_varscan)) +
  geom_density_2d_filled() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "True MAF", y = "Predicted MAF from VarScan", title = paste("rho =", signif(varscan_test$estimate, 4), "; p =", paste(format(signif(varscan_test$p.value, 2), digits = 2))))

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","varscan_maf_vs_truth_pool-seq.png", sep = ""), height= 7, width = 8)

# smudgeplot
ggplot(compare_maf_kmer, aes(x = kmer_maf, y = (c1+c2)/81.670932)) +
  geom_density_2d_filled() +
  #geom_hline(yintercept = 50) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "Minor k-mer frequency", y = "Total pair coverage")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","smudgeplot.png", sep = ""), height= 7, width = 8)

# site-frequency spectrum
plotdata = melt(as.data.table(compare_maf_kmer[,c("pos", "snp_maf", "kmer_maf", "snp_maf_varscan")]), id = "pos")

ggplot(plotdata, aes(x = value, fill = variable)) +
  geom_density(alpha = 0.5) +
  theme_classic()

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","kmer_vs_snp_sfs.png", sep = ""), height= 7, width = 8)

# folded sfs
kmer_pairs = kmer_pairs[(kmer_pairs$c1 >=4 & kmer_pairs$c2 >= 4),]

folded_sfs = data.frame(
    category=factor(c(rep("truth", times = nrow(true_af)), rep("kmer", times = nrow(kmer_pairs)), rep("varscan", times = nrow(varscan)))),
    value=c(true_af$snp_maf,kmer_pairs$kmer_maf, varscan$snp_maf_varscan)
)

ggplot(folded_sfs, aes(x = value, fill = category)) +
  geom_density(alpha = 0.5) +
  theme_classic() +
  theme(text = element_text(size = 20)) +
  labs(x = "minor allele frequency")

ggsave(paste("/mnt/home/robe1195/Josephs_Lab_Projects/poolseq-kmers/results/", today,  "/","kmer_vs_snp_sfs.png", sep = ""), height= 7, width = 8)

ggplot(true_af, aes(x = snp_maf)) +
  geom_density() +
  theme_classic()

ggplot(kmer_pairs, aes(x = kmer_maf)) +
  geom_density() +
  theme_classic()

ggplot(kmer_pairs, aes(x = kmer_maf)) +
  geom_density() +
  theme_classic()

ggplot(true_af[(true_af$snp_maf >= 0.05),], aes(x = snp_maf)) +
  geom_density() +
  theme_classic()

ggplot(compare_maf_kmer, aes(x = kmer_maf)) +
  geom_histogram(binwidth = 0.05) +
  theme_classic()

ggplot(compare_maf_kmer, aes(x = snp_maf)) +
  geom_histogram(breaks = seq(from = 0, to = 0.5, by = 0.02)) +
  theme_classic()

ggplot(compare_maf_kmer, aes(x = kmer_maf)) +
  geom_histogram(breaks = seq(from = 0, to = 0.5, by = 0.02)) +
  theme_classic()

# calculate error
compare_maf_kmer$error = abs(compare_maf_kmer$kmer_maf - compare_maf_kmer$snp_maf)

summary(compare_maf_kmer$error)

sum(compare_maf_kmer$error < 0.1)/nrow(compare_maf_kmer)

ggplot(confident_sites, aes(x = total_cov, y = error)) +
  geom_point() +
  theme_classic()

#
compare_maf_kmer$snp_maf_bins = cut(compare_maf_kmer$snp_maf, breaks = seq(from = 0, to = 0.5, by = 0.05))
compare_maf_kmer$kmer_maf_bins = cut(compare_maf_kmer$kmer_maf, breaks = seq(from = 0, to = 0.5, by = 0.05))

observed <- table(compare_maf_kmer$kmer_maf_bins)  # Observed frequencies
expected <- table(compare_maf_kmer$snp_maf_bins)  # Expected frequencies

observed = observed[(expected != 0)]
expected = expected[(expected != 0)]

# Calculate Chi-Square statistic manually
chi_sq_statistic <- sum((observed - expected)^2 / expected)
df <- length(observed) - 1
p_value <- 1 - pchisq(chi_sq_statistic, df)

# Print results
print(paste("Chi-Square Statistic:", chi_sq_statistic))
print(paste("Degrees of Freedom:", df))
print(paste("P-value:", p_value))

```

# correlations between true and predicted minor allele frequencies
```{r}
# load pairs of k-mers
kmer_pairs = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/", "kmerpairs_938_sequences.tsv", sep = ""), col.names = c("k1"))

kmer_covs = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/kmerpairs_938_coverages.tsv", sep = ""), col.names = c("c1", "c2"))

kmer_pairs = cbind(kmer_pairs, kmer_covs)

# calculate minor allele frequency
kmer_pairs$kmer_maf = pmin(kmer_pairs$c1, kmer_pairs$c2)/(kmer_pairs$c1 + kmer_pairs$c2)

# load in actual allele freqs from simulations
true_af = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_938.txt", sep = ""), col.names = c("chrom", "pos", "af", "ac"))

true_af$ac = as.numeric(true_af$ac)

true_af = true_af[complete.cases(true_af),]

# calculate minor allele frequencies
true_af$af = as.numeric(true_af$af)
true_af$snp_maf = pmin(true_af$af, 1 - true_af$af)

# get key for position to k-mers
pos_kmer_key = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/center_kmer_pairs_938.txt", sep = ""), fill = T, col.names = c("pos", "k1", "k2", "k3", "k4"))

pos_kmer_key$leftk = substr(pos_kmer_key$k1, 1, 15)
pos_kmer_key$rightk = substr(pos_kmer_key$k1, 17, 31)

# compare snp and k-mer mafs
kmer_pairs$leftk = substr(kmer_pairs$k1, 1, 15)
kmer_pairs$rightk = substr(kmer_pairs$k1, 17, 31)

pos_kmer_pairs = merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

# get snps that pair with reverse complements
pos_kmer_key$leftk_rc = chartr("ATGC","TACG",pos_kmer_key$leftk)
pos_kmer_key$rightk_rc = chartr("ATGC","TACG",pos_kmer_key$rightk)

pos_kmer_key$leftk_rc = strReverse(pos_kmer_key$leftk_rc)
pos_kmer_key$rightk_rc = strReverse(pos_kmer_key$rightk_rc)

pos_kmer_key$rightk = pos_kmer_key$leftk_rc 
pos_kmer_key$leftk = pos_kmer_key$rightk_rc

pos_kmer_pairs_rc = merge(pos_kmer_key, kmer_pairs, by = c("leftk", "rightk"))

# merge reverse complement and regular sets
any(duplicated(c(pos_kmer_pairs$pos, pos_kmer_pairs_rc$pos)))

pos_kmer_pairs = rbind(pos_kmer_pairs, pos_kmer_pairs_rc[,c(-8,-9)])

# do some filtering
# remove low coverage k-mer pairs
# remove tri-allelic k-mers
pos_kmer_pairs = pos_kmer_pairs[(pos_kmer_pairs$c1 >=4 & pos_kmer_pairs$c2 >= 4),]
pos_kmer_pairs = pos_kmer_pairs[( (pos_kmer_pairs$k3 == "") | is.na(pos_kmer_pairs$k3)),]

compare_maf_kmer = merge(true_af[,c("pos", "snp_maf")], pos_kmer_pairs, by = "pos")

# include varscan calls
varscan = read.table(paste("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/calls_938.tsv", sep = ""), header = T)

varscan$snp_maf_varscan = pmin(varscan$Reads1, varscan$Reads2)/(varscan$Reads1 + varscan$Reads2)

names(varscan)[2] = "pos"

# compare varscan and ground truth
compare_maf_varscan = merge(true_af[,c("pos", "snp_maf")], varscan, by = "pos")

# compare k-mer and varscan
compare_maf_kmer_varscan = merge(compare_maf_kmer, varscan, by = "pos")

# round k-mer mafs to nearest increment 
kmer_test = cor.test(compare_maf_kmer$snp_maf, compare_maf_kmer$kmer_maf, method = "pearson")

varscan_test = cor.test(compare_maf_varscan$snp_maf, compare_maf_varscan$snp_maf_varscan, method = "pearson")

# create plots
p1 = ggplot(compare_maf_kmer, aes(x = snp_maf, y = kmer_maf)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, lwd = 2, color = "purple") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(x = "True MAF", y = "MAF from k-mers", title = paste("rho =", signif(kmer_test$estimate, 3), "  p =", paste(format(signif(kmer_test$p.value, 3), digits = 3))))

p2 = ggplot(compare_maf_varscan, aes(x = snp_maf, y = snp_maf_varscan)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, lwd = 2, color = "purple") +
  theme_classic() +
  theme(text = element_text(size = 10)) +
  labs(x = "True MAF", y = "MAF from varscan", title = paste("rho =", signif(varscan_test$estimate, 3), "  p =", paste(format(signif(varscan_test$p.value, 3), digits = 3))))

ggarrange(p1, p2, nrow = 2, ncol = 1, labels = c("A", "B"), font.label = list(size = 12))

ggsave(paste("../results/", today, "/true_vs_predicted_maf.png", sep = ""), height = 6, width = 3, unit = "in")
```

```{r}
kmer_pairs = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/testpairs_sequences.tsv", col.names = c("k1"))

kmer_counts = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/testpairs_coverages.tsv", col.names = c("c1", "c2"))

kmer_pairs = cbind(kmer_pairs, kmer_counts)

# calculate minor allele frequency
nidv = 5
ploidy = 2

kmer_pairs$kmer_maf = round(pmin(kmer_pairs$c1, kmer_pairs$c2)/(kmer_pairs$c1 + kmer_pairs$c2)*nidv*ploidy)

# filter out k-mers below coverage thresholds
kmer_pairs = kmer_pairs[( (kmer_pairs$c1 > 3) & (kmer_pairs$c2 > 3) ),]



kmer_counts = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/kmer_counts_161.txt", col.names = c("k1", "c1"))

# frequency vs coverage
ggplot(kmer_counts[(kmer_counts$c1 >=4 ),], aes(x = c1)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 4) +
  theme_classic()

# frequency*coverage vs coverage
freq = table(kmer_counts$c1)

cov = as.numeric(names(freq))
freq = as.vector(freq)

trans_spectrum = data.frame(
  freq = freq,
  freq_cov = freq*cov,
  cov = cov
)

ggplot(trans_spectrum, aes(x = cov, y = freq_cov)) +
  geom_bar(stat = "identity") +
  theme_classic()

#
testmers = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/testmers.txt", col.names = c("k1", "c1"))

table(testmers$c1)/nrow(testmers) * 100

# try modeling coverage of a k-mer based on it's frequency in a pool
lamb = 60 # average haploid sequencing coverage
rho = 0.8 # dispersion due to sequencing biases
i = 3 # number of genomes where k-mer is present
n = 60 # number of genomes in pool

plotdata = data.frame(
  x = rnbinom(1000, mu = (i/n)*lamb, size = (i/n)*(lamb/rho))
)

ggplot(plotdata, aes(x)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = (i/n)*lamb)

# try fitting poisson to distribution
fitdistr(kmer_counts$c1, "poisson", method = "L-BFGS-B", lower = 0.1, upper = 0.5)

fitdistr(kmer_counts$c1[], "negative binomial")

ggplot(trans_spectrum, aes(x = cov, y = freq/sum(freq))) +
  geom_bar(stat = "identity") +
  geom_point(aes(y=dpois(cov, 6.7055415264)), colour="red") +
  theme_classic()

ggplot(kmer_counts, aes(x = c1)) +
  geom_histogram(binwidth = 1) +
  geom_point(aes(y=dpois(c1, 1.0344408452)), colour="red") +
  geom_vline(xintercept = 4) +
  theme_classic()

ggplot(data.frame(x=c(0:10)), aes(x)) +
    geom_point(aes(y=dpois(x, 1)), colour="red")
```

#
```{r}
disco = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/discoRes_840_ad.txt", col.names = c("CHROM", "POS", "c1", "c2"))

slim = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/poolseq-kmers/workflow/slim_allele_freqs_840.txt", col.names = c("CHROM", "POS", "snp_af", "ac"))

disco$kmer_min_af = pmin(disco$c1, disco$c2)/(disco$c1 + disco$c2)

slim$snp_min_af = pmin(slim$snp_af, 1 - slim$snp_af)

all_min_af = merge(slim, disco, by = c("CHROM", "POS"), all.x = T, all.y = T)

cor.test(all_min_af$snp_min_af, all_min_af$kmer_min_af, method = "spearman")

ggplot(all_min_af, aes(x = snp_min_af, y = kmer_min_af)) +
  geom_point() +
  theme_classic() +
  geom_abline(slope = 1, intercept = 0)
```

# DEPRECATED
```{r}
kmer_counts = read.table("/mnt/scratch/robe1195/Josephs_Lab_Projects/kmers_poolseq/kmer_counts_1.txt", col.names = c("k1", "c1"))

ggplot(kmer_counts, aes(x = c1)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 5) +
  theme_classic()

# loop over k-mers to find ones that putative snps
alphabet = c("A","T","G","C")

results = list()

j = 1
while(nrow(kmer_counts) > 0){
  # get focal k-mer
  focal_kmer = kmer_counts[1,"k1"]
  focal_counts = kmer_counts[1,"c1"]
  print(focal_kmer)

  # generate variant k-mers
  central_base = substr(focal_kmer, 16, 16)

  var_bases = alphabet[!(alphabet %in% central_base)]

  variant_kmers = rep(focal_kmer, times = 3)

  for(i in 1:3){
    var_base = var_bases[i]
    substr(variant_kmers[i], 16, 16) = var_base
  }

  # add reverse complements
  variant_kmers = c(variant_kmers, chartr("ATGC","TACG",variant_kmers))

  # 
  variant_kmers_index = which(kmer_counts$k1 %in% variant_kmers)
  print(variant_kmers_index)

  # store 
  found_variant_kmers = kmer_counts[variant_kmers_index,"k1"]
  found_variant_counts = kmer_counts[variant_kmers_index,"c1"]
  
  if(!identical(found_variant_kmers, character(0))){
      results[[j]] = data.frame(focal_kmer,
                            focal_counts,
                            found_kmers = paste(found_variant_kmers, sep = ";"),
                            found_counts = paste(found_variant_counts, sep = ";"))

    j = j + 1
  }

  # remove focal and variant k-mers from count table
  kmer_counts = kmer_counts[c(-1, -1*variant_kmers_index),]

}


```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
